---
title: "Predictive Modelling of Hearing Effect of Bioage"
author: "Landy Qu & Tom Naber"
date: "`r format(Sys.Date(), '%B %d, %Y')`"

knit: 
  (function(inputFile, encoding) { 
    source(here::here('Scripts','rmd_param_knitter.R'));
  
    filename_constructor <- function(params, inputFile) {
      parts <- c(
        params$hearing_var,
        paste0("k=", params$global_k),
        if (params$global_discrete) "discrete" else NULL,
        if (params$telolength) "telolength" else NULL,
        paste0(params$global_repeats, "r×", params$global_folds, "f"),
        if (params$run_optional_chunks) "incl_optional" else NULL
      );
      
      here::here("Output", "HTML", paste(parts, collapse = "_"));
    };
  
    smart_knitter(inputFile, out_path = filename_constructor, auto_close = TRUE);
  })
  
output:
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    df_ppi_print: paged
    highlight: tango
    self_contained: true
    fig_width: 7
    fig_height: 4
    code_folding: hide
    
editor_options:
  markdown:
    wrap: 80
  chunk_output_type: inline
  
params:
  hearing_var:
    label: "Hearing predictor variable"
    value: "FI_high"
    input: select
    choices: ["FI_high", "FI_low"]
    help: "Set predictor variable"
  telolength:
    label: "Analyze telomere length"
    input: checkbox
    value: FALSE
    help: "Check to use telolength dataset with smaller N"
  global_k:
    label: "K (n BAM base functions)"
    value: 50
    input: numeric
    help: "Set GAM wiggliness; higher K improves fit at higher computational cost"
  global_discrete: 
    label: "Discretize BAM"
    value: TRUE
    input: checkbox
    help: "Check to speed up fitting but create larger models that perform a little worse"
  global_folds: 
    label: "Number of folds"
    value: 10
    input: numeric
    help: "Set fold number"
  global_repeats: 
    label: "Number of repeats"
    value: 10
    input: numeric
    help: "Set repeat number"
  store_model_results:
    label: "Store model results"
    value: FALSE
    input: checkbox
    help: "Set to save large model results locally for fast reloading"
  load_model_results:
    label: "Load model results"
    value: TRUE
    input: checkbox
    help: "Set to skip refitting of k-fold validation if model results are present"
  keep_models:
    label: "Store models"
    value: FALSE
    input: checkbox
    help: "Set to store models, creating large files and filling memory if discrete = TRUE"
  confounders:
    label: "Included confounders"
    value: "gender_smoking_race_education_gender, race"
    input: text
    help: "Set confounder factor variables and interaction to be included in analysis separated by underscore"
  n_boot:
    label: "RMSE bootstrap N"
    value: 500
    input: numeric
    help: "Set number of samples to take when bootstrapping RMSEs for CI calculations"
  construct_data:
    label: "Build NHANES dataframe"
    value: FALSE
    input: checkbox
    help: "Set to import NHANES data from online website and build dataframe from scratch"
  run_optional_chunks:
    label: "Run tests and validations"
    value: FALSE
    input: checkbox
    help: "Set to run tests and validations, which can take a long time to compute"
  save_figures:
    label: "Save figures"
    value: FALSE
    input: checkbox
    help: "Set to store figures and overwrite old figures"
  single_column_width:
    label: "Single column figure width"
    value: 3.19
    input: numeric
    help: "Set single column figure width in inches"
  double_column_width:
    label: "Double column figure width"
    value: 6.96
    input: numeric   
    help: "Set double column figure width in inches"
  plot_output_type:
    label: "Set plot theme"
    value: "paper"
    input: radio
    choices: ["paper", "poster"]
    help: "Set theme preset to use for plots"
---

# Set-up

We set graphics device to ragg, set wd to here::here(), set contrasts by default to compare each level to the overall mean as opposed to reference, and import packages. The run_module() function checks if knitting or interactive session is in progress. If knitting, return markdown output, if interactive, only run code in environment and print a time message.

```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(
  fig.width = 6.4,
  fig.height = 4.8,
  dpi = 600,
  fig.retina = 1,
  out.width = "100%",
  out.height = "100%",
  dev = "ragg_png"
)

knitr::opts_knit$set(root.dir = here::here())

options(
  contrasts = c("contr.sum", "contr.poly"),
  knitr.duplicate.label = "allow"
)

run_module <- function(path) {
  if (isTRUE(getOption("knitr.in.progress"))) {
    cat(knitr::knit_child(path, quiet = TRUE))
  } else {
    runtime <- system.time(knitr::knit_child(path, quiet = TRUE))
    cat(basename(path), " completed in ", round(runtime["elapsed"], 1), "s.\n", sep = "")
  }
}
```

# Run modules

```{r run_modules, results='asis', echo=FALSE}
run_module(here::here("Scripts", "01_import.Rmd"))
run_module(here::here("Scripts", "02_data_preprocessing.Rmd"))
run_module(here::here("Scripts", "03_modeling.Rmd"))
run_module(here::here("Scripts", "04_data_processing.Rmd"))
run_module(here::here("Scripts", "05_statistics.Rmd"))
run_module(here::here("Scripts", "06_visualization.Rmd"))
```
