---
editor_options:
  markdown:
    wrap: 100
---

# Data Preprocessing

## Data cleaning

We make a list of files, store the year and filename from each file, then append the data to the
row. Next, we take all the files from one year and join the rows if they match or columns if they
don't match according to SEQN (ID) and create a `year` column and `cohort` column that takes the
first 4 characters from `year`. `imap_dfr()` applies this to each year (map), returns the index (i)
and row-binds into a dataframe (dfr).

```{r}
import_nhanes_data <- function(data_path, column_path, column_name_path) {
  files <- list.files(
    data_path,
    pattern = "\\.xpt$",
    full.names = TRUE,
    ignore.case = TRUE
  )
  # Read columns to select from text file
  col_spec <- readLines(column_path)

  # Build column spec sheet with metadata to select columns on
  col_spec <- tibble(line = col_spec) %>%
    mutate(
      dataset = str_extract(line, "^[^_]+"),
      year_group = str_extract(line, "\\d{4}"),
      cols = str_extract(line, "(?<=select\\().*(?=\\))") %>% str_split(",\\s*")
    ) %>%
      unnest(., cols)  %>% select(-line)
  
  # 
  # %>%
  #   group_by(year_group) %>%
  #   summarize(
  #     select_cols = list(unique(unlist(cols))),
  #     .groups = "drop"
  #   )
  # 
  col_name_spec <- readLines(column_name_path)

  browser()

  # Load data with metadata from filename
  data_map <- tibble(
    file = files,
    year = str_extract(
      tools::file_path_sans_ext(
        basename(files)
      ),
      "\\d{4}-\\d{4}$"
    ),
    cohort = substr(year, 1, 4),
    data = map(
      file, ~ read_xpt(.x)
    )
  )

  # Merge data from the same year into one dataframe
  data_map <- data_map %>%
    group_by(cohort) %>%
    summarise(
      data = list(reduce(data, full_join, by = "SEQN")),
      .groups = "drop"
    )

  # Select columns to keep per year
  data_map <- data_map %>%
    mutate(
      data = map2(data, cohort, \(df, yr) {
        cols_to_keep <- unlist(
          col_spec %>%
            filter(year_group == yr) %>%
            pull(select_cols)
        )

        df %>%
          select(all_of(cols_to_keep))
      })
    )


  browser()
}

```

```{r}
df_raw <- import_nhanes_data(
  data_path = here("Data", "Raw data"),
  column_path = here("Data", "columns_per_year.txt"),
  column_name_path = here("Data", "column_names_per_year.txt")
)
```

## Filter & rename columns

We filter out education level 9 because there are only 2 occurrences. Smoking has more than 2 levels
but we only need two. Race 1 (Mexican American) and 2 (other Hispanic) are merged. Income could be
added but we'd have to exclude 12, 13, 77, 99 and missing. 12 is over 20k, 13 is under 20k, 77 is
refused and 99 is don't know. This would drop the size of the data set by about 20%.

Code levels: **Smoking** 1. Smoker (L1) 2. Non-smoker L2) 7. Refused 9. Don't know **Gender** 1.
Male (L1) 2. Female (L2) **Race** 1. Mexican/other Hispanic (L1) 3. White (L2) 4. Black (L3) 5.
Other (L4) **Education** 1. Less than 9th grade (L1) 2. 9-11th grade (includes 12th grade with no
diploma) (L2) 3. High school (L3) 4. Some college or AA degree (L4) 5. College graduate or above 7.
Refused 9. Don't know

```{r format_df}
df <- df_raw %>%
  select(
    SEQN,
    age,
    all_of(confounder_vars),
    kdm,
    phenoage,
    hd_log,
    telolength,
    all_of(params$hearing_var),
    albumin_gL,
    alp,
    bun,
    creat,
    crp,
    glucose,
    hba1c,
    lymph,
    mcv,
    rdw,
    sbp,
    totchol,
    wbc,
    uap
  ) %>%
  filter(
    education != 9,
    smoking %in% c(1, 2)
  ) %>%
  mutate(
    gender = factor(gender),
    smoking = factor(smoking),
    race = case_when(race == 2 ~ 1, TRUE ~ race),
    race = factor(race),
    education = factor(education),
    phenoage_advance = phenoage - age
  )
```

## Optionally only include telolength rows

```{r telolength_only}
if (isTRUE(params$telolength)) {
  df <- df %>%
    filter(
      !is.na(telolength)
    )
}
```

## Prepare age group column

```{r set_age_groups}
df <- df %>% mutate(
  age_group = case_when(
    age >= 20 & age < 30 ~ "20-30",
    age >= 30 & age < 40 ~ "30-40",
    age >= 40 & age < 50 ~ "40-50",
    age >= 50 & age < 60 ~ "50-60",
    age >= 60 & age < 70 ~ "60-70",
    isFALSE(params$telolength) & age >= 70 ~ "70+",
  ),
  confounders = interaction(gender, smoking)
)
```
