---
editor_options:
  markdown:
    wrap: 100
---

# Data preprocessing

## Data selection & standardization

We make a list of files, store the year and filename from each file, then append the data to the
row. Next, we take all the files from one year and join the rows if they match or columns if they
don't match according to SEQN (ID) and create a `year` column and `cohort` column that takes the
first 4 characters from `year`. `imap_dfr()` applies this to each year (map), returns the index (i)
and row-binds into a dataframe (dfr).

### Select columns

We intentionally do not select creat_umol columns ([;) if we have to restandardize them later (1999
and 2005)

```{r}
select_columns <- function(data_map, column_path) {
  col_spec <- read.csv(column_path)

  col_spec <- col_spec %>%
    group_by(year_group) %>%
    summarize(
      select_cols = list(unique(cols)),
      .groups = "drop"
    )

  # Merge data from the same year into one dataframe
  data_map <- data_map %>%
    group_by(cohort) %>%
    summarise(
      data = list(reduce(data, full_join, by = "SEQN")),
      .groups = "drop"
    )

  # Select columns to keep per year
  data_map %>%
    mutate(
      data = map2(data, cohort, \(df, yr) {
        cols_to_keep <- unlist(
          col_spec %>%
            filter(year_group == yr) %>%
            pull(select_cols)
        )

        df %>%
          select(all_of(cols_to_keep)) %>%
          mutate(year = factor(yr))
      })
    )
}
```

### Rename columns

```{r}
rename_columns <- function(data_map, column_name_path) {
  col_name_spec <- read.csv(column_name_path)

  data_map %>%
    mutate(
      data = map2(data, cohort, \(df, yr) {
        rename_spec <- col_name_spec %>%
          filter(year == yr)

        df %>%
          rename_with(
            ~ rename_spec$new_name[which(rename_spec$old_name == .x)],
            .cols = rename_spec$old_name
          )
      })
    )
}
```

### Merge dataframe and create ID vars

```{r}
merge_data <- function(data_map) {
  data_map %>%
    pull(data) %>%
    bind_rows() %>%
    mutate(sampleID = paste(year, SEQN, sep = "_")) %>%
    relocate(SEQN, sampleID, year)
}
```

### Standardize values

Creatinine in 1999 and 2005 and CRP in 2015 and 2017 are standardized. SBP is averaged over multiple
measurements.

```{r}
standardize_values <- function(data_map) {
  # browser()
  data_map %>%
    mutate(
      creat = case_when(
        year == 1999 ~ LBXSCR * 1.013 + 0.147,
        year == 2005 ~ LBXSCR * 0.978 - 0.016,
        TRUE ~ creat
      ),
      creat_umol = case_when(
        year %in% c(1999, 2005) ~ creat * 88.4,
        TRUE ~ creat_umol
      ),
      crp = case_when(
        year %in% c(2015, 2017) ~ LBXHSCRP / 10,
        TRUE ~ crp
      ),
      lncrp = log(crp + 1),
      lncreat = log(creat + 1),
      lncreat_umol = log(creat_umol + 1),
      telolength_log = log(telolength)
    ) %>%
    rowwise() %>%
    mutate(
      sbp = case_when(
        year == 2017 ~ mean(c(BPXOSY1, BPXOSY2, BPXOSY3), na.rm = TRUE),
        TRUE ~ mean(c(BPXSY1, BPXSY2, BPXSY3, BPXSY4), na.rm = TRUE)
      )
    ) %>%
    ungroup() %>%
    select(
      -LBXSCR,
      -LBXHSCRP,
      -BPXSY1,
      -BPXSY2,
      -BPXSY3,
      -BPXSY4,
      -BPXOSY1,
      -BPXOSY2,
      -BPXOSY3
    )
}
```

### Data selection & renaming pipeline

```{r include=FALSE}
df_unfiltered <- df_raw %>%
  select_columns(
    column_path = here("Specs", "column_spec.csv")
  ) %>%
  rename_columns(
    column_name_path = here("Specs", "column_name_spec.csv")
  ) %>%
  merge_data() %>%
  standardize_values()
```

#### Function to construct table

Takes a dataframe and title and makes a scrollable table.

```{r fun_formula_table}
construct_table <- function(formula_list, title) {
  formula_list %>%
    kable(
      "html",
      caption = title
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE,
      position = "left",
      font_size = 12,
      fixed_thead = TRUE
    ) %>%
    scroll_box(width = "100%", height = "400px")
}
```

#### Check completeness

```{r}
df_unfiltered %>%
  group_by(year) %>%
  summarise(across(everything(), ~ sum(!is.na(.)), .names = "n_{.col}")) %>%
  select(year, where(~ any(. == 0))) %>%
  construct_table(title = "Check missingness")
```

Income is missing in 2017 but should be there according to the website as INDFMIN2. CRP is missing
in 2011, which seems to be correct. Telomere length was only assessed in early cohorts. Dialysis and
job exposure were only assessed in later cohorts.

## Calculate hearing metrics

```{r}
calculate_hearing <- function(data) {
  data %>%
    rowwise() %>%
    mutate(
      test_retest_diff_left = abs(audio_left_1000 - audio_left_1000_retest),
      test_retest_diff_right = abs(audio_right_1000 - audio_right_1000_retest),
      audio_right_1000_mean = mean(c(audio_right_1000, audio_right_1000_retest), na.rm = TRUE),
      audio_left_1000_mean = mean(c(audio_left_1000, audio_left_1000_retest), na.rm = TRUE),
      FI_low_r = mean(c(audio_right_500, audio_right_1000, audio_right_2000), na.rm = TRUE),
      FI_low_l = mean(c(audio_left_500, audio_left_1000, audio_left_2000), na.rm = TRUE),
      FI_low = mean(c(FI_low_r, FI_low_l), na.rm = TRUE),
      FI_high_r = mean(c(audio_right_4000, audio_right_6000, audio_right_8000), na.rm = TRUE),
      FI_high_l = mean(c(audio_left_4000, audio_left_6000, audio_left_8000), na.rm = TRUE),
      FI_high = mean(c(FI_high_r, FI_high_l), na.rm = TRUE),
      pta_r = mean(c(
        audio_right_500,
        audio_right_1000,
        audio_right_2000,
        audio_right_4000
      ), na.rm = TRUE),
      pta_l = mean(c(
        audio_left_500,
        audio_left_1000,
        audio_left_2000,
        audio_left_4000
      ), na.rm = TRUE),
      pta = mean(c(pta_r, pta_l)),
      pta_diff = abs(pta_r - pta_l)
    ) %>%
    ungroup()
}

df_unfiltered <- df_unfiltered %>% calculate_hearing()
```

## Data filtering

We keep any columns we may need at this stage or have already used, such as the individual
audiometry columns.

```{r}
filter_data <- function(data) {
  verbose_filter <- function(data, ..., label) {
    start_n <- nrow(data)
    data_filtered <- data %>% filter(...)
    message(start_n - nrow(data_filtered), "/", start_n, " excluded due to ", label)

    data_filtered
  }

  audiometry_cols <- data %>%
    select(contains("audio")) %>%
    colnames()

  biomarkers <- c(
    "albumin",
    "alp",
    "lncrp",
    "totchol",
    "lncreat",
    "hba1c",
    "sbp",
    "bun",
    "uap",
    "lymph",
    "mcv",
    "wbc"
  )

  data <- data %>%
    verbose_filter(
      case_when(
        year %in% c(1999, 2001, 2003, 2005) ~ age < 85,
        TRUE ~ age < 80
      ),
      label = "too old age."
    ) %>%
    verbose_filter(
      age >= 20,
      label = "too young age."
    ) %>%
    verbose_filter(
      !if_any(all_of(audiometry_cols), ~ . == 888),
      !if_any(all_of(audiometry_cols), ~ . == 666),
      !if_any(all_of(audiometry_cols), ~ is.na(.)),
      label = "missing audiometry data."
    ) %>%
    verbose_filter(
      test_retest_diff_left < 10,
      test_retest_diff_right < 10,
      label = "too large test-retest difference."
    ) %>%
    verbose_filter(
      pta_diff < 15,
      label = "too large imbalance in left/right hearing loss."
    ) %>%
    verbose_filter(
      pregnant != 1 | is.na(pregnant),
      label = "pregnancy."
    ) %>%
    verbose_filter(
      !if_any(all_of(biomarkers), ~ is.na(.)),
      label = "missing biomarkers."
    ) %>%
    group_by(gender) %>%
    verbose_filter(
      !if_any(
        all_of(biomarkers),
        ~ {
          m <- mean(.x, na.rm = TRUE)
          s <- sd(.x, na.rm = TRUE)
          abs(.x - m) > 5 * s
        }
      ),
      label = "biomarker outlier (greater or smaller than 5 SDs)."
    ) %>%
    ungroup()
}

df <- df_unfiltered %>% filter_data()
```

### Check data issues

```{r}
df_unfiltered %>%
  mutate(
    age_group = case_when(
      age < 6 ~ "00–06",
      age >= 6 & age <= 11 ~ "06–12",
      age >= 12 & age <= 19 ~ "12–19",
      age > 19 & age < 70 ~ "19–75",
      age >= 70 ~ "75+",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!year %in% c(2007, 2011)) %>%
  group_by(year, age_group) %>%
  summarise(
    n_na = sum(is.na(audio_left_1000)),
    n = n(),
    .groups = "drop"
  ) %>%
  mutate(prop_present = paste0(100 * round(1 - n_na / n, 2), "%")) %>%
  select(-n_na, -n) %>%
  pivot_wider(names_from = age_group, values_from = prop_present) %>%
  construct_table(title = "Show proportion audiometry present")

df %>%
  group_by(year) %>%
  summarize(
    age = paste0(round(mean(age), 1), "(", round(sd(age), 1), ")"),
    FI_high = paste0(round(mean(FI_high), 1), "(", round(sd(FI_high), 1), ")"),
    FI_low = paste0(round(mean(FI_low), 1), "(", round(sd(FI_low), 1), ")"),
    albumin = paste0(round(mean(albumin), 1), "(", round(sd(albumin), 1), ")"),
    alp = paste0(round(mean(alp), 1), "(", round(sd(alp), 1), ")"),
    lncrp = paste0(round(mean(lncrp), 1), "(", round(sd(lncrp), 1), ")"),
    totchol = paste0(round(mean(totchol), 1), "(", round(sd(totchol), 1), ")"),
    creat = paste0(round(mean(creat), 1), "(", round(sd(creat), 1), ")"),
    hba1c = paste0(round(mean(hba1c), 1), "(", round(sd(hba1c), 1), ")"),
    sbp = paste0(round(mean(sbp), 1), "(", round(sd(sbp), 1), ")"),
    bun = paste0(round(mean(bun), 1), "(", round(sd(bun), 1), ")"),
    uap = paste0(round(mean(uap), 1), "(", round(sd(uap), 1), ")"),
    lymph = paste0(round(mean(lymph), 1), "(", round(sd(lymph), 1), ")"),
    mcv = paste0(round(mean(mcv), 1), "(", round(sd(mcv), 1), ")"),
    wbc = paste0(round(mean(wbc), 1), "(", round(sd(wbc), 1), ")"),
    .groups = "drop"
  ) %>%
  construct_table(title = "Check cohort age and biomarker balance")

p <- df %>%
  group_by(age, gender) %>%
  summarize(n = n(), .groups = "drop") %>%
  group_by(age) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup() %>%
  filter(gender == "1") %>%
  mutate(gender = factor(gender)) %>%
  ggplot(aes(x = age, y = prop, colour = gender)) +
  geom_smooth(se = FALSE, span = 0.5, show.legend = FALSE) +
  geom_line(show.legend = FALSE) +
  coord_cartesian(ylim = c(0, 1)) +
  scale_color_manual(values = "#2986cc") +
  labs(title = "Proportion males in filtered dataset", x = "Age", y = "Proportion")

sameplot(p)

p <- df %>%
  group_by(age, gender) %>%
  summarize(n = n(), .groups = "drop") %>%
  mutate(gender = factor(gender)) %>%
  ggplot(aes(x = age, y = n, group = gender, color = gender)) +
  geom_line(show.legend = FALSE) +
  scale_y_continuous(breaks = seq(0, 100, 10)) +
  coord_cartesian(ylim = c(0, 110)) +
  scale_color_manual(values = c("1" = "#2986cc", "2" = "#c90076")) +
  labs(title = "Count of males and females in filtered dataset", x = "Age", y = "Count")

sameplot(p)
```

Audiometry was not performed on main cohort for 2005, 2009 and 2017, so only people \>= 70 years
were included from these years. Audiometry was only performed in 12-19 year olds in 2007 so none are
include from this cohort. 1/2 of the sample was included for audiometry in 1999, 2001, and 2003. ALP
was much higher in 1999 despite age being normal. CRP was not collected in 2011, so none are
included from this cohort. Totchol is lower in newer cohorts (even when included population is
older).

## Calculate biological age metrics

```{r warning=FALSE}
bioage_calc <- function(data, training_data_path, HD_training_data_path) {
  v1_biomarkers <- c(
    "fev", "sbp", "totchol", "hba1c", "albumin", "creat", "lncrp", "alp", "bun"
  )
  v2_biomarkers <- c(
    "albumin", "alp", "lncrp", "totchol", "lncreat", "hba1c", "sbp", "bun", "uap", "lymph", "mcv", "wbc"
  )
  v2_biomarkers_phenoage <- c(
    "albumin_gL", "alp", "lncrp", "totchol", "lncreat_umol", "hba1c", "sbp", "bun", "uap", "lymph", "mcv", "wbc"
  )

  load(training_data_path)
  load(HD_training_data_path)

  # Calculate KDM V2
  train_fem_kdm_v2 <- kdm_calc(
    data = NHANES3 %>%
      filter(gender == 2 & age >= 30 & age <= 75 & pregnant == 0),
    biomarkers = v2_biomarkers,
    fit = NULL,
    s_ba2 = NULL,
    n_allowed_missing_biomarkers = 0
  )
  train_male_kdm_v2 <- kdm_calc(
    data = NHANES3 %>%
      filter(gender == 1 & age >= 30 & age <= 75),
    biomarkers = v2_biomarkers,
    fit = NULL,
    s_ba2 = NULL,
    n_allowed_missing_biomarkers = 0
  )
  test_fem_kdm_v2 <- kdm_calc(
    data = data %>%
      filter(gender == 2),
    biomarkers = v2_biomarkers,
    fit = train_fem_kdm_v2$fit,
    s_ba2 = train_fem_kdm_v2$fit$s_ba2,
    n_allowed_missing_biomarkers = 0
  )
  test_male_kdm_v2 <- kdm_calc(
    data = data %>%
      filter(gender == 1),
    biomarkers = v2_biomarkers,
    fit = train_male_kdm_v2$fit,
    s_ba2 = train_male_kdm_v2$fit$s_ba2,
    n_allowed_missing_biomarkers = 0
  )

  test_kdm_v2 <- rbind(test_fem_kdm_v2$data, test_male_kdm_v2$data)

  data <- left_join(data, test_kdm_v2[, c("sampleID", "kdm", "kdm_advance")], by = "sampleID")

  # Calculate PhenoAge V2
  train_phenoage_v2 <- phenoage_calc(
    data = NHANES3 %>%
      filter(age >= 20 & age <= 84),
    biomarkers = v2_biomarkers_phenoage,
    fit = NULL
  )

  test_phenoage_v2 <- phenoage_calc(
    data = data,
    biomarkers = v2_biomarkers_phenoage,
    fit = train_phenoage_v2$fit
  )

  data <- left_join(data, test_phenoage_v2$data[, c("sampleID", "phenoage", "phenoage_advance")], by = "sampleID")

  # Calculate HD
  fem_hd <- hd_calc(
    data = data %>%
      filter(gender == 2),
    reference = NHANES3_HDTrain %>%
      filter(gender == 2),
    biomarkers = v2_biomarkers
  )
  male_hd <- hd_calc(
    data = data %>%
      filter(gender == 1),
    reference = NHANES3_HDTrain %>%
      filter(gender == 1),
    biomarkers = v2_biomarkers
  )

  hd <- rbind(fem_hd$data, male_hd$data)

  data <- left_join(data, hd[, c("sampleID", "hd", "hd_log")], by = "sampleID")

  data
}

df <- df %>% bioage_calc(
  training_data_path = here("Data", "Training data", "NHANES3.rda"),
  HD_training_data_path = here("Data", "Training data", "NHANES3_HDTrain.rda")
)
```

### Set data types

```{r}
df <- df %>%
  mutate(
    SEQN = as.character(SEQN),
    sampleID = as.character(sampleID),
    gender = factor(gender),
    pregnant = factor(pregnant),
    race = factor(race),
    education = factor(education),
    income = factor(income),
    angina = factor(angina),
    mi = factor(mi),
    stroke = factor(stroke),
    smoking = factor(smoking),
    diabetes = factor(diabetes),
    htn = factor(htn),
    kid_failure = factor(kid_failure),
    firearm = factor(firearm),
    offwork_exposure = factor(offwork_exposure),
    dialysis = factor(dialysis),
    job_exposure_loud = factor(job_exposure_loud),
    job_exposure_very_loud = factor(job_exposure_very_loud)
  )
```

## Cleanup environment

From this point onward, we only keep what will actually be used for analysis. We remove df_raw and
df_unfiltered, but don't run `gc()`, as the objects being removed from the environment already
labels the RAM as free to be allocated, thus it's not needed to run garbage collection manually.

```{r}
rm(df_raw)
rm(df_unfiltered)

df <- df %>%
  select(
    SEQN,
    age,
    all_of(confounder_vars),
    kdm,
    phenoage,
    hd_log,
    telolength,
    all_of(params$hearing_var),
    albumin_gL,
    alp,
    bun,
    creat,
    crp,
    glucose,
    hba1c,
    lymph,
    mcv,
    rdw,
    sbp,
    totchol,
    wbc,
    uap
  )
```

## Recode variables

We filter out education level 9 because there are only 2 occurrences. Smoking has more than 2 levels
but we only need two. Race 1 (Mexican American) and 2 (other Hispanic) are merged. Income could be
added but we'd have to exclude 12, 13, 77, 99 and missing. 12 is over 20k, 13 is under 20k, 77 is
refused and 99 is don't know. This would drop the size of the data set by about 20%.

Code levels: **Smoking** 1. Smoker (L1) 2. Non-smoker L2) 7. Refused 9. Don't know **Gender** 1.
Male (L1) 2. Female (L2) **Race** 1. Mexican/other Hispanic (L1) 3. White (L2) 4. Black (L3) 5.
Other (L4) **Education** 1. Less than 9th grade (L1) 2. 9-11th grade (includes 12th grade with no
diploma) (L2) 3. High school (L3) 4. Some college or AA degree (L4) 5. College graduate or above 7.
Refused 9. Don't know

```{r format_df}
df <- df %>%
  filter(
    education != 9,
    smoking %in% c(1, 2)
  ) %>%
  mutate(
    race = case_when(race == "2" ~ "1", TRUE ~ race),
  )
```

## Optionally only include telolength rows

```{r telolength_only}
if (isTRUE(params$telolength)) {
  df <- df %>%
    filter(
      !is.na(telolength)
    )
}
```

## Prepare age group column

```{r set_age_groups}
df <- df %>% mutate(
  age_group = case_when(
    age >= 20 & age < 30 ~ "20-30",
    age >= 30 & age < 40 ~ "30-40",
    age >= 40 & age < 50 ~ "40-50",
    age >= 50 & age < 60 ~ "50-60",
    age >= 60 & age < 70 ~ "60-70",
    isFALSE(params$telolength) & age >= 70 ~ "70+",
  ),
  confounders = interaction(gender, smoking)
)
```
