# Data Preprocessing

## Unpack confounders list

We take the param value for confounders and build one lists with all included interactions with age and one with just the variables

```{r confounder_list}
confounder_interactions <- trimws(
  strsplit(params$confounders, "_")[[1]]
)

confounder_vars <- confounder_interactions[!grepl(",", confounder_interactions)]
```

## Remove Unnecessary Columns & Rename

We filter out education level 9 because there are only 2 occurrences. Smoking
has more than 2 levels but we only need two. Race 1 (Mexican American) and 2
(other Hispanic) are merged. Income could be added but we'd have to exclude 12,
13, 77, 99 and missing. 12 is over 20k, 13 is under 20k, 77 is refused and 99 is
don't know. This would drop the size of the data set by about 20%.

Code levels: **Smoking** 1. Smoker (L1) 2. Non-smoker L2) 7. Refused 9. Don't
know **Gender** 1. Male (L1) 2. Female (L2) **Race** 1. Mexican/other Hispanic
(L1) 3. White (L2) 4. Black (L3) 5. Other (L4) **Education** 1. Less than 9th
grade (L1) 2. 9-11th grade (includes 12th grade with no diploma) (L2) 3. High
school (L3) 4. Some college or AA degree (L4) 5. College graduate or above 7.
Refused 9. Don't know

```{r format_df}
df <- df_raw %>%
  select(
    SEQN,
    age,
    all_of(confounder_vars),
    kdm,
    phenoage,
    hd_log,
    telolength,
    all_of(params$hearing_var),
    albumin_gL,
    alp,
    bun,
    creat,
    crp,
    glucose,
    hba1c,
    lymph,
    mcv,
    rdw,
    sbp,
    totchol,
    wbc,
    uap
  ) %>%
  filter(
    education != 9,
    smoking %in% c(1, 2)
  ) %>%
  mutate(
    gender = factor(gender),
    smoking = factor(smoking),
    race = case_when(race == 2 ~ 1, TRUE ~ race),
    race = factor(race),
    education = factor(education),
    phenoage_advance = phenoage - age
  )
```

## Optionally only include telolength rows

```{r telolength_only}
if (isTRUE(params$telolength)) {
  df <- df %>%
    filter(
      !is.na(telolength)
    )
}
```

## Prepare age group column

```{r set_age_groups}
df <- df %>% mutate(
  age_group = case_when(
    age >= 20 & age < 30 ~ "20-30",
    age >= 30 & age < 40 ~ "30-40",
    age >= 40 & age < 50 ~ "40-50",
    age >= 50 & age < 60 ~ "50-60",
    age >= 60 & age < 70 ~ "60-70",
    isFALSE(params$telolength) & age >= 70 ~ "70+",
  ),
  confounders = interaction(gender, smoking)
)
```
