# Modelling

## K fold splitting

### Fold-maker function

This function takes the indices of rows each age group, randomizes them, then
assigns numbers of 1-10. We could integrate this into the k-fold cross
validation function, but by doing it separately it's easier to inspect any
possible issues. We shuffle indices within each age group so that the numbers 1-10 will equally be spaced throughout the age group, creating even age group distributions in each fold.

```{r fun_make_folds}
make_repeated_folds <- function(data, grouping_variable, k, repeats) {
  # Helper to create random k-fold indices for one repeat
  kfold_split_stratified <- function(k, grouping_variable = NULL) {
    # Convert factor of age_group to number
    grouping_variable <- as.integer(
      as.factor(grouping_variable)
    )

    # Check number of age groups and observations
    group_len <- length(
      unique(grouping_variable)
    )
    observations <- length(grouping_variable)

    # Empty vector to add fold numbers to
    indices <- numeric()

    # Run loop for each age group
    for (l in 1:group_len) {
      # Find which rows have the age group
      indices_grouping_variable <- which(grouping_variable == l)

      # Shuffle indices within each age group and append
      indices <- c(indices, sample(indices_grouping_variable))
    }

    # Create an output with the length of the number of observations
    bins <- rep(observations)

    # Using the random order, 1-10 is repeated until we've reached the last observation
    bins[indices] <- rep(1:k, ceiling(observations / k))[1:observations]

    # Return shuffled k-folds, which is pasted into a new column in 'data'
    bins
  }

  # Per repeat, make shuffled k-folds of indices in a new column
  for (i in 1:repeats) {
    data[[paste0("fold_r", i)]] <- kfold_split_stratified(
      k = k,
      grouping_variable = data[[grouping_variable]]
    )
  }

  data
}
```

### Make folds per age group

```{r make_folds}
set.seed(2)
df <- make_repeated_folds(
  data = df,
  grouping_variable = "age_group",
  k = params$global_folds,
  repeats = params$global_repeats
)
```

## Formula building

### Formula builder function

The `k` determines the number of base functions (linear or non-linear).The parameter
with the lowest degrees of freedom (deg_free) determines the maximum basis dimensions
(k in mgcv). Albumin has 26 distinct levels, so the maximal k value/number of
basic functions, equivalent to the deg_free, is 25 if each level is in each fold. We
must compute the deg_free for each fold and pick the lowest to avoid errors. The GAM
is a composite of all these underlying functions, with different weights for
each function. Max degrees of freedom are only computed once per variable and cached for retrieval.

-   `k`: number of basis functions.
-   `max_k`: sets k to the theoretical limit for each parameter (deg_free), note that
    for non-integers, like the bioage measures, this will be equivalent to the
    length of the dataframe minus one.
-   `auto_limit_k`: if k exceeds the deg_free for a parameter (e.g., if k \> 25 for
    albumin), it will automatically be limited to the deg_free.
-   `equal_wiggliness`: gives an individual id to each sum-to-zero interaction term when FALSE which ensures they have the same smoothing parameter. Setting to TRUE marginally improves fit but uses many degrees of freedom.
-   `stepdown`: additionally creates formulas with age as predictor that each miss one confounder for a nested-CV comparison (i.e., compare, for example, the impact of adding 'gender' to a model that already includes other covariates).

```{r fun_formula_gen}
formula_generator <- function(
  data,
  formulas,
  confounders = NULL,
  k = 25,
  max_k = FALSE,
  auto_limit_k = FALSE,
  equal_wiggliness = TRUE,
  stepdown = FALSE
) {
  # Initialize the repeat columns with folds and unique fold numbers in folds
  fold_columns <- grep("^fold_r", names(data), value = TRUE)
  folds <- seq_along(fold_columns) # get each unique fold number in folds

  # Initialize a list for compiled formulas
  formula_list <- list()

  deg_free_store <- tibble(
    predictor = character(),
    deg_free = numeric()
  )

  for (formula in formulas) {
    # Unlist, split by comma, and get rid of white spaces for the current predictors string
    predictors <- trimws(
      unlist(
        strsplit(formula, ",")
      )
    )

    # Initialize a list for terms to compile formula from
    sm_terms <- list()

    # make a term for each predictor and load degrees of freedom if already computed for predictor
    for (predictor in predictors) {
      if (predictor %in% deg_free_store$predictor) {
        # Load degrees of freedom from cache if already computed else create list
        deg_free <- deg_free_store$deg_free[deg_free_store$predictor == predictor]
      } else {
        deg_free_list <- list()

        # Store the deg_free for each fold in each repeat in a list
        for (rep in fold_columns) {
          for (fold in folds) {
            deg_free_list <- append(
              deg_free_list,
              length(
                unique(
                  data[data[[rep]] != fold, ][[predictor]]
                )
              )
            )
          }
        }

        # Pick the smallest value in the list and subtract 1 to obtain degrees of freedom
        deg_free <- min(
          unlist(deg_free_list)
        ) - 1
        deg_free_store <- bind_rows(
          deg_free_store,
          tibble(predictor = predictor, deg_free = deg_free)
        )
      }

      # Overwrite k with max allowed k if desired
      k_formula <- if (max_k) deg_free else if (auto_limit_k && k > deg_free) deg_free else k

      if (predictor == "age") {
        # Construct age term
        sm_terms <- append(
          sm_terms,
          paste0(
            "s(age, k = ",
            k_formula,
            ")"
          )
        )

        if (!is.null(confounders)) {
          # construct Age * confounder terms
          for (i in seq_along(confounders)) {
            confounder <- confounders[i]
            if (isTRUE(equal_wiggliness)) {
              # Combine an age term with smooth terms with a unique ID for each
              sm_terms <- append(
                sm_terms,
                paste0(
                  "s(age, ",
                  confounder,
                  ", bs = 'sz', k = ",
                  k,
                  ", id = ",
                  i,
                  ")"
                )
              )
            } else {
              # Combine an age term with smooth terms
              sm_terms <- append(
                sm_terms,
                paste0(
                  "s(age, ",
                  confounder,
                  ", bs = 'sz', k = ",
                  k,
                  ")"
                )
              )
            }
          }
        }
      } else {
        # Construct predictor terms
        sm_terms <- append(
          sm_terms,
          paste0(
            "s(",
            predictor,
            ", k = ",
            k_formula,
            ")"
          )
        )
      }
    }

    # Put predictor terms in formula and separate by +
    formula_list <- append(
      formula_list,
      paste(
        params$hearing_var,
        "~",
        paste(unlist(sm_terms), collapse = " + ")
      )
    )

    # Create confounder stepdown vars
    if (isTRUE(stepdown) && formula == "age") {
      # Age will already have been created by the previous loop
      deg_free <- deg_free_store$deg_free[deg_free_store$predictor == "age"]

      k_formula <- if (max_k) deg_free else if (auto_limit_k && k > deg_free) deg_free else k

      # Remove one confounder at a time, construct the formula, and add it to the formula list
      for (i in seq_along(confounders)) {
        sm_terms <- list()

        # Add age term
        sm_terms <- append(
          sm_terms,
          paste0(
            "s(age, k = ",
            k_formula,
            ")"
          )
        )
        constrained_confounders <- confounders[-i]

        # Remove one confounder from formula at a time and create terms (with or without ID)
        for (j in seq_along(constrained_confounders)) {
          constrained_confounder <- constrained_confounders[j]
          if (isTRUE(equal_wiggliness)) {
            sm_terms <- append(
              sm_terms,
              paste0(
                "s(age, ",
                constrained_confounder,
                ", bs = 'sz', k = ",
                k_formula,
                ", id = ",
                j,
                ")"
              )
            )
          } else {
            sm_terms <- append(
              sm_terms,
              paste0(
                "s(age, ",
                constrained_confounder,
                ", bs = 'sz', k = ",
                k_formula,
                ")"
              )
            )
          }
        }

        formula_list <- append(
          formula_list,
          paste(
            params$hearing_var,
            "~",
            paste(
              unlist(sm_terms),
              collapse = " + "
            )
          )
        )
      }
    }
  }

  formula_list
}
```

### Extract variables from formula function

Helper used to turn a formula back into a list of variables.

```{r fun_extr_pred}
extract_predictors <- function(formula, collapse = "") {
  s_terms <- unlist(
    str_extract_all(
      formula,
      "s\\([^\\)]*\\)"
    )
  )
  keep <- !str_detect(
    s_terms,
    paste0(
      "\\b(",
      paste(confounder_vars, collapse = "|"),
      ")\\b"
    )
  )
  str_match(
    s_terms[keep],
    "^s\\(\\s*([^,\\)\\s]+)"
  )[, 2]
}
```

### Get formula table function

Takes a dataframe and title and makes a scrollable table.

```{r fun_formula_table}
get_formula_table <- function(formula_list, title) {
  tibble(
    Formulas = unlist(formula_list)
  ) %>%
    kable(
      "html",
      caption = title
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE,
      position = "left",
      font_size = 12,
      fixed_thead = TRUE
    ) %>%
    scroll_box(width = "100%", height = "400px")
}
```

### Generate formulas bioage

```{r build_bioage_forms}
formula_list_bioage <- formula_generator(
  data = df,
  k = params$global_k,
  auto_limit_k = TRUE,
  equal_wiggliness = TRUE,
  formulas = c(
    "age",
    "age, phenoage",
    "age, kdm",
    "age, hd_log",
    if (isTRUE(params$telolength)) {
      "age, telolength"
    } else {
      NULL
    }
  ),
  confounders = confounder_interactions
)

get_formula_table(
  formula_list = formula_list_bioage,
  title = "Bioage Formulas"
)
```

### Generate formulas biomarkers

```{r build_biomark_forms}
formula_list_biomarkers <- formula_generator(
  data = df,
  k = params$global_k,
  auto_limit_k = TRUE,
  equal_wiggliness = TRUE,
  formulas = c(
    "age",
    "age, phenoage",
    "age, albumin_gL",
    "age, alp",
    "age, creat",
    "age, glucose",
    "age, crp",
    "age, lymph",
    "age, mcv",
    "age, rdw",
    "age, wbc",
    "age, sbp",
    "age, bun",
    "age, totchol",
    "age, hba1c",
    "age, uap",
    "age, crp, hba1c, albumin_gL",
    "age, crp, hba1c, albumin_gL, phenoage"
  ),
  confounders = confounder_interactions
)

get_formula_table(
  formula_list = formula_list_biomarkers,
  title = "Biomarker Formulas"
)
```

### Generate formulas confounders

```{r build_conf_forms}
if (isFALSE(params$telolength)) {
  formula_list_confounders <- formula_generator(
    data = df,
    k = params$global_k,
    auto_limit_k = TRUE,
    equal_wiggliness = TRUE,
    formulas = c("age", "age, phenoage"),
    confounders = confounder_interactions,
    stepdown = TRUE
  )

  get_formula_table(
    formula_list = formula_list_confounders,
    title = "Confounder Formulas"
  )
}
```

## Repeated k-fold modeling

### Repeated k-fold validation function per age group

Fits a BAM() to the formulas generated splitting data into folds. Predictions are made using the models and MSEs are calculated with these predictions per age group. A list is returned with a tibble with results (MSE per fold per formula), predictions (each individual prediction made), and optionally the actual models (uses very large amounts of memory when discrete = TRUE). When discrete = TRUE, discretizes fitting using fREML (much faster but slightly worse fit). Confounder_analysis does nothing to the fitting but returns names based on nested covariate models instead of the regular predictors (which exclude covariates in the label).

```{r fun_CV}
cross_validation <- function(
  data,
  formulas,
  k,
  repeats,
  discrete = FALSE,
  keep_models = FALSE,
  confounder_analysis = FALSE
) {
  # Initialize age groups to stratify by and repeat column names that contain fold specification
  age_groups <- c(
    "20-30",
    "30-40",
    "40-50",
    "50-60",
    "60-70",
    if (isFALSE(params$telolength)) "70+"
  )
  repeat_columns <- paste0("fold_r", seq_len(repeats))
  folds <- k

  # Initialize list of models, CV results (MSEs), predictions, and indices of lists
  models <- list()
  results_list <- list()
  predicted_list <- list()
  index_results <- 1L
  index_predicted <- 1L

  # Initialize vector of confounders and store predictor column names for prediction dataset
  confounders <- confounder_vars
  predictor_columns <- unique(
    unlist(
      lapply(formulas, extract_predictors)
    )
  )

  # Set predictor combination names
  if (isTRUE(confounder_analysis)) {
    predictor_labels <- c(
      "full",
      paste0("no_", confounder_interactions),
      "full_with_phenoage"
    )
  } else {
    predictor_labels <- unique(
      sapply(formulas, function(f) {
        paste(
          extract_predictors(f),
          collapse = ", "
        )
      })
    )
  }

  # Only select the columns that we need
  data <- data %>% select(
    all_of(
      c(
        "SEQN",
        params$hearing_var,
        "age_group",
        predictor_columns,
        confounder_vars,
        repeat_columns
      )
    )
  )

  for (formula in formulas) {
    start_time <- Sys.time()
    for (repeat_column in repeat_columns) {
      for (fold in seq_len(folds)) {
        # Subset training and testing data
        train <- data %>% filter(
          .data[[repeat_column]] != fold
        )
        test <- data %>% filter(
          .data[[repeat_column]] == fold
        )

        # Fit GAM
        mod <- bam(
          as.formula(formula),
          data = train,
          family = gaussian(),
          method = if (isTRUE(discrete)) "fREML" else "REML",
          discrete = if (isTRUE(discrete)) TRUE else FALSE
        )

        # Extract predictor or get from predefined variable for confounder analysis
        if (isTRUE(confounder_analysis)) {
          predictors <- predictor_labels[which(formulas == formula)]
        } else {
          predictors <- paste(extract_predictors(formula), collapse = ", ")
        }

        # Extract repetition number only
        repeat_value <- as.numeric(
          sub(".*fold_r(\\d+).*", "\\1", repeat_column)
        )

        # Wrap model into a list to keep the models separated, attach, and name
        if (keep_models) {
          model_id <- paste0(predictors, "_r", repeat_value, "_f", fold)
          models[[model_id]] <- mod
        }

        # calculate performance per age group
        for (age_group_val in age_groups) {
          # Subset testing and training data for age group
          train_subset <- train %>%
            filter(
              age_group == age_group_val
            )
          test_subset <- test %>%
            filter(
              age_group == age_group_val
            )

          # Use GAM to predict responses in current age group
          pred <- predict(mod, newdata = test_subset, type = "response")

          # Calculate squared difference between predicted vs real in this age group
          results_list[[index_results]] <- tibble(
            predictors = predictors,
            age_group = age_group_val,
            r = repeat_value,
            k = fold,
            n_train = nrow(train_subset),
            n_test = nrow(test_subset),
            mse = mean((test_subset[[params$hearing_var]] - pred)^2)
          )
          index_results <- index_results + 1L

          # add predictions used to calculate RMSE later to a dataframe
          predicted_list[[index_predicted]] <- test_subset %>%
            as_tibble() %>%
            mutate(
              predictors = predictors,
              age_group = age_group_val,
              r = repeat_value,
              k = fold,
              predicted = pred
            )
          index_predicted <- index_predicted + 1L
        }
      }
    }
    print(paste(formula, Sys.time() - start_time))
  }

  # Bind results/predictions together and set factors when needed
  results <- bind_rows(results_list)
  results <- results %>%
    mutate(
      predictors = factor(
        predictors,
        levels = predictor_labels
      ),
      age_group = factor(
        age_group,
        levels = age_groups
      )
    )

  predictions <- bind_rows(predicted_list)
  predictions <- predictions %>%
    mutate(
      predictors = factor(
        predictors,
        levels = predictor_labels
      ),
      age_group = factor(age_group, levels = age_groups)
    )

  list(
    results = results,
    cv_models = models,
    predictions = predictions
  )
}
```

### Generate model path function

```{r fun_gen_m_paths}
generate_model_path <- function(output_type) {
  folder_name <- paste0(
    params$hearing_var,
    if (isTRUE(params$telolength)) "_telolength" else NULL
  )

  file_name <- paste0(
    output_type,
    "_k=",
    params$global_k,
    "_discrete=",
    params$global_discrete,
    ".RDS"
  )

  model_path <- here("Models", folder_name, file_name)

  dir.create(dirname(model_path), recursive = TRUE, showWarnings = FALSE)

  model_path
}
```

### Model bioage

Cross-validations rerun if an .RDS file with the corresponding name does not exist or if load_model_results is FALSE. Results are stored if store_model_results is TRUE.

#### Bioage K-fold CV

```{r bioage_cv}
system.time({
  output_bioage <- if (
    isTRUE(params$load_model_results) &&
      file.exists(generate_model_path("bioage"))
  ) {
    readRDS(generate_model_path("bioage"))
  } else {
    cross_validation(
      df,
      formula_list_bioage,
      k = params$global_folds,
      repeats = params$global_repeats,
      discrete = params$global_discrete,
      keep_models = params$keep_models
    )
  }
})
```

#### Store bioage modeling results

```{r store_bioage_res}
if (isTRUE(params$store_model_results)) {
  saveRDS(
    object = output_bioage,
    file = generate_model_path("bioage"),
    compress = "xz"
  )
}
```

### Model biomarkers

#### Biomarkers K-fold CV

```{r biomarkers_cv}
if (isFALSE(params$telolength)) {
  system.time({
    output_biomarkers <- if (
      isTRUE(params$load_model_results) &&
        file.exists(generate_model_path("biomarkers"))
    ) {
      readRDS(generate_model_path("biomarkers"))
    } else {
      cross_validation(
        df,
        formula_list_biomarkers,
        k = params$global_folds,
        repeats = params$global_repeats,
        discrete = params$global_discrete,
        keep_models = params$keep_models
      )
    }
  })
}
```

#### Store bioage modeling results

```{r store_biomarkers_cv}
if (isFALSE(params$telolength) && isTRUE(params$store_model_results)) {
  saveRDS(
    object = output_biomarkers,
    file = generate_model_path("biomarkers"),
    compress = "xz"
  )
}
```

### Model confounders

#### Confounders K-fold CV

```{r confounders_cv}
if (isFALSE(params$telolength)) {
  system.time({
    output_confounders <- if (
      isTRUE(params$load_model_results) &&
        file.exists(generate_model_path("confounders"))
    ) {
      readRDS(generate_model_path("confounders"))
    } else {
      cross_validation(
        df,
        formula_list_confounders,
        k = params$global_folds,
        repeats = params$global_repeats,
        discrete = params$global_discrete,
        confounder_analysis = TRUE
      )
    }
  })
}
```

#### Build confounder named step-down models

Build models for covariate analysis to use for in-sample statistics or predictions.

```{r fit_conf_models}
if (isFALSE(params$telolength) && is.null(output_confounders$models)) {
  for (formula in formula_list_confounders) {
    output_confounders$models <- append(
      output_confounders$models,
      list(
        bam(
          as.formula(formula),
          data = df,
          family = gaussian(),
          method = "REML",
          discrete = FALSE,
          confounder_analysis = TRUE
        )
      )
    )
  }

  names(output_confounders$models) <- c(
    "full",
    paste0("no_", confounder_interactions),
    "full_with_phenoage"
  )
}
```

#### Store confounders modeling results

```{r store_conf_res}
if (isFALSE(params$telolength) && isTRUE(params$store_model_results)) {
  saveRDS(
    object = output_confounders,
    file = generate_model_path("confounders"),
    compress = "xz"
  )
}
```

## Model age rate

### Build GAM and Linear Age Models in a List

When solely wanting to predict an outcome variable, the wiggliness penalty of a
GAM prevents overfitting, so k can be as large as we'd like. However, with a
larger k, we have more parameters, so if we take the derivative of the GAM and
we have more parameters, each with its own uncertainty, the total uncertainty
will grow, even if the extra base functions don't contribute. Therefore, a
larger k past the number that captures all global trends decreases statistical
power. This is a problem when assessing the rate of aging. Therefore, we need to
have a penalty for practically useless base functions so that they shrink to 0. Default shrinkage solution still permits large complexity despite the RMSE improving little over a simpler model.

We can use k.check with a small k to see if its enough. With
a k of 10 we already get a k index of 0.9998, so k = 10 should be sufficient.

```{r aging_models}
output_age_rate <- list(models = list(), predictions = list())

model_age_rate <- function(terms) {
  gam(
    reformulate(terms, response = params$hearing_var),
    data = df,
    method = "REML",
    family = gaussian()
  )
}

output_age_rate$models <- append(
  output_age_rate$models,
  list(
    linear = model_age_rate("age"),
    quadratic = model_age_rate(c("age", "I(age^2)")),
    smooth = model_age_rate(c("age", "I(age^2)", "s(age, k = 10)"))
  )
)

set.seed(1)
k.check(output_age_rate$models$smooth)
```

## Optional tests

### (Optional) Inspect folds

We check if each fold number is assigned equal amounts of times in two repeats
for each age group. Thereafter, we save the row number and plot the fold number
of points (with some jitter to spread things out) to check if bands appear
homogenous. We know it should be homogenous in the y-axis within a band because
the jitter is random. If there are similarities between bands, then the shuffle
isn't randomized for each repeat. If the bands don't appear of similar intensity
over the x-axis, then the number assignment is not random.

```{r inspect_folds, fig.height = 10, fig.width = 10}
if (isTRUE(params$run_optional_chunks)) {
  df %>%
    group_by(age_group) %>%
    mutate(total_fold1 = n()) %>%
    group_by(age_group, fold_r1, total_fold1) %>%
    summarize(
      n_fold1 = n(),
      .groups = "drop"
    ) %>%
    rename("fold" = fold_r1) %>%
    left_join(
      df %>%
        group_by(age_group) %>%
        mutate(total_fold2 = n()) %>%
        group_by(age_group, fold_r2, total_fold2) %>%
        summarize(
          n_fold2 = n(),
          .groups = "drop"
        ) %>%
        select(age_group, fold_r2, total_fold2, n_fold2, ) %>%
        rename("fold" = fold_r2),
      by = c("age_group", "fold")
    ) %>%
    kable(
      "html",
      caption = "Inspect Fold Number Frequencies"
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE,
      position = "left",
      font_size = 12,
      fixed_thead = TRUE
    ) %>%
    scroll_box(width = "100%", height = "400px")

  p <- df %>%
    mutate(row_index = row_number()) %>%
    pivot_longer(
      cols = starts_with("fold_r"),
      names_to = "r",
      values_to = "fold"
    ) %>%
    ggplot(aes(x = row_index, y = fold)) +
    geom_point(
      size = 1,
      alpha = 0.1,
      color = "grey80",
      position = position_jitter(width = 0, height = 0.4)
    ) +
    facet_wrap(~r) +
    theme_few()

  plot(p)
}
```

Fold numbers are consistent between groups and correctly split upon multiple
repeats. each fold number has the same (within 1) frequency, so there are no
imbalances between folds and repeats. Plot shows no non-random changes in
intensity over the X dimension and random dissimilarities between bands in the Y
dimension (indicating repeats have different samples).

### (Optional) Check confounder two-way interactions vs offsets

We want to get a sense which confounders improve the model and which interact
with age.

```{r assess_confounders}
if (isTRUE(params$run_optional_chunks)) {
  assess_confounders <- function(df, confounders, k) {
    f0 <- paste0(params$hearing_var, " ~ s(age, k = ", k, ")")
    m0 <- bam(as.formula(f0), data = df, method = "REML")

    results <- data.frame()

    for (confounder in confounders) {
      f1 <- paste0(
        params$hearing_var,
        " ~ ",
        confounder,
        " + s(age, k = ",
        k,
        ")"
      )
      f2 <- paste0(
        params$hearing_var,
        " ~ s(age) + s(age, ",
        confounder,
        ", bs = 'sz', k = ",
        k,
        ")"
      )

      m1 <- bam(
        as.formula(f1),
        data = df,
        method = "REML"
      )
      m2 <- bam(
        as.formula(f2),
        data = df,
        method = "REML"
      )

      aov <- as.data.frame(
        anova(m0, m1, m2)
      )
      aic <- as.data.frame(
        AIC(m0, m1, m2)
      )
      bic <- as.data.frame(
        BIC(m0, m1, m2)
      )

      results <- rbind(
        results,
        cbind(
          tibble(
            confounder = confounder,
            formula = c(f0, f1, f2)
          ),
          aic,
          bic,
          aov
        )
      )
    }

    return(results)
  }

  assess_confounders(df, confounders = confounder_vars, k = params$global_k)
}
```

**We see that for each confounder, the interaction model fits better than the
intercept-only model, so we should use interaction terms instead of intercepts for covariates.**

### (Optional) Investigate Additional Information

First we see if the confounders are correlated (e.g., smoking is more prevalent
in males than females so gender is partly coded in smoking), to see if all
confounders actually contribute.

```{r compare_added_info}
if (isTRUE(params$run_optional_chunks)) {
  compare_confounders <- function(df, k) {
    m0 <- bam(
      reformulate("s(age, k = k)", response = params$hearing_var),
      data = df,
      method = "REML"
    )
    m1 <- bam(
      reformulate(
        c(
          "s(age, k = k)",
          "s(age, gender, bs = 'sz', k = k)"
        ),
        response = params$hearing_var
      ),
      data = df,
      method = "REML"
    )
    m2 <- bam(
      reformulate(
        c(
          "s(age, k = k)",
          "s(age, gender, bs = 'sz', k = k)",
          "s(age, race, bs = 'sz', k = k)"
        ),
        response = params$hearing_var
      ),
      data = df,
      method = "REML"
    )
    m3 <- bam(
      reformulate(
        c(
          "s(age, k = k)",
          "s(age, gender, bs = 'sz', k = k)",
          "s(age, race, bs = 'sz', k = k)",
          "s(age, smoking, bs = 'sz', k = k)"
        ),
        response = params$hearing_var
      ),
      data = df,
      method = "REML"
    )
    m4 <- bam(
      reformulate(c(
        "s(age, k = k)",
        "s(age, gender, bs = 'sz', k = k)",
        "s(age, race, bs = 'sz', k = k)",
        "s(age, smoking, bs = 'sz', k = k)",
        "s(age, education, bs = 'sz', k = k)"
      ), response = params$hearing_var),
      data = df,
      method = "REML"
    )

    print(
      cbind(
        as.data.frame(
          AIC(m0, m1, m2, m3, m4)
        ),
        as.data.frame(
          BIC(m0, m1, m2, m3, m4)
        ),
        as.data.frame(
          anova(m0, m1, m2, m3, m4)
        )
      )
    )
    print(
      summary(m4)
    )
    plot(m4)
  }

  compare_confounders(df, k = params$global_k)
}
```

**Each factor significantly improves the model and seems to interact with age in
some way**

### (Optional) Check shared wiggliness

We could insist that different smooths have the same amount of smoothness, i.e.,
that there's no difference in the wiggliness of the GAM between men and women.
The plots from the previous chunk (where different wiggliness was permitted)
suggest that there may be equal amount of wiggliness, so that each smooth has
the same basis functions but with different coefficients. There's no specific
reason to assume that wiggliness is the same in many cases, so if aiming to
maximize model flexibility/predictive power, we should leave this unconstrained,
but we do eat up a lot of degrees of freedom and may be 'wasting' the data
estimating these global trends. It may also be better to spend the processing
time on a higher k rather than estimating smooth parameters.

See the link below for more info on constructing interaction terms:
<https://stats.stackexchange.com/questions/637423/conceptual-interpretation-of-bs-fs-and-by-term-in-gam>.


```{r compare_wiggliness}
if (isTRUE(params$run_optional_chunks)) {
  compare_wiggliness <- function(df, k) {
    m1 <- bam(
      as.formula(
        paste0(
          params$hearing_var,
          ' ~
      s(age, k = k) +
      s(age, gender, bs = "sz", k = k, id = 1) +
      s(age, race, bs = "sz", k = k, id = 2) +
      s(age, smoking, bs = "sz", k = k, id = 3) +
      s(age, education, bs = "sz", k = k, id = 4)'
        )
      ),
      data = df,
      method = "REML"
    )

    m2 <- bam(
      as.formula(
        paste0(
          params$hearing_var,
          ' ~
      s(age, k = k) +

      s(age, gender, bs = "sz", k = k) +
      s(age, race, bs = "sz", k = k) +
      s(age, smoking, bs = "sz", k = k) +
      s(age, education, bs = "sz", k = k)'
        )
      ),
      data = df,
      method = "REML"
    )

    print(
      cbind(
        as.data.frame(
          AIC(m1, m2)
        ),
        as.data.frame(
          BIC(m1, m2)
        ),
        as.data.frame(
          anova(m1, m2)
        )
      )
    )
  }

  compare_wiggliness(df, k = params$global_k)
}
```

**The free wiggliness model takes about 15 times longer to compute and barely has a
drop in residual deviance. The improvement in fit is not significant, so enforce shared smoothing parameters.**

### (Optional) Check factor three-way interactions

```{r conf_interactions}
if (isTRUE(params$run_optional_chunks)) {
  compare_confounder_interactions <- function(df, k, equal_wiggliness, return_model = FALSE) {
    # Pull confounders, cross them, keep copy with var1 being first (alphabetical), paste by comma
    terms <- c(
      confounder_vars,
      crossing(var1 = confounder_vars, var2 = confounder_vars) %>%
        filter(var1 < var2) %>%
        mutate(pair = paste(var1, var2, sep = ", ")) %>%
        pull(pair)
    )

    # Combine an age term with smooth terms with or without a unique ID for each
    if (isTRUE(equal_wiggliness)) {
      smooth_terms <- c(
        paste0("s(age, k = ", k, ")"),
        paste0(
          "s(",
          "age, ",
          terms,
          ", bs = 'sz', k = ",
          k,
          ", id = ",
          seq_along(smooth_terms),
          ")"
        )
      )
    } else {
      smooth_terms <- c(
        paste0("s(age, k = ", k, ")"),
        paste0(
          "s(",
          "age, ",
          terms,
          ", bs = 'sz', k = ",
          k,
          ")"
        )
      )
    }

    m0 <- bam(
      as.formula(
        paste0(
          params$hearing_var,
          " ~ ",
          paste(smooth_terms, collapse = " + ")
        )
      ),
      data = df,
      method = "REML"
    )

    print(summary(m0))
    plot(m0)

    if (return_model) {
      m0
    }
  }

  system.time({
    compare_confounder_interactions(
      df,
      k = 5,
      equal_wiggliness = TRUE
    )
  })
}
```

We see that `age*gender*race` is a significant three-way interaction but it
takes 3 times longer to compute. Bam() speeds things up enough over gam() that this processing cost is acceptable.

### (Optional) Inspect k-values

To assess k-value, we fit a model on the whole dataset using the formula lists,
build a list of models, make a helper function to run the k.check function from
mgcv, then turn it in a dataframe and return the output. The k-value determines
the number of base functions that the GAM is constructed from. A GAM adds these
different base functions together, some of which capture linear trends, some of
which are just splines or polynomials with a bump in one region. The GAM gives
different weights to each model to get as good a fit as possible, but a GAM also
measures wiggliness and gives a penalty to wiggliness. This is what makes it
generalized. I.e., to perform well in general, even to unseen data. Therefore, a
GAM optimizes to get as good a fit as possible with as little wiggliness as
possible. Because of this wiggliness penalty, adding more base functions to the
GAM doesn't make it overfit. Therefore, k should be as large as possible within
computation power or degrees of freedom. If k isn't large enough, there will be
patterns left in the residuals. Therefore, by measuring patterns in the
residuals, we can assess if k is large enough. We want to capture all global
patterns without fitting to any noise. This is what we assess when we look at
k-index.

Models are fit on whole dataset.

```{r inspect_k}
if (isTRUE(params$run_optional_chunks)) {
  inspect_k_values <- function(data, formulas, discrete = FALSE, subsample = 5000, reps = 400) {
    set.seed(2)

    models <- list()

    for (formula in formulas) {
      mod <- bam(
        as.formula(formula),
        data = data,
        family = gaussian(),
        method = if (isTRUE(discrete)) "fREML" else "REML",
        discrete = if (isTRUE(discrete)) TRUE else FALSE
      )

      predictors <- paste(
        extract_predictors(formula),
        collapse = ", "
      )

      models[[predictors]] <- mod
    }

    inspect_k <- function(model) {
      k.check(
        model,
        subsample = subsample,
        n.rep = reps
      )
    }

    k_output <- lapply(models, inspect_k)
    k_output <- imap_dfr(
      k_output,
      ~ {
        as.data.frame(.x) %>%
          rownames_to_column(var = "Predictors") %>%
          mutate(model = .y) %>%
          filter(!is.na(`k-index`))
      }
    )

    k_output
  }

  inspect_k_values(df,
    formula_list_bioage,
    discrete = params$global_discrete,
    subsample = 5000,
    reps = 400
  ) %>%
    kable(
      "html",
      caption = "Inspect K Bioage"
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE,
      position = "left",
      font_size = 12,
      fixed_thead = TRUE
    ) %>%
    scroll_box(width = "100%", height = "400px")

  inspect_k_values(df,
    formula_list_biomarkers,
    discrete = params$global_discrete,
    subsample = 5000,
    reps = 400
  ) %>%
    kable(
      "html",
      caption = "Inspect K Biomarkers"
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE,
      position = "left",
      font_size = 12,
      fixed_thead = TRUE
    ) %>%
    scroll_box(width = "100%", height = "400px")
}
```

### (Optional) Test k-fold validation

Test k-fold validation on a single formula with specified k.

```{r test_k_fold_val}
if (isTRUE(params$run_optional_chunks)) {
  test_cv <- function(run = FALSE, k = 5) {
    if (isTRUE(run)) {
      formula_list_test <- formula_generator(
        data = df,
        k = k,
        auto_limit_k = TRUE,
        equal_wiggliness = TRUE,
        formulas = c("age, phenoage")
      )

      system.time({
        output_test <- cross_validation(
          df,
          formula_list_test,
          k = params$global_folds,
          repeats = params$global_repeats,
          discrete = params$global_discrete
        )
      })
    }
  }

  test_cv(run = TRUE, k = 5)
}
```
