# Modelling

## K fold splitting

### Fold-maker function

This function takes the indices of rows each age group, randomizes them, then
assigns numbers of 1-10. We could integrate this into the k-fold cross
validation function, but by doing it separately it's easier to inspect any
possible issues.

```{r fun_make_folds}
make_repeated_folds <- function(data, grouping_variable, k, repeats) {
  # helper function to make random k-fold indices for one repeat
  kfold_split_stratified <- function(k, grouping_variable = NULL) {
    grouping_variable <- as.integer(as.factor(grouping_variable)) # convert factor of age_group to number
    group_len <- length(unique(grouping_variable)) # check number of age groups
    observations <- length(grouping_variable) # check how many observations there are to get a number

    indices <- numeric() # empty list to add fold numbers to

    for (l in 1:group_len) {
      # run loop for each age group
      indices_grouping_variable <- which(grouping_variable == l) # find which rows have the age group
      indices <- c(indices, sample(indices_grouping_variable)) # randomly shuffle indegrouping_variablees within each age group, so that the numbers 1-10 will equally be spaced throughout the age group, creating even age group distributions in each fold
    }

    bins <- rep(observations) # create an output with the length of the number of observations
    bins[indices] <- rep(1:k, ceiling(observations / k))[1:observations] # using the random order, 1-10 is repeated until we've reached the last observation
    return(bins) # return shuffled k-folds, which is pasted into a new column in 'data'
  }

  for (i in 1:repeats) {
    # per repeat, make shuffled k-folds of indices in a new column
    data[[paste0("fold_r", i)]] <- kfold_split_stratified(k = k, grouping_variable = data[[grouping_variable]]) # writes k-folds to a new column for each repeat
  }

  return(data) # return the dataframe with a k-fold column per repeat
}
```

### Make folds

If k-fold validation throws errors that testing data contains factor levels that
weren't in the training data, it may be needed to make an interaction variable
of age_group with the problematic factor (or condense levels) to ensure the
smaller factor levels are represented in each fold.

```{r make_folds}
set.seed(2)
df <- make_repeated_folds(
  data = df,
  grouping_variable = "age_group",
  k = params$global_folds,
  repeats = params$global_repeats
)
```

## Formula building

### Formula builder function

The `k` determines the number of functions (linear or non-linear).The parameter
with the lowest degrees of freedom (DF) determines the maximum basis dimensions
(k in mgcv). Albumin has 26 distinct levels, so the maximal k value/number of
basic functions, equivalent to the DF, is 25 if each level is in each fold. We
must compute the DF for each fold and pick the lowest to avoid errors. The GAM
is a composite of all these underlying functions, with different weights for
each function.

-   `k`: number of basis functions.
-   `max_k`: sets k to the theoretical limit for each parameter (DF), note that
    for non-integers, like the bioage measures, this will be equivalent to the
    length of the dataframe minus one.
-   `auto_limit_k`: if k exceeds the DF for a parameter (e.g., if k \> 25 for
    albumin), it will automatically be limited to the DF.

```{r fun_formula_gen}
formula_generator <- function(data,
                              formulas,
                              confounders = NULL,
                              k = 25,
                              max_k = FALSE,
                              auto_limit_k = FALSE,
                              equal_wiggliness = TRUE,
                              stepdown = FALSE) {
  fold_columns <- grep("^fold_r", names(data), value = TRUE) # initialize the repeat columns with folds
  folds <- 1:length(fold_columns) # get each unique fold number in folds
  formula_list <- list() # initialize a list to store the compiled formulas in

  DF_store <- tibble(
    predictor = character(),
    DF = numeric()
  )

  for (formula in formulas) {
    predictors <- trimws(unlist(strsplit(formula, ","))) # unlist, split by comma, and get rid of white spaces for the current predictors string
    sm_terms <- list()

    # make a term for each predictor
    for (predictor in predictors) {
      if (predictor %in% DF_store$predictor) {
        DF <- DF_store$DF[DF_store$predictor == predictor]
      } else {
        DF_list <- list() # initialize a list to store the DF for each fold in

        for (rep in fold_columns) {
          for (fold in folds) {
            DF_list <- append(DF_list, length(unique(data[data[[rep]] != fold, ][[predictor]]))) # store the DF for each fold in each repeat in a list
          }
        }
        DF <- min(unlist(DF_list)) - 1 # pick the smallest value in the list and subtract 1 to obtain degrees of freedom
        DF_store <- bind_rows(DF_store, tibble(predictor = predictor, DF = DF)) # Store in list
      }

      if (max_k) {
        k_formula <- DF
      } else {
        k_formula <- k
      }

      if (auto_limit_k) {
        if (k_formula > DF) {
          k_formula <- DF
        }
      }
      if ((predictor == "age") && (equal_wiggliness)) {
        sm_terms <- append(sm_terms, paste0("s(age, k = ", k_formula, ")"))


        if (!is.null(confounders)) {
          for (i in seq_along(confounders)) {
            confounder <- confounders[i]
            sm_terms <- append(
              sm_terms,
              paste0(
                "s(age, ",
                confounder,
                ", bs = 'sz', k = ",
                k_formula,
                ", id = ",
                i,
                ") "
              )
            )
          }
        }
      } else if ((predictor == "age") && !(equal_wiggliness)) {
        sm_terms <- append(sm_terms, paste0("s(age, k = ", k_formula, ")"))
        if (!is.null(confounders)) {
          for (confounder in confounders) {
            sm_terms <- append(
              sm_terms,
              paste0(
                "s(age, ",
                confounder,
                ", bs = 'sz', k = ",
                k_formula,
                ") "
              )
            )
          }
        }
      } else {
        sm_terms <- append(
          sm_terms,
          paste0("s(", predictor, ", k = ", k_formula, ")")
        )
      }
    }
    formula_list <- append(formula_list, paste(params$hearing_var, "~", paste(unlist(sm_terms), collapse = " + "))) # put predictor terms in formula and separate by +
    if (isTRUE(stepdown) && formula == "age") {
      if ("age" %in% DF_store$predictor) {
        DF <- DF_store$DF[DF_store$predictor == "age"]
      } else {
        DF_list <- list()
        for (rep in fold_columns) {
          for (fold in folds) {
            DF_list <- append(DF_list, length(unique(data[data[[rep]] != fold, ][["age"]])))
          }
        }

        DF <- min(unlist(DF_list)) - 1
        DF_store <- bind_rows(DF_store, tibble(predictor = "age", DF = DF)) # Store in list
      }
      if (max_k) {
        k_formula <- DF
      } else {
        k_formula <- k
      }

      if (auto_limit_k) {
        if (k_formula > DF) {
          k_formula <- DF
        }
      }
      for (i in seq_along(confounders)) {
        sm_terms <- list()
        sm_terms <- append(sm_terms, paste0("s(age, k = ", k_formula, ")"))
        constrained_confounders <- confounders[-i]

        for (j in seq_along(constrained_confounders)) {
          constrained_confounder <- constrained_confounders[j]
          sm_terms <- append(
            sm_terms,
            paste0(
              "s(age, ",
              constrained_confounder,
              ", bs = 'sz', k = ",
              k_formula,
              ", id = ",
              j,
              ") "
            )
          )
        }

        formula_list <- append(formula_list, paste(params$hearing_var, "~", paste(unlist(sm_terms), collapse = " + ")))
      }
    }
  }
  return(formula_list)
}
```

### Extract variables from formula function

```{r fun_extr_pred}
extract_predictors <- function(formula, collapse = "") {
  confounders <- confounder_vars

  s_terms <- unlist(str_extract_all(formula, "s\\([^\\)]*\\)"))
  keep <- !str_detect(s_terms, paste0("\\b(", paste(confounders, collapse = "|"), ")\\b"))
  str_match(s_terms[keep], "^s\\(\\s*([^,\\)\\s]+)")[, 2]
}
```

### Generate formulas bioage

```{r build_bioage_forms}
formula_list_bioage <- formula_generator(
  data = df,
  k = params$global_k,
  auto_limit_k = TRUE,
  equal_wiggliness = TRUE,
  formulas = c(
    "age",
    "age, phenoage",
    "age, kdm",
    "age, hd_log",
    if (isTRUE(params$telolength)) {
      "age, telolength"
    } else {
      NULL
    }
  ),
  confounders = confounder_interactions
)

as.data.frame(cbind(Formulas = unlist(formula_list_bioage))) %>%
  kable("html", caption = "Bioage Formulas") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position = "left",
    font_size = 12,
    fixed_thead = TRUE
  ) %>%
  scroll_box(width = "100%", height = "400px")
```

### Generate formulas biomarkers

```{r build_biomark_forms}
formula_list_biomarkers <- formula_generator(
  data = df,
  k = params$global_k,
  auto_limit_k = TRUE,
  equal_wiggliness = TRUE,
  formulas = c(
    "age",
    "age, phenoage",
    "age, albumin_gL",
    "age, alp",
    "age, creat",
    "age, glucose",
    "age, crp",
    "age, lymph",
    "age, mcv",
    "age, rdw",
    "age, wbc",
    "age, sbp",
    "age, bun",
    "age, totchol",
    "age, hba1c",
    "age, uap",
    "age, crp, hba1c, albumin_gL",
    "age, crp, hba1c, albumin_gL, phenoage"
  ),
  confounders = confounder_interactions
)

as.data.frame(cbind(Formulas = unlist(formula_list_biomarkers))) %>%
  kable("html", caption = "Biomarker Formulas") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position = "left",
    font_size = 12,
    fixed_thead = TRUE
  ) %>%
  scroll_box(width = "100%", height = "400px")
```

### Generate formulas confounders

```{r build_conf_forms}
if (isFALSE(params$telolength)) {
  formula_list_confounders <- formula_generator(
    data = df,
    k = params$global_k,
    auto_limit_k = TRUE,
    equal_wiggliness = TRUE,
    formulas = c("age", "age, phenoage"),
    confounders = confounder_interactions,
    stepdown = TRUE
  )
}
```


## Repeated k-fold modeling

### Repeated k-fold validation function per age group

```{r fun_CV}
cross_validation <- function(data,
                             formulas,
                             k,
                             repeats,
                             discrete = FALSE,
                             keep_models = FALSE,
                             confounder_analysis = FALSE) {
  age_groups <- c("20-30", "30-40", "40-50", "50-60", "60-70", if (isFALSE(params$telolength)) "70+" else NULL) # initialize age groups to stratify by explicitly
  fold_columns <- paste0("fold_r", 1:repeats) # initialize names of repeat columns to iterate over

  models <- list() # intialize list of models
  results_list <- list() # Initialize list for results
  predicted_list <- list() # Initialize list for predictions
  index_results <- 1L
  index_predicted <- 1L

  confounders <- confounder_vars # initialize vector of confounders
  names_of_predictors <- unique(unlist(lapply(formulas, extract_predictors))) # save names of all predictor columns to select for the pred vs obs dataset

  if (isTRUE(confounder_analysis)) {
    predictor_combinations <- c(
      "full",
      paste0("no_", confounder_interactions),
      "full_with_phenoage"
    )
  } else {
    predictor_combinations <- unique(sapply(formulas, function(f) {
      paste(extract_predictors(f), collapse = ", ")
    }))
  }

  data <- data %>% select(all_of(
    c(
      params$hearing_var,
      "age_group",
      names_of_predictors,
      confounder_vars,
      fold_columns
    )
  )) # Only take the columns that we need

  for (formula in formulas) {
    for (r in fold_columns) {
      for (f in 1:k) {
        train <- data %>% filter(.data[[r]] != f) # set training data
        test <- data %>% filter(.data[[r]] == f) # set testing data

        if (discrete) {
          mod <- bam(
            as.formula(formula),
            data = train,
            family = gaussian(),
            method = "fREML",
            discrete = TRUE
          ) # fit GAM
        } else {
          mod <- bam(
            as.formula(formula),
            data = train,
            family = gaussian(),
            method = "REML",
            discrete = FALSE
          ) # fit GAM
        }
        if (isTRUE(confounder_analysis)) {
          predictors <- predictor_combinations[which(formulas == formula)]
        } else if (isFALSE(confounder_analysis)) {
          predictors <- paste(extract_predictors(formula), collapse = ", ") # extract predictors from formula
        }

        rep <- as.numeric(sub(".*fold_r(\\d+).*", "\\1", r)) # extract repetition number only

        if (keep_models) {
          model_id <- paste0(predictors, "_r", rep, "_f", f)
          models[[model_id]] <- mod # wrap model into a list to keep the models separated, attach it to the list of models, and name the model according to predictors, repeat, and fold
        }

        # calculate performance per age group
        for (age_group_val in age_groups) {
          test_subset <- test %>%
            filter(age_group == age_group_val) # subset testing data for age group
          train_subset <- train %>%
            filter(age_group == age_group_val) # subset training data for age group

          pred <- predict(mod, newdata = test_subset, type = "response") # use GAM to predict responses in current age group
          results_list[[index_results]] <- tibble(
            predictors = predictors,
            age_group = age_group_val,
            r = rep,
            k = f,
            n_train = nrow(train_subset),
            n_test = nrow(test_subset),
            mse = mean((test_subset[[params$hearing_var]] - pred)^2) # calculate squared difference between predicted vs real in this age group
          )
          index_results <- index_results + 1L

          predicted_list[[index_predicted]] <- test_subset %>%
            as_tibble() %>%
            mutate(
              predictors = predictors,
              age_group = age_group_val,
              r = rep,
              k = f,
              predicted = pred
            ) # add predictions used to calculate MSE to a dataframe
          index_predicted <- index_predicted + 1L
        }
      }
    }
  }

  results <- bind_rows(results_list)
  results <- results %>%
    mutate(
      predictors = factor(predictors, levels = predictor_combinations),
      age_group = factor(age_group, levels = age_groups)
    ) # Factorize

  pred_vs_obs <- bind_rows(predicted_list)
  pred_vs_obs <- pred_vs_obs %>%
    mutate(
      predictors = factor(predictors, levels = predictor_combinations),
      age_group = factor(age_group, levels = age_groups)
    ) # Factorize

  return(list(
    results = results,
    cv_models = models,
    predictions = pred_vs_obs
  )) # return models, dataframe with results, and a fitted vs predicted dataframe
}
```

### Generate model path function

```{r fun_gen_m_paths}
generate_model_path <- function(output_type) {
  model_path <- here(
    "Models",
    paste0(params$hearing_var, if (isTRUE(params$telolength)) {
      "_telolength"
    } else {
      NULL
    }),
    paste0(
      output_type,
      "_k=",
      params$global_k,
      "_discrete=",
      params$global_discrete,
      ".RDS"
    )
  )

  dir.create(dirname(model_path), recursive = TRUE, showWarnings = FALSE)

  return(model_path)
}
```

### Model bioage

#### Bioage K-fold CV
With k = 40 discrete = FALSE takes 6759s With k = 50 discrete = FALSE takes
13506s

```{r bioage_cv}
system.time({
  output_bioage <- if (isTRUE(params$load_model_results) && file.exists(generate_model_path("bioage"))) {
    readRDS(generate_model_path("bioage"))
  } else {
    cross_validation(
      df,
      formula_list_bioage,
      k = params$global_folds,
      repeats = params$global_repeats,
      discrete = params$global_discrete,
      keep_models = params$keep_models
    )
  }
})
```

#### Store bioage modeling results

```{r store_bioage_res}
if (isTRUE(params$store_model_results)) {
  saveRDS(output_bioage, generate_model_path("bioage"), compress = "xz")
}
```


### Model biomarkers

#### Biomarkers K-fold CV

With k = 40 discrete = FALSE takes 52564s WIth k = 50 discrete = FALSE takes
110443s

```{r biomarkers_cv}
if (isFALSE(params$telolength)) {
  system.time({
    output_biomarkers <- if (isTRUE(params$load_model_results) &&
      file.exists(generate_model_path("biomarkers"))) {
      readRDS(generate_model_path("biomarkers"))
    } else {
      cross_validation(
        df,
        formula_list_biomarkers,
        k = params$global_folds,
        repeats = params$global_repeats,
        discrete = params$global_discrete,
        keep_models = params$keep_models
      )
    }
  })
}
```

#### Store bioage modeling results

```{r store_biomarkers_cv}
if (isFALSE(params$telolength)) {
  if (isTRUE(params$store_model_results)) {
    saveRDS(output_biomarkers,
      generate_model_path("biomarkers"),
      compress = "xz"
    )
  }
}
```

#### Fit full biomarker model

```{r fit_full_biomark_m}
if (isFALSE(params$telolength)) {
  fit_full_model <- function(fit) {
    if (fit) {
      full_biomarker_model <-
        bam(
          as.formula(
            formula_generator(
              data = df,
              k = params$global_k,
              auto_limit_k = TRUE,
              equal_wiggliness = TRUE,
              formulas = c(
                "age, crp, hba1c, albumin_gL, alp, wbc, glucose, uap, creat, totchol, rdw, lymph, mcv, bun, sbp, phenoage"
              ),
              confounders = confounder_interactions
            )
          ),
          data = df,
          method = "REML",
          discrete = FALSE
        )

      print(summary(full_biomarker_model))

      return(full_biomarker_model)
    }
  }

  fit_full_model(fit = FALSE)
}
```

### Model confounders

#### Confounders K-fold CV

```{r confounders_cv}
if (isFALSE(params$telolength)) {
  system.time({
    output_confounders <- if (isTRUE(params$load_model_results) && file.exists(generate_model_path("confounders"))) {
      readRDS(generate_model_path("confounders"))
    } else {
      cross_validation(
        df,
        formula_list_confounders,
        k = params$global_folds,
        repeats = params$global_repeats,
        discrete = params$global_discrete,
        confounder_analysis = TRUE
      )
    }
  })
}
```

#### Build confounder named step-down models

```{r fit_conf_models}
if (isFALSE(params$telolength) &&
  is.null(output_confounders$models)) {
  for (formula in formula_list_confounders) {
    output_confounders$models <- append(output_confounders$models, list(
      bam(
        as.formula(formula),
        data = df,
        family = gaussian(),
        method = "REML",
        discrete = FALSE,
        confounder_analysis = TRUE
      )
    ))
  }

  names(output_confounders$models) <-
    c(
      "full",
      paste0("no_", confounder_interactions),
      "full_with_phenoage"
    )
}
```

#### Store confounders modeling results

```{r store_conf_res}
if (isFALSE(params$telolength)) {
  if (isTRUE(params$store_model_results)) {
    saveRDS(output_confounders, generate_model_path("confounders"), compress = "xz")
  }
}
```

## Model age rate

### Build GAM and Linear Age Models in a List

When solely wanting to predict an outcome variable, the wiggliness penalty of a
GAM prevents overfitting, so k can be as large as we'd like. However, with a
larger k, we have more parameters, so if we take the derivative of the GAM and
we have more parameters, each with its own uncertainty, the total uncertainty
will grow, even if the extra base functions don't contribute much. Therefore, a
larger k past the number that captures all global trends decreases statistical
power. This is a problem when assessing the rate of aging. Therefore, we need to
have a penalty for practically useless base functions so that they shrink to 0.
As an alternative, we can use k.check with a small k to see if its enough. With
a k of 10 we already get an edf of almost 8, so k = 10 should be sufficient.

```{r aging_models}
output_age_rate <- list(models = list(), predictions = list())

output_age_rate$models <- append(output_age_rate$models, list(
  linear =
    gam(
      reformulate(c("age"), response = params$hearing_var),
      data = df,
      method = "REML",
      family = gaussian()
    )
))

output_age_rate$models <- append(output_age_rate$models, list(
  quadratic =
    gam(
      reformulate(c("age", "I(age^2)"), response = params$hearing_var),
      data = df,
      method = "REML",
      family = gaussian()
    )
))

output_age_rate$models <- append(output_age_rate$models, list(
  smooth =
    gam(
      reformulate(c("age", "I(age^2)", "s(age, k = 10)"), response = params$hearing_var),
      data = df,
      method = "REML",
      family = gaussian()
    )
))

set.seed <- 1
k.check(output_age_rate$models$smooth)
```

## Optional tests

### (Optional) Inspect folds

We check if each fold number is assigned equal amounts of times in two repeats
for each age group. Thereafter, we save the row number and plot the fold number
of points (with some jitter to spread things out) to check if bands appear
homogenous. We know it should be homogenous in the y-axis within a band because
the jitter is random. If there are similarities between bands, then the shuffle
isn't randomized for each repeat. If the bands don't appear of similar intensity
over the x-axis, then the number assignment is not random.

```{r inspect_folds, fig.height = 10, fig.width = 10}
if (isTRUE(params$run_optional_chunks)) {
  df %>%
    group_by(age_group) %>%
    mutate(total_fold1 = n()) %>%
    group_by(age_group, fold_r1, total_fold1) %>%
    summarize(
      n_fold1 = n(),
      .groups = "drop"
    ) %>%
    rename("fold" = fold_r1) %>%
    left_join(
      df %>%
        group_by(age_group) %>%
        mutate(total_fold2 = n()) %>%
        group_by(age_group, fold_r2, total_fold2) %>%
        summarize(
          n_fold2 = n(),
          .groups = "drop"
        ) %>% select(age_group, fold_r2, total_fold2, n_fold2, ) %>%
        rename("fold" = fold_r2),
      by = c("age_group", "fold")
    ) %>%
    kable("html", caption = "Inspect Fold Number Frequencies") %>%
    kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE,
      position = "left",
      font_size = 12,
      fixed_thead = TRUE
    ) %>%
    scroll_box(width = "100%", height = "400px")

  p <- df %>%
    mutate(row_index = row_number()) %>%
    pivot_longer(
      cols = starts_with("fold_r"),
      names_to = "r",
      values_to = "fold"
    ) %>%
    ggplot(aes(x = row_index, y = fold)) +
    geom_point(
      size = 1,
      alpha = 0.1,
      color = "grey80",
      position = position_jitter(width = 0, height = 0.4)
    ) +
    facet_wrap(~r) +
    theme_few()

  plot(p)
}
```

Fold numbers are consistent between groups and correctly split upon multiple
repeats. each fold number has the same (within 1) frequency, so there are no
imbalances between folds and repeats. Plot shows no non-random changes in
intensity over the X dimension and random dissimilarities between bands in the Y
dimension (indicating repeats have different samples). **Stratified fold
randomization succeeded**

### (Optional) Check Interaction vs Offsets

We want to get a sense which confounders improve the model and which interact
with age.

```{r assess_confounders}
if (isTRUE(params$run_optional_chunks)) {
  assess_confounders <- function(df, confounders, k) {
    f0 <- paste0(params$hearing_var, " ~ s(age, k = ", k, ")")
    m0 <- bam(as.formula(f0), data = df, method = "REML")

    results <- data.frame()

    for (confounder in confounders) {
      f1 <- paste0(params$hearing_var, " ~ ", confounder, " + s(age, k = ", k, ")")
      f2 <- paste0(
        params$hearing_var, " ~ s(age) + s(age, ", confounder, ", bs = 'sz', k = ",
        k,
        ")"
      )

      m1 <- bam(as.formula(f1), data = df, method = "REML")
      m2 <- bam(as.formula(f2), data = df, method = "REML")

      aov <- as.data.frame(anova(m0, m1, m2))
      aic <- as.data.frame(AIC(m0, m1, m2))
      bic <- as.data.frame(BIC(m0, m1, m2))

      results <- rbind(results, cbind(tibble(
        confounder = confounder, formula = c(f0, f1, f2)
      ), aic, bic, aov))
    }

    return(results)
  }

  assess_confounders(df, confounders = c(
    "gender", "smoking", "race", "education"
  ), k = params$global_k)
}
```

**We see that for each confounder the interaction model fits better than the
intercept-only model**

### (Optional) Investigate Additional Information

First we see if the confounders are correlated (e.g., smoking is more prevalent
in males than females so gender is partly coded in smoking), to see if all
confounders actually contribute.

```{r compare_added_info}
if (isTRUE(params$run_optional_chunks)) {
  compare_confounders <- function(df, k) {
    m0 <- bam(reformulate("s(age, k = k)", response = params$hearing_var), data = df, method = "REML")
    m1 <- bam(
      reformulate(c("s(age, k = k)", "s(age, gender, bs = 'sz', k = k)"), response = params$hearing_var),
      data = df,
      method = "REML"
    )
    m2 <- bam(
      reformulate(c("s(age, k = k)", "s(age, gender, bs = 'sz', k = k)", "s(age, race, bs = 'sz', k = k)"), response = params$hearing_var),
      data = df,
      method = "REML"
    )
    m3 <- bam(
      reformulate(c("s(age, k = k)", "s(age, gender, bs = 'sz', k = k)", "s(age, race, bs = 'sz', k = k)", "s(age, smoking, bs = 'sz', k = k)"), response = params$hearing_var),
      data = df,
      method = "REML"
    )
    m4 <- bam(
      reformulate(c("s(age, k = k)", "s(age, gender, bs = 'sz', k = k)", "s(age, race, bs = 'sz', k = k)", "s(age, smoking, bs = 'sz', k = k)", "s(age, education, bs = 'sz', k = k)"), response = params$hearing_var),
      data = df,
      method = "REML"
    )

    print(cbind(
      as.data.frame(AIC(m0, m1, m2, m3, m4)),
      as.data.frame(BIC(m0, m1, m2, m3, m4)),
      as.data.frame(anova(m0, m1, m2, m3, m4))
    ))
    print(summary(m4))
    plot(m4)
  }

  compare_confounders(df, k = params$global_k)
}
```

**Each factor significantly improves the model and seems to interact with age in
some way**

### (Optional) Check shared wiggliness

We could insist that different smooths have the same amount of smoothness, i.e.,
that there's no difference in the wiggliness of the GAM between men and women.
The plots from the previous chunk (where different wiggliness was permitted)
suggest that there may be equal amount of wiggliness, so that each smooth has
the same basis functions but with different coefficients. There's no specific
reason to assume that wiggliness is the same in many cases, so if aiming to
maximize model flexibility/predictive power, we should leave this unconstrained,
but we do eat up a lot of degrees of freedom and may be 'wasting' the data
estimating these global trends. It may also be better to spend the processing
time on a higher k rather than estimating smooth parameters.

```{r compare_wiggliness}
if (isTRUE(params$run_optional_chunks)) {
  compare_wiggliness <- function(df, k) {
    m1 <- bam(
      as.formula(paste0(params$hearing_var, ' ~
      s(age, k = k) +

      s(age, gender, bs = "sz", k = k, id = 1) +
      s(age, race, bs = "sz", k = k, id = 2) +
      s(age, smoking, bs = "sz", k = k, id = 3) +
      s(age, education, bs = "sz", k = k, id = 4)')),
      data = df,
      method = "REML"
    )

    m2 <- bam(
      as.formula(paste0(params$hearing_var, ' ~
      s(age, k = k) +

      s(age, gender, bs = "sz", k = k) +
      s(age, race, bs = "sz", k = k) +
      s(age, smoking, bs = "sz", k = k) +
      s(age, education, bs = "sz", k = k)')),
      data = df,
      method = "REML"
    )

    print(cbind(
      as.data.frame(AIC(m1, m2)),
      as.data.frame(BIC(m1, m2)),
      as.data.frame(anova(m1, m2))
    ))
  }

  compare_wiggliness(df, k = params$global_k)
}
```

### (Optional) Check factor three-way interactions

<https://stats.stackexchange.com/questions/637423/conceptual-interpretation-of-bs-fs-and-by-term-in-gam>.
The free wiggliness model takes about 15 times longer to compute but has a 0.37%
drop in residual deviance. That is significant but probably not worth it
compared to a higher k. Moreover, the improvement in fit must stem entirely from
three-way interactions, which are likely noisy, so it is unlikely that out of
fold performance would improve substantially. BIC prefers simpler model and AIC
prefers more complex model.

```{r conf_interactions}
if (isTRUE(params$run_optional_chunks)) {
  compare_confounder_interactions <- function(df, k, equal_wiggliness, return_model = FALSE) {
    if (equal_wiggliness) {
      m0 <- bam(
        as.formula(paste0(params$hearing_var, ' ~
        s(age, k = k) +

        s(gender, age, bs = "sz", k = k, id = 1) +
        s(race, age, bs = "sz", k = k, id = 2) +
        s(smoking, age, bs = "sz", k = k, id = 3) +
        s(education, age, bs = "sz", k = k, id = 4) +

        s(gender, race, age, bs = "sz", k = k, id = 5) +
        s(gender, smoking, age, bs = "sz", k = k, id = 6) +
        s(gender, education, age, bs = "sz", k = k, id = 7) +
        s(race, smoking, age, bs = "sz", k = k, id = 8) +
        s(race, education, age, bs = "sz", k = k, id = 9) +
        s(smoking, education, age, bs = "sz", k = k, id = 10)')),
        data = df,
        method = "REML"
      )
    } else {
      m0 <- bam(
        as.formula(paste0(params$hearing_var, ' ~
        s(age, k = k) +

        s(gender, age, bs = "sz", k = k) +
        s(race, age, bs = "sz", k = k) +
        s(smoking, age, bs = "sz", k = k) +
        s(education, age, bs = "sz", k = k) +

        s(gender, race, age, bs = "sz", k = k) +
        s(gender, smoking, age, bs = "sz", k = k) +
        s(gender, education, age, bs = "sz", k = k) +
        s(race, smoking, age, bs = "sz", k = k) +
        s(race, education, age, bs = "sz", k = k) +
        s(smoking, education, age, bs = "sz", k = k)')),
        data = df,
        method = "REML"
      )
    }

    print(summary(m0))
    plot(m0)

    if (return_model) {
      return(m0)
    }
  }

  system.time({
    compare_confounder_interactions(df, k = params$global_k, equal_wiggliness = TRUE)
  })
}
```

We see that `age*gender*race` is a significant three-way interaction but it
takes 3 times longer to compute. Thanks to bam() we can afford this cost.

### (Optional) Inspect k-values

To assess k-value, we fit a model on the whole dataset using the formula lists,
build a list of models, make a helper function to run the k.check function from
mgcv, then turn it in a dataframe and return the output. The k-value determines
the number of base functions that the GAM is constructed from. A GAM adds these
different base functions together, some of which capture linear trends, some of
which are just splines or polynomials with a bump in one region. The GAM gives
different weights to each model to get as good a fit as possible, but a GAM also
measures wiggliness and gives a penalty to wiggliness. This is what makes it
generalized, so to perform well in general, even to unseen data. Therefore, a
GAM optimizes to get as good a fit as possible with as little wiggliness as
possible. Because of this wiggliness penalty, adding more base functions to the
GAM doesn't make it overfit. Therefore, k should be as large as possible within
computation power or degrees of freedom. If k isn't large enough, there will be
patterns left in the residuals. Therefore, by measuring patterns in the
residuals, we can assess if k is large enough. We want to capture all global
patterns without fitting to any noise. This is what we assess when we look at
k-index.

For reference, k = 10 takes \~ 15 min, k = 25 takes \~ 1h, k = 50 takes \~ 4h.

```{r inspect_k}
if (isTRUE(params$run_optional_chunks)) {
  inspect_k_values <- function(data, formulas, subsample, reps) {
    set.seed(2)

    models <- list()

    for (formula in formulas) {
      mod <- bam(
        as.formula(formula),
        data = data,
        method = "REML",
        family = gaussian()
      )

      predictors <- paste(extract_predictors(formula), collapse = ", ")

      models[[predictors]] <- mod
    }

    inspect_k <- function(model) {
      k.check(model, subsample = subsample, n.rep = reps)
    }

    k_output <- lapply(models, inspect_k)
    k_output <- imap_dfr(k_output, ~ {
      as.data.frame(.x) %>%
        rownames_to_column(var = "Predictors") %>%
        mutate(model = .y)
    })

    return(k_output)
  }

  inspect_k_values(df,
    formula_list_bioage,
    subsample = 5000,
    reps = 400
  ) %>%
    kable("html", caption = "Inspect K Bioage") %>%
    kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE,
      position = "left",
      font_size = 12,
      fixed_thead = TRUE
    ) %>%
    scroll_box(width = "100%", height = "400px")

  inspect_k_values(df,
    formula_list_biomarkers,
    subsample = 5000,
    reps = 400
  ) %>%
    kable("html", caption = "Inspect K Biomarkers") %>%
    kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE,
      position = "left",
      font_size = 12,
      fixed_thead = TRUE
    ) %>%
    scroll_box(width = "100%", height = "400px")
}
```

### (Optional) Estimate modeling time

bam speeds up the fitting compared to gam and limits the memory footprint.
Data.table is negligibly faster than data.frame (less than 1%) because the
majority of computing time is spent on fitting the model as opposed to data
operations.

With discrete = FALSE: k = 5, age + phenoage: 0.5 k = 10, age + phenoage: 0.78 k
= 15, age + phenoage: 1.34 k = 20, age + phenoage: 2.76 k = 25, age + phenoage:
4.98 k = 30, age + phenoage: 8.16 k = 40, age + phenoage: 18.28 k = 50, age +
phenoage: 31.97

With discrete = TRUE: k = 5, age + phenoage: 0.04 k = 10, age + phenoage: 0.06 k
= 15, age + phenoage: 0.09 k = 20, age + phenoage: 0.15 k = 25, age + phenoage:
0.24 k = 30, age + phenoage: 0.35 k = 40, age + phenoage: 0.78 k = 50, age +
phenoage: 1.28

```{r check_modeling_time}
if (isTRUE(params$run_optional_chunks)) {
  estimate_modeling_time <- function(df, k, discrete = FALSE) {
    system.time({
      bam(as.formula(paste0(params$hearing_var, ' ~
        s(age, k = k) +

        s(gender, age, bs = "sz", k = k, id = 1) +
        s(race, age, bs = "sz", k = k, id = 2) +
        s(smoking, age, bs = "sz", k = k, id = 3) +
        s(education, age, bs = "sz", k = k, id = 4) +

        s(gender, race, age, bs = "sz", k = k, id = 5) +

        s(phenoage, k = k)')),
        data = df, discrete = discrete, method = "fREML"
      )
    })
  }

  estimate_modeling_time(df, k = params$global_k, discrete = params$global_discrete)
}
```

*Without id = x, k = 15 took 39.65 theoretically, but in practice the 4 bioage
formulas took 58s each. Similarly, k = 5 with id should have taken \~400s for
one formula but took 650s, so add \~50% time to estimate processing time.*

### (Optional) Test k-fold validation

Test k-fold validation on a single formula with small k. k = 5, discrete = FALSE
takes 50s, k = 5, discrete = TRUE takes 15s

```{r test_k_fold_val}
if (isTRUE(params$run_optional_chunks)) {
  formula_list_test <- formula_generator(
    data = df,
    k = params$global_k,
    auto_limit_k = TRUE,
    equal_wiggliness = TRUE,
    formulas = c("age, phenoage")
  )

  system.time({
    output_test <- cross_validation(
      df,
      formula_list_test,
      k = params$global_folds,
      repeats = params$global_repeats,
      discrete = params$global_discrete
    )
  })
}
```
