---
editor_options:
  markdown:
    wrap: 100
---

# Data Processing

## Split up bioage results into two

```{r}
if (isFALSE(params$telolength)) {
split_bioage_reverse <- function(output) {
  list(
    results = output$results %>%
      filter(predictors != "age") %>%
      mutate(predictors = fct_drop(predictors)),
    predictions = output$predictions %>%
      filter(predictors != "age") %>%
      mutate(predictors = fct_drop(predictors))
  )
}

split_bioage <- function(output) {
list(
    results = output$results %>%
      filter(!predictors %in% c("phenoage", "kdm", "hd_log")) %>%
      mutate(predictors = fct_drop(predictors)),
    predictions = output$predictions %>%
      filter(!predictors %in% c("phenoage", "kdm", "hd_log")) %>%
      mutate(predictors = fct_drop(predictors))
  )
}

output_bioage_reverse <- split_bioage_reverse(output_bioage_bidirectional)
output_bioage <- split_bioage(output_bioage_bidirectional)
}
```

## Process result comparison dfs

### Reformat df, calculate delta MSE bioage, and set data types

We flip the sign for confounder analysis `delta_mse` and `additional_variance` when we subtract a
variable.

```{r reformat_df}
reformat_df <- function(data) {
  if ("full" %in% data$results$predictors) {
    data$results %>%
    left_join(
      data$results %>%
        filter(
            predictors == "full"
        ) %>%
        mutate(mse_ref = mse) %>%
        select(-mse, -predictors, -n_train, -n_test),
      by = c("age_group", "r", "k")
    ) %>%
    mutate(
      reference = "full",
      delta_mse = mse_ref - mse,
      additional_variance = delta_mse / mse_ref * 100,
              delta_mse = case_when(
          !predictors %in% c("full", "full_with_phenoage") ~ delta_mse * -1,
          TRUE ~ delta_mse
        ),
        additional_variance = case_when(
          !predictors %in% c("full", "full_with_phenoage") ~ additional_variance * -1,
          TRUE ~ additional_variance
        ),
      fold_repeat_id = factor(
        interaction(r, k)
      ),
    )
  } else {
  single_predictors <- data$results %>%
    filter(!grepl(", ", predictors)) %>%
    pull(predictors) %>%
    unique()
  
  map(single_predictors, function(ref) {
    data$results %>%
      filter(grepl(paste0("(^|, )", ref, "(, |$)"), predictors) | predictors == ref) %>%
      left_join(
        data$results %>%
          filter(predictors == ref) %>%
          mutate(mse_ref = mse) %>%
          select(-mse, -predictors, -n_train, -n_test),
        by = c("age_group", "r", "k")
      ) %>%
      mutate(
        reference = ref,
        delta_mse = mse_ref - mse,
        additional_variance = delta_mse / mse_ref * 100,
        fold_repeat_id = factor(interaction(r, k)),
      )
  }) %>%
    list_rbind()
  }
}

df_results_bioage <- output_bioage %>% reformat_df()
df_results_bioage_reverse <- output_bioage_reverse %>% reformat_df()

if (isFALSE(params$telolength)) {
  df_results_biomarkers <- output_biomarkers %>% reformat_df()
  df_results_confounders <- output_confounders %>% reformat_df()
}
```

## Build predictions dataframe age rate

We sample 1 point for each age and predict the FI using the smooth-only model and generate
simultaneous confidence intervals with 'mvt'. We could ask for a single point and ask where the true
mean would be with 95% certainty, but if we'd apply that to all 60 points, there's a large chance
that somewhere we get a false positive. With the simultaneous interval or with multivariate t
distribution error correction, we can say that we're 95% sure that the **entire** derivative in the
true population that was sampled from (all people on earth if NHANES is representative) is within
the band.

To calculate the quadratic derivative, we can either calculate the slope at each point manually with

`coef(m)["age"] + 2 * coef(m)["I(age^2)"] * seq(min(df$age), max(df$age), 1)}`

Or we use emtrends to calculate the slope at each point. The result is the same. However, using
emtrends we can calculate simultaneous intervals with "mvt". Emtrends also works with smooth terms.

```{r predict_aging}
predict_age_rate <- function(
  df,
  model,
  predictor = "age",
  type = c("predict", "predict_derivative")
) {
  ages <- setNames(
    list(
      seq(
        min(df[[predictor]]),
        max(df[[predictor]]),
        by = 1
      )
    ),
    predictor
  )
  formula <- as.formula(
    paste("~", predictor)
  )

  if (type == "predict") {
    em <- emmeans(
      model,
      formula,
      at = ages,
      type = "response",
      data = df
    )
  } else {
    em <- emtrends(
      model,
      formula,
      var = predictor,
      at = ages,
      data = df
    )
  }

  predictions <- tibble(
    confint(em, adjust = "mvt")
  )

  if (type == "predict") {
    predictions <- predictions %>%
      rename(
        "estimate" = emmean,
        "ci_lower" = lower.CL,
        "ci_upper" = upper.CL
      )
  } else {
    predictions <- predictions %>%
      rename(
        "derivative" = !!paste0(predictor, ".trend"),
        "ci_lower" = lower.CL,
        "ci_upper" = upper.CL
      )
  }
}

output_age_rate$predictions <- append(
  output_age_rate$predictions,
  list(
    linear = predict_age_rate(
      df,
      output_age_rate$models$linear,
      type = "predict"
    ),
    linear_derivative = predict_age_rate(
      df,
      output_age_rate$models$linear,
      type = "predict_derivative"
    ),
    quadratic = predict_age_rate(
      df,
      output_age_rate$models$quadratic,
      type = "predict"
    ),
    quadratic_derivative = predict_age_rate(
      df,
      output_age_rate$models$quadratic,
      type = "predict_derivative"
    ),
    smooth = predict_age_rate(
      df,
      output_age_rate$models$smooth,
      type = "predict"
    ),
    derivative_smooth = predict_age_rate(
      df,
      output_age_rate$models$smooth,
      type = "predict_derivative"
    )
  )
)

output_age_rate$predictions$derivative_smooth <- output_age_rate$predictions$derivative_smooth %>%
  mutate(
    derivative_adj = derivative - output_age_rate$predictions$quadratic_derivative$derivative,
    ci_lower_adj = ci_lower - output_age_rate$predictions$quadratic_derivative$derivative,
    ci_upper_adj = ci_upper - output_age_rate$predictions$quadratic_derivative$derivative
  )
```

## Calculate coefficients

We estimate the rate of change in dB/year and acceleration in dB/year² for the quadratic model and
GAM. For the GAM, we estimate the acceleration by fitting a linear model to the derivative and
calculating its coefficient.

```{r calc_coeffs}
with(output_age_rate$predictions$linear, cov(age, estimate) / var(age))
with(output_age_rate$predictions$quadratic_derivative, cov(age, derivative) / var(age))
coef(lm(derivative ~ age, data = output_age_rate$predictions$derivative_smooth))[["age"]]
```
