# Data Processing

## Process Result Comparison dfs

### Reformat df, Calculate Delta MSE Bioage, and set data types

We flip the sign for confounder analysis `delta_mse` and `additional_variance` when we subtract a variable. 

```{r reformat_df}
reformat_df <- function(data) {
  reformatted_data <- data$results %>%
    left_join(
      data$results %>%
        filter(if ("age" %in% data$results$predictors) {
          predictors == "age"
        } else {
          predictors == "full"
        }) %>%
        mutate(mse_age = mse) %>%
        select(-mse, -predictors, -n_train, -n_test),
      by = c("age_group", "r", "k")
    ) %>%
    mutate(
      delta_mse = mse_age - mse,
      additional_variance = delta_mse / mse_age * 100,
      fold_repeat_id = factor(interaction(r, k)),
    )

  if ("full" %in% data$results$predictors) {
    reformatted_data <- reformatted_data %>%
      rename("mse_full" = mse_age) %>%
      mutate(
        delta_mse = case_when(
          !predictors %in% c("full", "full_with_phenoage") ~ delta_mse * -1,
          TRUE ~ delta_mse
        ),
        additional_variance = case_when(
          !predictors %in% c("full", "full_with_phenoage") ~ additional_variance * -1,
          TRUE ~ additional_variance
        )
      )
  }
  return(reformatted_data)
}

df_results_bioage <- reformat_df(output_bioage)

if (isFALSE(params$telolength)) {
  df_results_biomarkers <- reformat_df(output_biomarkers)
  df_results_confounders <- reformat_df(output_confounders)
}
```


## Build predictions dataframe age rate

We sample 1 point for each age and predict the FI using the smooth-only
model and generate simultaneous confidence intervals. We could ask for a single
point and ask where the true mean would be with 95% certainty, but if we'd apply
that to all 60 points, there's a large chance that somewhere we get a false
positive. With the simultaneous interval or with multivariate t distribution error correction, we can say that we're 95% sure that the ENTIRE derivative in the true population that was sampled from (all people
on earth, assuming NHANES is representative) is within the band. It's
essentially a multiple comparisons test for confidence intervals.

To calculate the quadratic derivative, we can either calculate the slope at each
point manually with

`coef(m)["age"] + 2 * coef(m)["I(age^2)"] * seq(min(df$age), max(df$age), 1)}`

Or we use emtrends to calculate the slope at each point. The result is the same.
However, using emtrends we can calculate simultaneous intervals with "mvt". Emtrends also works with smooth terms.

```{r predict_aging}
build_parametric_predictions <- function(df, model, predictor) {
  as.data.frame(confint(
    emmeans(
      model,
      as.formula(paste("~", predictor)),
      at = setNames(list(seq(
        min(df[[predictor]]), max(df[[predictor]]),
        by = 1
      )), predictor),
      type = "response",
      data = df
    ),
    adjust = "mvt"
  )) %>%
    rename(
      "estimate" = emmean,
      "ci_lower" = lower.CL,
      "ci_upper" = upper.CL
    )
}

build_parametric_derivative_predictions <- function(df, model, predictor) {
  data.frame(confint(
    emtrends(
      model,
      as.formula(paste("~", predictor)),
      var = predictor,
      at = setNames(list(seq(
        min(df[[predictor]]), max(df[[predictor]]),
        by = 1
      )), predictor),
      data = df
    ),
    adjust = "mvt"
  )) %>%
    rename(
      "derivative" = !!paste0(predictor, ".trend"),
      "ci_lower" = lower.CL,
      "ci_upper" = upper.CL
    )
}

output_age_rate$predictions <- append(
  output_age_rate$predictions,
  list(linear = build_parametric_predictions(df, output_age_rate$models$linear, "age"))
)

output_age_rate$predictions <- append(
  output_age_rate$predictions,
  list(quadratic = build_parametric_predictions(df, output_age_rate$models$quadratic, "age"))
)

output_age_rate$predictions <- append(
  output_age_rate$predictions,
  list(linear_derivative = build_parametric_derivative_predictions(df, output_age_rate$models$linear, "age"))
)

output_age_rate$predictions <- append(
  output_age_rate$predictions,
  list(
    quadratic_derivative = build_parametric_derivative_predictions(df, output_age_rate$models$quadratic, "age")
  )
)

output_age_rate$predictions <- append(
  output_age_rate$predictions,
  list(
    smooth =
      build_parametric_predictions(
        df, output_age_rate$models$smooth, "age"
      )
  )
)

output_age_rate$predictions <- append(
  output_age_rate$predictions,
  list(
    derivative_smooth =
      build_parametric_derivative_predictions(
        df, output_age_rate$models$smooth, "age"
      ) %>%
        mutate(
          derivative_adj = derivative - output_age_rate$predictions$quadratic_derivative$derivative,
          ci_lower_adj = ci_lower - output_age_rate$predictions$quadratic_derivative$derivative,
          ci_upper_adj = ci_upper - output_age_rate$predictions$quadratic_derivative$derivative
        )
  )
)
```
