# Data Processing

## Process Result Comparison dfs

### Reformat df, Calculate Delta MSE Bioage, and set data types

We flip the sign for confounder analysis `delta_mse` and `additional_variance` when we subtract a variable. 

```{r reformat_df}
reformat_df <- function(data) {
  reformatted_data <- data$results %>%
    left_join(
      data$results %>%
        filter(if ("age" %in% data$results$predictors) {
          predictors == "age"
        } else {
          predictors == "full"
        }) %>%
        mutate(mse_age = mse) %>%
        select(-mse, -predictors, -n_train, -n_test),
      by = c("age_group", "r", "k")
    ) %>%
    mutate(
      delta_mse = mse_age - mse,
      additional_variance = delta_mse / mse_age * 100,
      fold_repeat_id = factor(interaction(r, k)),
    )
  
  if ("full" %in% data$results$predictors) {
    reformatted_data <- reformatted_data %>% rename("mse_full" = mse_age) %>%
      mutate(
        delta_mse = case_when(
          !predictors %in% c("full", "full_with_phenoage") ~ delta_mse * -1,
          TRUE ~ delta_mse
        ),
        additional_variance = case_when(
          !predictors %in% c("full", "full_with_phenoage") ~ additional_variance * -1,
          TRUE ~ additional_variance
        )
      )
  }
  return(reformatted_data)
}

df_results_bioage <- reformat_df(output_bioage)

if (isFALSE(params$telolength)) {
  df_results_biomarkers  <- reformat_df(output_biomarkers)
  df_results_confounders <- reformat_df(output_confounders)
}
```

### Calculate CI

For the overall, non-age-group-stratified summary statistics, we first calculate
mse per fold by weighting the age groups. Then, we calculate the number of
repeats and folds, the mse of each outcome variable (weighted per number of
observations, which produces a small difference over unweighted because the
n_test per fold differs by 1 at most), and the variance of mse `s2_mse`, also
weighted per n_test, grouped by predictors and possibly age group.

The Nadeau Bengio corrected t-test estimates variance using the training and
testing dataset sizes, because cross validations are not independent, violating
the assumption of a paired t-test, inflating type 1 errors (false positives). We
can use the corrected variance estimate for the confidence intervals too. First,
we calculate t-crit using degrees of freedom (n-1). The t-test should be
one-tailed because we only want to know if age + bioage is better than age, not
if one is worse/different, so then a tcrit of 0.05 is appropriate, but for the
confidence intervals we want to know whether the true mean is within 95% of the
sample mean, not just where the true mean would be if it would only be above.
The variance is used based on `n_train` and `n_test`. It doesn't matter whether
we use the mean n per fold or the total over all folds because the ratio is the
same.

In <https://www.jmlr.org/papers/volume6/markatou05a/markatou05a.pdf>, Analysis of
Variance of Cross-Validation Estimators of the Generalization Error, the authors
also report the variance estimate as `S² * (1/J + n2/n1)`, where `S2` is the
variance (squared standard deviation), `J` is the total number of folds, `n2` is the
validation/test set and `n1` is the training set size. They also show that when `J`
is appropriate (`10*10`), it provides a good variance estimate. Variance is
already in squared units, as its just the square of the standard deviation.

```{r calculate_ci}
calculate_nb_corrected_cis <- function(data, per_age_group) {
  if (per_age_group) {
    grouping_variables <- c("predictors", "age_group")
  } else {
    grouping_variables <- c("predictors")
  }
  
  if (!per_age_group) {
    data <- data %>%
      group_by(predictors, r, k) %>%
      summarize(
        mse = weighted.mean(mse, w = n_test),
        delta_mse = weighted.mean(delta_mse, w = n_test),
        additional_variance = weighted.mean(additional_variance, w = n_test),
        n_train = sum(n_train),
        n_test = sum(n_test),
        .groups = "drop"
      )
  }
  
  summary_data <- data %>%
    group_by(all_of(grouping_variables)) %>%
    summarize(
      r = length(unique(r)),
      k = length(unique(k)),
      mean_mse = weighted.mean(mse, w = n_test),
      mean_delta_mse = weighted.mean(delta_mse, w = n_test),
      mean_additional_variance = weighted.mean(additional_variance, w = n_test),
      s2_mse = weightedVar(mse, w = n_test),
      s2_delta_mse = weightedVar(delta_mse, w = n_test),
      s2_additional_variance = weightedVar(additional_variance, w = n_test),
      sum_n_train = sum(n_train),
      sum_n_test = sum(n_test),
      .groups = "drop"
    ) %>%
    mutate(
      n = r * k,
      df = n - 1,
      t_crit = qt(1 - 0.025, df),
      ci_low_mse = mean_mse - t_crit * (sqrt(s2_mse * (
        1 / n + sum_n_test / sum_n_train
      ))),
      ci_high_mse = mean_mse + t_crit * (sqrt(s2_mse * (
        1 / n + sum_n_test / sum_n_train
      ))),
      ci_low_delta_mse = mean_delta_mse - t_crit * (sqrt(
        s2_delta_mse * (1 / n + sum_n_test / sum_n_train)
      )),
      ci_high_delta_mse = mean_delta_mse + t_crit * (sqrt(
        s2_delta_mse * (1 / n + sum_n_test / sum_n_train)
      )),
      ci_low_additional_variance = mean_additional_variance - t_crit * (sqrt(
        s2_additional_variance * (1 / n + sum_n_test / sum_n_train)
      )),
      ci_high_additional_variance = mean_additional_variance + t_crit * (sqrt(
        s2_additional_variance * (1 / n + sum_n_test / sum_n_train)
      ))
    ) %>% select(-r, -k, -s2_mse, -s2_additional_variance, -df, -t_crit)
  
  return(summary_data)
}

df_summary_bioage <- calculate_nb_corrected_cis(df_results_bioage, per_age_group = TRUE)
df_summary_bioage_overall <- calculate_nb_corrected_cis(df_results_bioage, per_age_group = FALSE)

if (isFALSE(params$telolength)) {
df_summary_biomarkers <- calculate_nb_corrected_cis(df_results_biomarkers, per_age_group = TRUE)
df_summary_biomarkers_overall <- calculate_nb_corrected_cis(df_results_biomarkers, per_age_group = FALSE)

df_summary_confounders <- calculate_nb_corrected_cis(df_results_confounders, per_age_group = TRUE)
df_summary_confounders_overall <- calculate_nb_corrected_cis(df_results_confounders, per_age_group = FALSE)
}
```

### Calculate R2 and RMSE

**Calculating descriptives**: First we import `n_train` and `n_test` from the
results dataframe per fold, either grouped or not grouped by `age_group`. If
grouped by `age_group`, we can import directly because the results dataframe is
already binned by age group. Then, we import the mean Fletcher Index from the
whole prediction dataset for each model/set of predictors, repeat, and fold. As
a sanity check, because each model should have the same folds and repeats, I
checked whether the mean Fletcher Index is always the same at repeat 1 fold 1
(it is). The mean FI is then called in the summarize function using median(),
which gives the same result as mean(), because all predicted/real values within
a model, fold, and repeat have the same `mean_FI`, as told by the left_join
function. Thereafter, we group over the 100 (10x10) folds per model to calculate
the total number of predictions (on the testing data), the mean training data
size, mean testing data size, R² (using sum of squares), MSE, and RMSE. For
RMSE, we could either calculate it using `sqrt(mse)` or `sqrt(sse/n_predictions)`.
It gives the same result.

**Total sum of squares calculation when binning R² per age group**: It's
imperative that for the age group stratified descriptives, the global mean is
used. Otherwise it's as if the total sum of squares is calculated within each
age-bin, which is as if a linear regression is fitted through the mean Fletcher
Index per age group, explaining a lot of variation while it shouldn't. We want
to assess the predictive effect on age, so the total sum of squares should be
calculated by taking the distance from each Fletcher Index within an age group
to the global mean. Of course, the models are trained on the whole dataset, not
just one age group. We're just assessing performance per age group.

<https://chatgpt.com/share/68da6507-0008-8010-857e-bc6e9b76df58> - see
interpretation of R² differences


```{r calculate_descr}
calculate_RMSE_R2 <- function(summary_data,
                              results_data,
                              predictions,
                              per_age_group) {
  if (per_age_group) {
    grouping_variables <- c("predictors", "age_group")
  } else {
    grouping_variables <- c("predictors")
  }
  
  if (!per_age_group) {
    results_data <- results_data %>%
      group_by(predictors, r, k) %>%
      summarize(
        n_train = sum(n_train),
        n_test = sum(n_test),
        .groups = "drop"
      )
  }
  
  sample_sizes <- results_data %>%
    select(all_of(grouping_variables), r, k, n_train, n_test)
  
  mean_hearing_var_df <- predictions %>%
    group_by(predictors, k, r) %>%
    summarize(mean_hearing_var = mean(.data[[params$hearing_var]]), .groups = "drop")
  
  updated_summary_data <- summary_data %>%
    left_join(
      predictions %>%
        left_join(sample_sizes, by = c(grouping_variables, "r", "k")) %>%
        left_join(mean_hearing_var_df, by = c("predictors", "r", "k")) %>%
        group_by(all_of(
          grouping_variables
        )) %>%
        summarize(
          n_predictions = n(),
          mean_n_train = round(mean(n_train), 1),
          mean_n_test = round(mean(n_test), 1),
          sse = sum((.data[[params$hearing_var]] - predicted) ^ 2),
          sst = sum((.data[[params$hearing_var]] - median(mean_hearing_var)) ^ 2),
          .groups = "drop"
        )
      %>% mutate(
          r2  = 1 - sse / sst,
          rmse = sqrt(sse / n_predictions))
      %>% select(-sse, -sst),
      by = c(grouping_variables)
    )
  
  if (per_age_group) {
    updated_summary_data <- updated_summary_data %>%
      left_join(
        updated_summary_data %>%
          filter(if ("age" %in% summary_data$predictors) {
            predictors == "age"
          } else {
            predictors == "full"
          }) %>%
          mutate(rmse_age = rmse) %>%
          select(age_group, rmse_age),
        by = c("age_group")
      ) %>%
      mutate(delta_rmse = rmse_age - rmse) %>%
      select(-rmse_age)
  } else {
    updated_summary_data <- updated_summary_data %>% cross_join(
      updated_summary_data %>%
     filter(if ("age" %in% summary_data$predictors) {
            predictors == "age"
          } else {
            predictors == "full"
          }) %>%
        mutate(rmse_age = rmse) %>%
        select(rmse_age)
    ) %>%
      mutate(delta_rmse = rmse_age - rmse) %>%
      select(-rmse_age)
  }
  
  return(updated_summary_data)
}

df_summary_bioage <- calculate_RMSE_R2(df_summary_bioage,
                                       df_results_bioage,
                                       output_bioage$predictions,
                                       per_age_group = TRUE)

df_summary_bioage_overall <- calculate_RMSE_R2(
  df_summary_bioage_overall,
  df_results_bioage,
  output_bioage$predictions,
  per_age_group = FALSE
)

if (isFALSE(params$telolength)) {
  df_summary_biomarkers <- calculate_RMSE_R2(
  df_summary_biomarkers,
  df_results_biomarkers,
  output_biomarkers$predictions,
  per_age_group = TRUE
)

df_summary_biomarkers_overall <- calculate_RMSE_R2(
  df_summary_biomarkers_overall,
  df_results_biomarkers,
  output_biomarkers$predictions,
  per_age_group = FALSE
)

  df_summary_confounders <- calculate_RMSE_R2(
  df_summary_confounders,
  df_results_confounders,
  output_confounders$predictions,
  per_age_group = TRUE
)

df_summary_confounders_overall <- calculate_RMSE_R2(
  df_summary_confounders_overall,
  df_results_confounders,
  output_confounders$predictions,
  per_age_group = FALSE
)
}
```


### Calculate RMSE CIs

```{r calc_RMSE_CI}
compute_rmse_CI <- function(data, summary_data, n_boot, per_age_group = FALSE, confounders = FALSE) {
  ref_var <- if (isTRUE(confounders)) "full" else "age"
  
  predictor_list <- setdiff(unique(data$predictors), ref_var)
  
  ci_df <- data.frame()
  
  if (isTRUE(per_age_group)) {
    for (predictor in predictor_list) {
      for (age_bin in unique(data$age_group)) {
        subset_ref <- data %>%
          filter(predictors == ref_var, age_group == age_bin) %>%
          select(all_of(params$hearing_var), predicted)
        
        subset <- data %>%
          filter(predictors == predictor, age_group == age_bin) %>%
          select(all_of(params$hearing_var), predicted)
        
        length <- nrow(subset)
        
        rmses <- replicate(n_boot, {
          i <- sample(length, replace = TRUE)
          c(sqrt(mean((
            subset[[params$hearing_var]][i] - subset[["predicted"]][i]
          )^2)), sqrt(mean((subset_ref[[params$hearing_var]][i] - subset_ref[["predicted"]][i])^2
          )))
        })
        
        ci_lower <- quantile(rmses[1, ], 0.025)
        ci_upper <- quantile(rmses[1, ], 0.975)
        
        delta_ci_lower <- quantile(rmses[2, ] - rmses[1, ], 0.025)
        delta_ci_upper <- quantile(rmses[2, ] - rmses[1, ], 0.975)
        
        ci_subset <- data.frame(
          ci_low_rmse = ci_lower,
          ci_high_rmse = ci_upper,
          ci_low_delta_rmse = delta_ci_lower,
          ci_high_delta_rmse = delta_ci_upper,
          predictors = predictor,
          age_group = age_bin
        )
        ci_df <- rbind(ci_df, ci_subset)
      }
    }
  } else {
    for (predictor in predictor_list) {
      subset_ref <- data %>%
        filter(predictors == ref_var) %>%
        select(all_of(params$hearing_var), predicted)
      
      subset <- data %>%
        filter(predictors == predictor) %>%
        select(all_of(params$hearing_var), predicted)
      
      length <- nrow(subset)
      
      rmses <- replicate(n_boot, {
        i <- sample(length, replace = TRUE)
        c(sqrt(mean((
          subset[[params$hearing_var]][i] - subset[["predicted"]][i]
        )^2)), sqrt(mean((
          subset_ref[[params$hearing_var]][i] - subset_ref[["predicted"]][i]
        )^2)))
      })
      
      ci_lower <- quantile(rmses[1, ], 0.025)
      ci_upper <- quantile(rmses[1, ], 0.975)
      
      delta_ci_lower <- quantile(rmses[2, ] - rmses[1, ], 0.025)
      delta_ci_upper <- quantile(rmses[2, ] - rmses[1, ], 0.975)
      
      ci_subset <- data.frame(
        ci_low_rmse = ci_lower,
        ci_high_rmse = ci_upper,
        ci_low_delta_rmse = delta_ci_lower,
        ci_high_delta_rmse = delta_ci_upper,
        predictors = predictor
      )
      ci_df <- rbind(ci_df, ci_subset)
    }
  }
  
  return(ci_df)
}

test <- compute_rmse_CI(output_bioage$predictions, n_boot = 500, per_age_group = FALSE, confounders = FALSE)
```


## Build predictions dataframe age rate

We sample 1 point for each age and predict the FI using the smooth-only
model and generate simultaneous confidence intervals. We could ask for a single
point and ask where the true mean would be with 95% certainty, but if we'd apply
that to all 60 points, there's a large chance that somewhere we get a false
positive. With the simultaneous interval or with multivariate t distribution error correction, we can say that we're 95% sure that the ENTIRE derivative in the true population that was sampled from (all people
on earth, assuming NHANES is representative) is within the band. It's
essentially a multiple comparisons test for confidence intervals.

To calculate the quadratic derivative, we can either calculate the slope at each
point manually with

`coef(m)["age"] + 2 * coef(m)["I(age^2)"] * seq(min(df$age), max(df$age), 1)}`

Or we use emtrends to calculate the slope at each point. The result is the same.
However, using emtrends we can calculate simultaneous intervals with "mvt". Emtrends also works with smooth terms.

```{r predict_aging}
build_parametric_predictions <- function(df, model, predictor) {
  as.data.frame(confint(
    emmeans(
      model,
      as.formula(paste("~", predictor)),
      at = setNames(list(seq(
        min(df[[predictor]]), max(df[[predictor]]), by = 1
      )), predictor)
      ,
      type = "response",
      data = df
    ),
    adjust = "mvt"
  )) %>%
    rename("estimate" = emmean,
           "ci_lower" = lower.CL,
           "ci_upper" = upper.CL)
}

build_parametric_derivative_predictions <- function(df, model, predictor) {
  data.frame(confint(
    emtrends(
      model,
      as.formula(paste("~", predictor)),
      var = predictor,
      at = setNames(list(seq(
        min(df[[predictor]]), max(df[[predictor]]), by = 1
      )), predictor),
      data = df
    ),
    adjust = "mvt"
  )) %>%
    rename(
      "derivative" = !!paste0(predictor, ".trend"),
      "ci_lower" = lower.CL,
      "ci_upper" = upper.CL
    )
}

output_age_rate$predictions <- append(output_age_rate$predictions,
                                      list(linear = build_parametric_predictions(df, output_age_rate$models$linear, "age")))

output_age_rate$predictions <- append(output_age_rate$predictions,
                                      list(quadratic = build_parametric_predictions(df, output_age_rate$models$quadratic, "age")))

output_age_rate$predictions <- append(
  output_age_rate$predictions,
  list(linear_derivative = build_parametric_derivative_predictions(df, output_age_rate$models$linear, "age"))
)

output_age_rate$predictions <- append(
  output_age_rate$predictions,
  list(
    quadratic_derivative = build_parametric_derivative_predictions(df, output_age_rate$models$quadratic, "age")
  )
)

output_age_rate$predictions <- append(output_age_rate$predictions,
                                      list(
                                        smooth =
build_parametric_predictions(
        df, output_age_rate$models$smooth, "age")))

output_age_rate$predictions <- append(
  output_age_rate$predictions,
  list(
    derivative_smooth =
      build_parametric_derivative_predictions(
        df, output_age_rate$models$smooth, "age") %>%
      mutate(
        derivative_adj = derivative - output_age_rate$predictions$quadratic_derivative$derivative,
        ci_lower_adj = ci_lower - output_age_rate$predictions$quadratic_derivative$derivative,
        ci_upper_adj = ci_upper - output_age_rate$predictions$quadratic_derivative$derivative
      )
  )
)
```
