# Statistics

## Assess Rate of Aging

### Compare Linear vs Quadratic vs Smooth Age Fits

We do a partial F-test to compare a full model with a constrained model, to see if the added term significantly improves fit.

```{r compare_aging_models}
anova(
  output_age_rate$models$linear,
  output_age_rate$models$quadratic
)

anova(
  output_age_rate$models$quadratic,
  output_age_rate$models$smooth
)

c(linear = sqrt(mean(
  residuals(output_age_rate$models$linear, type = "response") ^ 2
)),quadratic = sqrt(mean(
  residuals(output_age_rate$models$quadratic, type = "response") ^ 2
)), smooth = sqrt(mean(
  residuals(output_age_rate$models$smooth, type = "response") ^ 2
))) %>% kable("html", caption = "RMSE (In-Sample)") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position = "left",
    font_size = 12,
    fixed_thead = TRUE
  ) %>%
  scroll_box(width = "100%", height = "400px")
```

*We see that the quadratic curve fits better than the linear one, so the rate of
aging is not linear, and the smooth curve fits better than the quadratic one, so there are deviations from a quadratic pattern*

### Assess Where Rate of Aging Differs

By taking the derivative of the full model (with smooth) and subtracting the derivative of the nested quadratic model, we can test against zero to test deviations from a quadratic pattern.

When making a smooth (or linear) model, we
fit a line to noisy data as best we can, but other lines with different
coefficients might fit almost as well and be better for some data points then
for others. Therefore, there is an uncertainty. We can say that, for example,
the best-guess slope is 5 (5 increase fletcher index per year). But because the
data is noisy, assuming a normal distribution, we know that the true slope that
would actually fit the sampling population may be something different within a
certain range of uncertainty (say, 2), which we can express as a standard error.
When figuring out the derivative, we inherent this uncertainty. This standard
error is what we can use to assess the likelihood of the null-hypothesis:

H0: The rate of aging is quadratic, so the GAM derivative minus the quadratic coefficients equals 0. H1: The rate of aging is non-quadratic, so the GAM derivative minus the quadratic coefficients is not 0.

With a p-value of 0.05, we can say that if we randomly sample 100 times from the
population and the rate of aging IS linear, then 5 times we'll see a significant
non-zero smooth term.

Depending on the variation/noise in the data, the smooth term derivative needs
to be larger or smaller compared to 0 to get a p-value of 0.05. If the variation
in the data is higher, the standard error is higher, and the derivative of the
smooth term must be higher to get a p-value of 0.05 or lower.

Of course, we want to know at which age ranges we have a non-zero slope, so we
can either test for each 10 year age group or test for each individual age
integer with correct pairwise comparison corrections.

A two-tailed p-value can be calculated when we know a z-score, which is
calculated as sample mean - hypothesized mean divided by the standard error. The
hypothesized mean is 0. For a two-tailed t-test, we then take `2 * pnorm(abs(z_score), lower.tail = FALSE)`. If we'd want to only know the left
tail, we'd use `lower.tail = TRUE`.

Because the t-tests are dependent and are likely to be either higher or lower
when one is higher (because ages are adjacent and the aging rate can be either
higher or lower), we should use Holm's correction.

To identify significant boundaries, check which pairwise comparisons are
significant, then check if the previous row was significant, then whenever a row
is significant but the previous one was, change the value by 1 (using cumsum),
or if the previous row was not significant but the current one is, also change
by 1 to identify stretches. Then take the first and last values of the stretches
to identify boundaries. Only keep the significant stretches.

```{r find_sig_ages}
pwc_age_rate <- output_age_rate$predictions$derivative_smooth %>%
  mutate(
    z_score = derivative_adj / SE,
    p = 2 * pnorm(abs(z_score), lower.tail = FALSE),
    p_adj = p.adjust(p, method = "holm"),
    signif = case_when(
      p_adj <= 0.001 ~ "***",
      p_adj <= 0.01 ~ "**",
      p_adj <= 0.05 ~ "*",
      TRUE ~ "ns"
    )) %>%
  select(age, derivative, ci_lower, ci_upper, derivative_adj, ci_lower_adj, ci_upper_adj, p, p_adj, signif)

pwc_age_rate_boundaries <- pwc_age_rate %>%
  mutate(significant = (p_adj < 0.05),
         run = cumsum(significant != lag(significant, default = first(significant)))) %>%
  group_by(run, significant) %>%
  summarise(
    x_min = first(age),
    x_max = last(age),
    .groups = "drop") %>%
  mutate(
    `Significantly non-linear age ranges` = paste0(x_min, "–", x_max)
  ) %>%
  filter(significant)

pwc_age_rate_boundaries %>%
        select(`Significantly non-linear age ranges`) %>% kable("html", caption = "Boundaries") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position = "left",
    font_size = 12,
    fixed_thead = TRUE
  ) %>%
  scroll_box(width = "100%", height = "400px")
```

## Repeated Nadeau-Bengio Corrected T-Tests

### Check dependence

Holm error correction is the best method to control family-wise error rate when
comparisons are negatively dependent. However, if tests are either independent
or positively dependent, we can use Hommel, which has more power. To test
dependence, we correlate MSE of each metric over the same fold.

```{r assess_cor}
test_correlation <- function(data, grouping_variables, exclusions = NULL) {
  data <- data %>%
    filter(!predictors %in% exclusions) %>%
    group_by(all_of(grouping_variables), r, k) %>%
    summarize(
      mse = weighted.mean(mse, w = n_test),
      delta_mse = weighted.mean(delta_mse, w = n_test),
      additional_variance = weighted.mean(additional_variance, w = n_test),
      n_train = sum(n_train),
      n_test = sum(n_test),
      .groups = "drop"
    ) %>%
    unite("interaction_column", all_of(grouping_variables), sep = "_") %>%
    mutate(n = (r - 1) * 10 + k) %>%
    select(n, interaction_column, delta_mse) %>%
    pivot_wider(names_from = interaction_column, values_from = delta_mse) %>%
    select(-n)
  
  correlation_df <- as.data.frame(cor(data, use = "pairwise.complete.obs"))
  
  eps = 0.2
  
  upper_indices <- which(upper.tri(correlation_df), arr.ind = TRUE)
  
  df_long <- tibble(
    row = rownames(correlation_df)[upper_indices[, 1]],
    col = colnames(correlation_df)[upper_indices[, 2]],
    values = correlation_df[upper_indices]
  )
  
  p <- ggplot(df_long, aes(values)) +
      geom_density(alpha = 0.15) +
      geom_vline(
        xintercept = c(-eps, eps),
        colour = "red",
        linetype = "dashed"
      ) + theme_few() + labs(title =  paste0("Mean correlation: ", round(
        mean(correlation_df[upper.tri(correlation_df)], na.rm = TRUE), 2
      )))
  
  return(p)
}

test_correlation(df_results_bioage,
                 grouping_variables = "predictors",
                 exclusions = "age")

test_correlation(
  df_results_bioage,
  grouping_variables = c("age_group", "predictors"),
  exclusions = "age"
)

if (isFALSE(params$telolength)) {
test_correlation(
  df_results_biomarkers,
  grouping_variables = "predictors",
  exclusions = c(
    "age",
    "age, phenoage",
    "age, crp, hba1c, albumin_gL",
    "age, crp, hba1c, albumin_gL, phenoage"
  )
)

test_correlation(
  df_results_biomarkers,
  grouping_variables = c("age_group", "predictors"),
  exclusions = c(
    "age",
    "age, phenoage",
    "age, crp, hba1c, albumin_gL",
    "age, crp, hba1c, albumin_gL, phenoage"
  )
)

test_correlation(
  df_results_confounders,
  grouping_variables = "predictors",
  exclusions = c(
    "full"
  )
)

test_correlation(
  df_results_confounders,
  grouping_variables = c("age_group", "predictors"),
  exclusions = c(
    "full"
  )
)
}
```

All correlations trend positive. Overall bioage methods have positive
dependence, so we can use Hommel, for the other comparisons, correlations center
at 0 and are either mildly positively dependent or independent. There are some
negative correlations but practically all are between -0.25 and 0.25, so there
is probably no violation of the positive regression dependency in which it is
unlikely that an increase in one test statistic would lead to a decrease in
another. This also makes sense for bioage models. If phenoage works, kdm may
also work a little, because they are based on the same blood markers. Similarly,
if we get positive predictive value from one bioage method in one age group, it
probably positively affects predictive power in another age group. For
biomarkers, an increase in predictive power of one marker doesn't per se affect
another. If one of two closely-related biomarkers has higher predictive power,
it doesn't mean the other one won't. The only way to get a negative outcome is
to get overfitting, and I don't see in which scenario a higher predictive power
in one biomarker would increase the odds of overfitting for another biomarker,
so we can assume independence and use Hommel.

### Calculate p-values

Calculate t-statistic according to nadeau bengio, which is` t = m/(√(variance(1/n + n2/n1)))`. m is the difference between the sample mean and
hypothesized population mean. In this case, we've normalized it to age so the
null hypothesis (the hypothesized population mean) is 0, i.e., there is no
difference from age only. The variance we already calculated. n is equivalent to
the number of folds times the number of repeats. n2 is the number of
testing/validation data and n1 the number of training data. We then calculate
degrees of freedom by taking n - 1. Then we use pt (which automatically
calculates the critical value) using a one-tailed t-test of the upper tail. This
is because we're not interested in whether an age + bioage model is different
from age-only. We only want to know if its better. We then adjust p values with
Hommel or Holm error correction.

First, the question is, 'Does this bioage/biomarker add predictive value for
hearing loss in the population while controlling for chronological age?', which
we test with the overall comparisons. Only if it's significantly different do we
look at per age-group tests, also to control power. Another option is to ask
'Might bioage/biomarker only matter for specific age ranges, even if it doesn't
have an overall predictive effect?', in which case we could consider doing
pairwise on all, but we'd lose a lot of power and would have to use FDR and
accept a 5% false discovery rate to salvage it.

The CorrectR library has an implementation of the Nadeau-Bengio corrected t-test
(with support for repeated k-fold validations), but this makes the false
assumption that fold sizes are always equal, so there's a small error in
calculation of the sample mean and variance. The sample mean and variance should
be weighted by fold size. This gets relevant for age-group delta MSEs, because
for the 50-60 age group, some folds have a test size of 99 and others of 98.
That means that some folds weigh 1% more than another. Not correcting for that
could make the difference between significant or not signficant.

```{r t_test_models}
nb_corrected_repeated_t_test <- function(summary_data,
                                         exceptions = NULL,
                                         exception_include_exclude,
                                         adjust_method) {
  if (exception_include_exclude == "include") {
    summary_data <- summary_data %>%
      filter(predictors %in% exceptions)
  } else {
    summary_data <- summary_data %>%
      filter(!predictors %in% exceptions)
  }
  
  summary_data <- summary_data %>%
    mutate(
      t = mean_delta_mse / (sqrt(
        s2_delta_mse * (1 / n + sum_n_test / sum_n_train)
      )),
      df = n - 1,
      p = pt(t, df = df, lower.tail = FALSE),
      p_adj = p.adjust(p, method = adjust_method),
      signif = case_when(
        p_adj <= 0.001 ~ "***",
        p_adj <= 0.01 ~ "**",
        p_adj <= 0.05 ~ "*",
        TRUE ~ "ns"
      )
    ) %>%
    select(-t, -df, -p, -s2_delta_mse, -sum_n_test, -sum_n_train, -n)
  
  return(summary_data)
}

pwc_bioage_overall <- nb_corrected_repeated_t_test(
  df_summary_bioage_overall,
  exceptions = "age",
  exception_include_exclude = "exclude",
  adjust_method = "holm"
)

pwc_bioage <- nb_corrected_repeated_t_test(
  df_summary_bioage,
  exceptions = "age, phenoage",
  exception_include_exclude = "include",
  adjust_method = "holm"
)

if (isFALSE(params$telolength)) {
pwc_biomarkers_overall <- nb_corrected_repeated_t_test(
  df_summary_biomarkers_overall,
  exceptions = c(
    "age",
    "age, phenoage",
    "age, crp, hba1c, albumin_gL",
    "age, crp, hba1c, albumin_gL, phenoage"
  ),
  exception_include_exclude = "exclude",
  adjust_method = "holm"
)  %>%
  arrange(p_adj)

pwc_biomarkers <- nb_corrected_repeated_t_test(
  df_summary_biomarkers,
  exceptions = c("age",
    "age, phenoage",
    "age, crp, hba1c, albumin_gL",
    "age, crp, hba1c, albumin_gL, phenoage"),
  exception_include_exclude = "exclude",
  adjust_method = "holm"
)

pwc_confounders_overall <- nb_corrected_repeated_t_test(
  df_summary_confounders_overall,
  exceptions = c(
    "full"
  ),
  exception_include_exclude = "exclude",
  adjust_method = "holm"
)  %>%
  arrange(p_adj)

pwc_confounders <- nb_corrected_repeated_t_test(
  df_summary_confounders,
  exceptions = c("full", "full_with_phenoage"),
  exception_include_exclude = "exclude",
  adjust_method = "holm"
)
}

pwc_bioage_overall %>% kable("html", caption = "Overall Bioage Pairwise Comparisons") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position = "left",
    font_size = 12,
    fixed_thead = TRUE
  ) %>%
  scroll_box(width = "100%", height = "400px")

if (isFALSE(params$telolength)) {
pwc_biomarkers_overall %>% kable("html", caption = "Overall Biomarker Pairwise Comparisons") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position = "left",
    font_size = 12,
    fixed_thead = TRUE
  ) %>%
  scroll_box(width = "100%", height = "400px")
  
pwc_confounders_overall %>% kable("html", caption = "Overall Confounder Pairwise Comparisons") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position = "left",
    font_size = 12,
    fixed_thead = TRUE
  ) %>%
  scroll_box(width = "100%", height = "400px")
}
```

### Add p-values to summary dataframe

```{r add_p_vals}
import_p_values <- function(summary_data, pwc_data, per_age_group) {
  if (per_age_group) {
    grouping_variables <- c("predictors", "age_group")
  } else {
    grouping_variables <- c("predictors")
  }
  
  updated_summary_data <- summary_data %>%
    left_join(pwc_data %>%
                select(all_of(grouping_variables), p_adj, signif), by = c(grouping_variables)) %>%
    mutate(
      p_adj = as.character(p_adj),
      p_adj = case_when(signif == "ns" ~ "ns", is.na(p_adj) ~ "ns", TRUE ~ p_adj),
      signif = case_when(is.na(signif) ~ "ns", TRUE ~ signif)
    )
  
  return(updated_summary_data)
}

df_summary_bioage <- import_p_values(df_summary_bioage, pwc_bioage, per_age_group = TRUE)
df_summary_bioage_overall <- import_p_values(df_summary_bioage_overall,
                                             pwc_bioage_overall,
                                             per_age_group = FALSE)

if (isFALSE(params$telolength)) {
df_summary_biomarkers <- import_p_values(df_summary_biomarkers, pwc_biomarkers, per_age_group = TRUE)
df_summary_biomarkers_overall <- import_p_values(df_summary_biomarkers_overall,
                                                 pwc_biomarkers_overall,
                                                 per_age_group = FALSE)

df_summary_confounders <- import_p_values(df_summary_confounders, pwc_confounders, per_age_group = TRUE)
df_summary_confounders_overall <- import_p_values(df_summary_confounders_overall,
                                                 pwc_confounders_overall,
                                                 per_age_group = FALSE)
}
```

**It doesn't matter whether we use holm or hommel, so perhaps we should use hommel.**

### Add raw Ns to dataframe

```{r add_n}
add_n_age_groups <- function(data, summary_data, per_age_group) {
  if (per_age_group) {
    summary_data <- summary_data %>%
      rename("n_folds_repeats" = n) %>%
      left_join(data %>% select(age_group) %>%
                  group_by(age_group) %>%
                  summarize(n = n()),
                by = "age_group")
  } else {
    summary_data <- summary_data %>%
      rename("n_folds_repeats" = n) %>%
      mutate(n = nrow(data))
  }
  
  return(summary_data)
}

df_summary_bioage <- add_n_age_groups(df, df_summary_bioage, per_age_group = TRUE)
df_summary_bioage_overall <- add_n_age_groups(df, df_summary_bioage_overall, per_age_group = FALSE)

if (isFALSE(params$telolength)) {
df_summary_biomarkers <- add_n_age_groups(df, df_summary_biomarkers, per_age_group = TRUE)
df_summary_biomarkers_overall <- add_n_age_groups(df, df_summary_biomarkers_overall, per_age_group = FALSE)

df_summary_confounders <- add_n_age_groups(df, df_summary_confounders, per_age_group = TRUE)
df_summary_confounders_overall <- add_n_age_groups(df, df_summary_confounders_overall, per_age_group = FALSE)
}
```

## Confounder model comparison

```{r}
if (isFALSE(params$telolength)) {
anova(output_confounders$models$full, output_confounders$models$full_with_phenoage)
anova(output_confounders$models$full, output_confounders$models$no_gender)
anova(output_confounders$models$full, output_confounders$models$no_smoking)
anova(output_confounders$models$full, output_confounders$models$no_race)
anova(output_confounders$models$full, output_confounders$models$no_education)
anova(output_confounders$models$full, output_confounders$models$`no_gender, race`)
}
```