# Visualization

## Tables

#### Descriptives table

#### Format table

Round R² and delta RMSE, rename variables, set p values to bold and to " <
0.001" when very small.

```{r format_table}
format_table <- function(summary_data, per_age_group) {
  table <- summary_data %>%
    mutate(
      r2 = round(r2, 3),
      delta_rmse = round(delta_rmse, 2)
    ) %>%
    rename(
      "Predictors" = predictors,
      "∆RMSE (vs age, dB HL)" = delta_rmse,
      "R²" = r2,
      "p (vs age)" = p_adj
    ) %>%
    mutate(
      `p (vs age)` = case_when(
        signif == "***" ~
          cell_spec("< 0.001", "html", bold = TRUE),
        suppressWarnings(!is.na(as.numeric(`p (vs age)`)) &
          as.numeric(`p (vs age)`) < 0.05) &
          `p (vs age)` != "< 0.001" ~
          cell_spec(suppressWarnings(round(
            as.numeric(`p (vs age)`), 3
          )), "html", bold = TRUE),
        TRUE ~ `p (vs age)`
      ),
      `p (vs age)` = as.character(`p (vs age)`),
      `p (vs age)` = case_when(Predictors == "age" ~ "", TRUE ~ `p (vs age)`),
      `∆RMSE (vs age, dB HL)` = as.character(`∆RMSE (vs age, dB HL)`),
      `∆RMSE (vs age, dB HL)` = case_when(
        Predictors == "age" ~ "",
        TRUE ~ `∆RMSE (vs age, dB HL)`
      )
    )

  if (isTRUE(per_age_group)) {
    table <- table %>%
      rename("Age" = age_group) %>%
      select(Predictors, Age, n, `R²`, `∆RMSE (vs age, dB HL)`, `p (vs age)`)
  } else {
    table <- table %>%
      select(Predictors, n, `R²`, `∆RMSE (vs age, dB HL)`, `p (vs age)`)
  }

  table <- table %>%
    rename(
      "p (vs full model)" = `p (vs age)`,
      "∆RMSE (vs full model, dB HL)" = `∆RMSE (vs age, dB HL)`,
    )

  return(table)
}

table_bioage <- format_table(df_summary_bioage, per_age_group = TRUE)
table_bioage_overall <- format_table(df_summary_bioage_overall, per_age_group = FALSE)

if (isFALSE(params$telolength)) {
  table_biomarkers <- format_table(df_summary_biomarkers, per_age_group = TRUE)
  table_biomarkers_overall <- format_table(df_summary_biomarkers_overall, per_age_group = FALSE)

  table_confounders <- format_table(df_summary_confounders, per_age_group = TRUE)
  table_confounders_overall <- format_table(df_summary_confounders_overall, per_age_group = FALSE)
}
```

#### Order, Filter, and Rename Tables

```{r order_table}
rename_bioage_table <- function(bioage_table) {
  bioage_table <- bioage_table %>%
    mutate(
      Predictors = case_when(
        Predictors == "age" ~ "Age",
        Predictors == "age, hd_log" ~ "Age + HD",
        Predictors == "age, kdm" ~ "Age + KDM",
        Predictors == "age, phenoage" ~ "Age + PhenoAge",
        isTRUE(params$telolength) & Predictors == "age, telolength" ~ "Age + Telolength",
        TRUE ~ Predictors
      ),
      Predictors = factor(
        Predictors,
        levels = if (isTRUE(params$telolength)) {
          c("Age", "Age + HD", "Age + KDM", "Age + PhenoAge", "Age + Telolength")
        } else {
          c("Age", "Age + HD", "Age + KDM", "Age + PhenoAge")
        }
      )
    ) %>%
    arrange(Predictors)

  return(bioage_table)
}

rename_biomarker_table <- function(biomarker_table) {
  biomarker_table <- biomarker_table %>%
    filter(!(
      Predictors %in% c(
        "age",
        "age, phenoage",
        "age, crp, hba1c, albumin_gL",
        "age, crp, hba1c, albumin_gL, phenoage"
      )
    )) %>%
    mutate(
      Predictors = case_when(
        Predictors == "age, rdw" ~ "Age + RDW",
        Predictors == "age, alp" ~ "Age + ALP",
        Predictors == "age, crp" ~ "Age + CRP",
        Predictors == "age, totchol" ~ "Age + TC",
        Predictors == "age, albumin_gL" ~ "Age + ALB",
        Predictors == "age, hba1c" ~ "Age + HbA1c",
        Predictors == "age, glucose" ~ "Age + Glc",
        Predictors == "age, sbp" ~ "Age + SBP",
        Predictors == "age, bun" ~ "Age + BUN",
        Predictors == "age, mcv" ~ "Age + MCV",
        Predictors == "age, wbc" ~ "Age + WBC",
        Predictors == "age, lymph" ~ "Age + Lymphs",
        Predictors == "age, uap" ~ "Age + UA",
        Predictors == "age, creat" ~ "Age + Creat"
      ),
      Predictors = factor(Predictors, levels = c(
        "Age + SBP",
        "Age + BUN",
        "Age + MCV",
        "Age + Lymphs",
        "Age + RDW",
        "Age + TC",
        "Age + Creat",
        "Age + UA",
        "Age + Glc",
        "Age + WBC",
        "Age + ALP",
        "Age + ALB",
        "Age + HbA1c",
        "Age + CRP"
      ))
    ) %>%
    arrange(Predictors)

  return(biomarker_table)
}

rename_confounder_table <- function(confounders_table) {
  confounders_table <- confounders_table %>%
    filter(!(
      Predictors %in% c(
        "full"
      )
    )) %>%
    mutate(
      Predictors = case_when(
        Predictors == "full_with_phenoage" ~ "Phenoage",
        Predictors == "no_gender" ~ "Gender",
        Predictors == "no_smoking" ~ "Smoking",
        Predictors == "no_race" ~ "Race",
        Predictors == "no_education" ~ "Education",
        Predictors == "no_gender, race" ~ "Gender * Race"
      ),
      Predictors = factor(Predictors, levels = c(
        "Phenoage",
        "Gender",
        "Smoking",
        "Race",
        "Education",
        "Gender * Race"
      ))
    ) %>%
    arrange(Predictors)

  return(confounders_table)
}

table_bioage <- rename_bioage_table(table_bioage)
table_bioage_overall <- rename_bioage_table(table_bioage_overall)

if (isFALSE(params$telolength)) {
  table_biomarkers <- rename_biomarker_table(table_biomarkers)
  table_biomarkers_overall <- rename_biomarker_table(table_biomarkers_overall)

  table_confounders <- rename_confounder_table(table_confounders)
  table_confounders_overall <- rename_confounder_table(table_confounders_overall)
}
```

#### Render table

```{r store_tables}
render_and_save_table <- function(table, path) {
  html_table <- table %>%
    kable(
      format = "html",
      align = c("l", "l", "r", "r", "r", "r", "r", "r", "r"),
      escape = FALSE
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE,
      position = "center"
    ) %>%
    collapse_rows(columns = 1, valign = "top")

  if (isTRUE(params$save_figures)) {
    save_kable(html_table, path)
  }

  return(html_table)
}

render_and_save_table(table_bioage, path = here("Output", "Figures", params$hearing_var, "table_bioage.html"))
render_and_save_table(table_bioage_overall, path = here("Output", "Figures", params$hearing_var, "table_bioage_overall.html"))

if (isFALSE(params$telolength)) {
  render_and_save_table(table_biomarkers, path = here("Output", "Figures", params$hearing_var, "table_biomarkers.html")) %>%
    scroll_box(width = "100%", height = "400px")
  render_and_save_table(table_biomarkers_overall, path = here("Output", "Figures", params$hearing_var, "table_biomarkers_overall.html"))

  render_and_save_table(table_confounders, path = here("Output", "Figures", params$hearing_var, "table_confounders.html")) %>%
    scroll_box(width = "100%", height = "400px")
  render_and_save_table(table_confounders_overall, path = here("Output", "Figures", params$hearing_var, "table_confounders_overall.html"))
}
```

### Confounders table

#### Build table and recode values

```{r build_conf_table}
build_confounder_table <- function(data, y = params$hearing_var) {
  variables <- c("Gender", "Smoking", "Race", "Education")

  data <- data %>%
    mutate(
      gender = dplyr::recode(gender, "1" = "Male", "2" = "Female"),
      smoking = dplyr::recode(smoking, "1" = "Smoker", "2" = "Non-smoker"),
      race = dplyr::recode(
        race,
        "1" = "Hispanic",
        "3" = "White",
        "4" = "Black",
        "5" = "Other"
      ),
      education = dplyr::recode(
        education,
        "1" = "< 9th grade",
        "2" = "9-11th grade",
        "3" = "High school",
        "4" = "Some college or AA degree",
        "5" = "College graduate or higher"
      )
    ) %>%
    rename(
      "Smoking" = smoking,
      "Gender" = gender,
      "Race" = race,
      "Education" = education
    )

  bind_rows(lapply(variables, function(variable) {
    data %>%
      mutate(level = .data[[variable]]) %>%
      group_by(age_group, level) %>%
      summarise(
        mean = mean(.data[[y]], na.rm = TRUE),
        sd = sd(.data[[y]], na.rm = TRUE),
        .groups = "drop"
      ) %>%
      mutate(
        variable = factor(variable, levels = c(
          "Gender", "Smoking", "Race", "Education"
        )),
        .before = 1
      )
  }))
}

if (isFALSE(params$telolength)) {
  table_confounders_descriptive <- build_confounder_table(df)
}
```

#### Pivot table and merge mean and standard deviation

```{r pivot_conf_table}
if (isFALSE(params$telolength)) {
  table_confounders_descriptive <- table_confounders_descriptive %>%
    mutate(stat = sprintf("%.2f (%.2f)", mean, sd)) %>%
    select(variable, level, age_group, stat) %>%
    pivot_wider(names_from = age_group, values_from = stat)
}
```

#### Format table

```{r format_Conf_table}
if (isFALSE(params$telolength)) {
  table_confounders_descriptive <- table_confounders_descriptive %>%
    group_by(variable) %>%
    mutate(variable = if_else(row_number() == 1L, variable, "")) %>%
    ungroup() %>%
    rename(" " = variable, "  " = level)
}
```

#### Save table

```{r store_conf_table}
if (isFALSE(params$telolength)) {
  render_and_save_table(
    table_confounders_descriptive,
    path = here(
      "Output",
      "Figures",
      params$hearing_var,
      "table_confounders_descriptives.html"
    )
  )
}
```

## Figures

### Set plot config

#### Set plot styling variables

We store a separate configuration depending on output type of figure for easy switching between output types. Only the currently selected output type variables in params + global variables for plotting we may need later is returned. These variables are used to construct ggplot protoobjects.

```{r}
get_plot_config <- function(params) {
  p_cfg_vars <- list(
    global = list(
    axis_unit_color = "#4D4D4D" # used later when constructing axes from HTML
    ),
    poster = list(),
    paper = list(
      figure_dir = here(
        "Output",
        "Figures",
        if (isTRUE(params$telolength)) {
          paste0(params$hearing_var, "_telolength")
        } else {
          params$hearing_var
        }
      ),
      text_size_primary = 8,
      text_size_secondary = 9,
      signif_size = 14
    )
  )
  
  c(p_cfg_vars$global, p_cfg_vars[[params$plot_output_type]])
}
```

#### Store general plot theme

We first store the theme preset (theme_few), then theme overrides, then other theme items like guides. Upon initial assignment of the theme settings, we read directly from get_plot_config(). The plot output type variables set in the previous chunk are also appended, so that they can be accessed either with p_cfg or with get_plot_config(params). All items in the plot_theme list must be ggplot proto objects. Repeated get_plot_config() calls are used to avoid creating another unnecessary variable. Setting the theme with repeated calls takes < 200ms, so clarity of intent with only declaring one list of p_cfg is preferred

```{r}
p_cfg <-
  c(list(plot_theme = list(
    theme_few(),
    theme(
      axis.text.x = element_markdown("sans", size = get_plot_config(params)$text_size_primary),
      axis.text.y = element_markdown("sans", size = get_plot_config(params)$text_size_primary),
      strip.text.x = element_markdown(
        "sans",
        size = get_plot_config(params)$text_size_secondary,
        lineheight = 1.1
      ),
      strip.text.y = element_markdown(
        "sans",
        size = get_plot_config(params)$text_size_secondary,
        lineheight = 1.1
      ),
      axis.title.x = element_markdown(
        "sans",
        size = get_plot_config(params)$text_size_secondary,
        lineheight = 1.1
      ),
      axis.title.y = element_markdown(
        "sans",
        size = get_plot_config(params)$text_size_secondary,
        lineheight = 1.1
      ),
      legend.title = element_blank(),
      legend.background = element_rect(fill = "transparent", color = NA),
      legend.box.background = element_rect(fill = "transparent", color = NA),
      legend.key = element_rect(fill = "transparent"),
      legend.position = "inside",
      legend.justification.inside = c(0, 1),
      legend.key.spacing.y = unit(1, "pt"),
      panel.border = element_part_rect(
        colour = "grey30",
        side = "lb",
        linewidth = 0.5
      ),
      plot.margin = margin(1, 1, 0, 1, unit = "pt")
    ),
    guides(
      colour = guide_legend(
        override.aes = list(size = 1.25),
        keywidth = unit(0.3, "cm"),
        keyheight = unit(0.3, "cm"),
        label.theme = element_markdown(
          "sans",
          size = get_plot_config(params)$text_size_secondary,
          margin = margin(l = 3, t = 1.5, unit = "pt")
        )
      )
    )
  )), get_plot_config(params))
```

#### Set axis scales

```{r}
p_cfg <- modifyList(
  p_cfg,
  list(
    x = list(
      age = set_axis(p_cfg, "x", "continuous", breaks = seq(20, 80, 10), limits = c(20, 82), label = "Age", unit = "Years"),
      age_group = set_axis(p_cfg, "x", "discrete", limits = c(1.4, 5.6), label = "Age", unit = "Years"),
      age_accel = set_axis(p_cfg, "x", "continuous", limits = c(-20, 20), label = "Age Acceleration", unit = "PhenoAge – Age (Years)")
    ),
    y = list(
      age_group = set_axis(p_cfg, "y", "discrete", limits = c(1, 7), label = "Age", unit = "Years"),
      hl = set_axis(p_cfg, "y", "continuous", breaks = seq(0, 90, 30), limits = c(-10, 100), label = paste0(if (params$hearing_var == "FI_high") "High" else "Low", " Fletcher Index"), unit = "dB HL"),
      hl_rate = set_axis(p_cfg, "y", "continuous", breaks = seq(0, 2, 1), limits = c(-0.5, 2.6), label = "Rate of HL", unit = "dB HL/Year"),
      hl_diff = set_axis(p_cfg, "y", "continuous", breaks = seq(0, 2, 1), limits = c(-0.5, 2.6), label = "Deviation Rate of HL", unit = "dB HL/Year"),
      dmse_bioage = set_axis(p_cfg, "y", "continuous", breaks = seq(0, 8, 2), limits = c(-1, 8.5), label = "Additional Predictive Capacity", unit = "Δ MSE (dB HL)<sup>2</sup>"),
      drmse_bioage = set_axis(p_cfg, "y", "continuous", breaks = seq(0, 0.15, 0.05), limits = c(-0.05, 0.15), label = "Additional Predictive Capacity", unit = "Δ RMSE (dB HL)"),
      dmse_biomark = set_axis(p_cfg, "y", "continuous", breaks = seq(0, 2, 1), limits = c(-0.25, 1.5), label = "Additional Predictive Capacity", unit = "Δ MSE (dB HL)<sup>2</sup>"),
      dmse_conf = set_axis(p_cfg, "y", "continuous", breaks = seq(0, 40, 5), limits = c(-1, 40), label = "Additional Predictive Capacity", unit = "Δ MSE (dB HL)<sup>2</sup>"),
      dmse_best_biomark = set_axis(p_cfg, "y", "continuous", breaks = c(0, 2), limits = c(-1.5, 4), label = "Additional Predictive Capacity", unit = "Δ MSE (dB HL)<sup>2</sup>"),
      add_pred_conf = set_axis(p_cfg, "y", "continuous", breaks = seq(0, 14, 2), limits = c(-1, 15), label = "Additional Predictive Capacity", unit = "% Change MSE")
    )
  )
)
```

#### Set colors

```{r}
p_cfg <- modifyList(
  p_cfg,
  list(
    colors = list(
      bioage = set_colors(
        colors = c(
          "age" = "#000000",
          "age, hd_log" = "#256DAB",
          "age, kdm" = "#059748",
          "age, phenoage" = "#CB2027",
          "age, telolength" = if (isTRUE(params$telolength)) "#7B3A96" else NULL
        ),
        fill_colors = c(
          "age" = "#8C8C8C",
          "age, hd_log" = "#88BDE6",
          "age, kdm" = "#90CD97",
          "age, phenoage" = "#F07E6E",
          "age, telolength" = if (isTRUE(params$telolength)) "#B276B2" else NULL
        ),
        labels = c("age" = "Age", "age, hd_log" = "Age + HD", "age, kdm" = "Age + KDM", "age, phenoage" = "Age + PhenoAge", "age, telolength" = if (isTRUE(params$telolength)) "Age + Telolength" else NULL)
      ),
      confounders = set_colors(
        colors = c(
          "full" = "#000000",
          "no_gender" = "#059748",
          "no_smoking" = "#9D722A",
          "no_race" = "#E5126F",
          "no_education" = "#DF5C24",
          "no_gender, race" = "#7B3A96",
          "full_with_phenoage" = "#256DAB"
        ),
        fill_colors = c(
          "full" = "#8C8C8C",
          "no_gender" = "#90CD97",
          "no_smoking" = "#BFA554",
          "no_race" = "#F6AAC9",
          "no_education" = "#FBB258",
          "no_gender, race" = "#BC99C7",
          "full_with_phenoage" = "#88BDE6"
        ),
        labels = c(
          "full" = "Full",
          "no_gender" = "Gender",
          "no_smoking" = "Smoking",
          "no_race" = "Race",
          "no_education" = "Education",
          "no_gender, race" = "Gender*Race",
          "full_with_phenoage" = "Phenoage"
        )
      ),
      phenoage = set_colors(
        colors = c(
          "age" = "#000000", "age, phenoage" = "#CB2027"
        ),
        fill_colors = c(
          "age" = "#8C8C8C", "age, phenoage" = "#F07E6E"
        ),
        labels = c(
          "age" = "Age", "age, phenoage" = "Age + PhenoAge"
        )
      ),
      gender = set_colors(
        colors = c(
          "1" = "#2986cc", "2" = "#c90076"
        ),
        fill_colors = c(
          "1" = "#6fa8dc", "2" = "#d5a6bd"
        ),
        labels = c("1" = "Male", "2" = "Female")
      ),
      race = set_colors(
        colors = c(
          "1" = "#E5126F",
          "3" = "#DF5C24",
          "4" = "#7B3A96",
          "5" = "#256DAB"
        ),
        fill_colors = c(
          "1" = "#F6AAC9",
          "3" = "#FBB258",
          "4" = "#BC99C7",
          "5" = "#88BDE6"
        ),
        labels = c("1" = "Hispanic", "3" = "White", "4" = "Black", "5" = "Other")
      ),
      biomarker_diff = set_colors(
        colors = c(
          "age, crp, hba1c, albumin_gL, phenoage" = "#CB2027",
          "age, crp, hba1c, albumin_gL" = "#059748",
          "age, phenoage" = "grey65"
        ),
        fill_colors = c(
          "age, crp, hba1c, albumin_gL, phenoage" = "#F07E6E",
          "age, crp, hba1c, albumin_gL" = "#90CD97",
          "age, phenoage" = "grey80"
        ),
        labels = c(
          "age, crp, hba1c, albumin_gL, phenoage" = "Age + CRP + HbA1c + ALB + PhenoAge",
          "age, crp, hba1c, albumin_gL" = "Age + CRP + HbA1c + ALB",
          "age, phenoage" = "Age + PhenoAge"
        ),
        breaks = c(
          "age, crp, hba1c, albumin_gL, phenoage",
          "age, crp, hba1c, albumin_gL",
          "age, phenoage"
        )
      ),
      best_biomarkers = set_colors(
        colors = c(
          "age, crp" = "#256DAB",
          "age, hba1c" = "#E5126F",
          "age, albumin_gL" = "#DF5C24"
        ),
        fill_colors = c(
          "age, crp" = "#88BDE6",
          "age, hba1c" = "#F6AAC9",
          "age, albumin_gL" = "#FBB258"
        )
      )
    )
  )
)
```

#### Set geom style

Note that using update_geom_defaults() is global and requires reset_geom_defaults() to be reset. update_geom_defaults() is only set once when switching to a different plot type and subsequent plots with the same styling will not have repeated update_geom_defaults() calls to avoid overhead and clutter. Plots must therefore be run in order.

```{r}
p_cfg <- modifyList(
  p_cfg,
  list(
    geom_args = list(
      ribbon = list(
        alpha = 0.15
      ),
      highlight_rect = list(
        alpha = 0.125,
        fill = "#ff0000"
      ),
      hline = list(
        linetype = "22",
        color = "grey20",
        linewidth = 0.2
      ),
      age_rate_point = list(
        alpha = 0.1,
        size = 0.5,
        shape = 1,
        stroke = 0.25,
        color = "grey30"
      ),
      age_rate_line = list(
        linewidth = 0.6,
        alpha = 1
      ),
      mse_point = list(
        size = 1.25,
        alpha = 1,
        shape = 16,
        stroke = 0.5
      ),
      mse_line = list(
        linewidth = 0.7,
        alpha = 1
      ),
      biomarker_linerange = list(
        linewidth = 0.75
      ),
      biomarker_point = list(
        size = 2.25,
        alpha = 1,
        shape = 16,
        stroke = 0.5
      )
    )
  )
)

update_geom_defaults("ribbon", p_cfg$geom_args$ribbon)
update_geom_defaults("rect", p_cfg$geom_args$highlight_rect)
update_geom_defaults("hline", p_cfg$geom_args$hline)
```


### Rate of aging panel

#### Main panel (paper)

```{r plot_aging, warning=FALSE}
update_geom_defaults("point", p_cfg$geom_args$age_rate_point)
update_geom_defaults("line", p_cfg$geom_args$age_rate_line)

p0 <- ggplot() +
  geom_point(
    data = df,
    aes(age, .data[[params$hearing_var]])
  ) +
  geom_line(
    data = output_age_rate$predictions$linear,
    aes(x = age, y = estimate)
  ) +
  geom_ribbon(
    data = output_age_rate$predictions$linear,
    aes(x = age, ymin = ci_lower, ymax = ci_upper)
  ) +
  construct_axes(p_cfg$x$age, p_cfg$y$hl) +
  p_cfg$plot_theme +
  theme(
    plot.margin = margin(1, 0, 0, 1, unit = "pt"),
    axis.title.x = element_blank(),
    geom = element_geom()
  )

p1 <- ggplot() +
  geom_point(
    data = df,
    aes(age, .data[[params$hearing_var]])
  ) +
  geom_line(
    data = output_age_rate$predictions$quadratic,
    aes(x = age, y = estimate)
  ) +
  geom_ribbon(
    data = output_age_rate$predictions$quadratic,
    aes(x = age, ymin = ci_lower, ymax = ci_upper)
  ) +
  construct_axes(p_cfg$x$age, p_cfg$y$hl) +
  p_cfg$plot_theme +
  theme(
    plot.margin = margin(1, 0, 0, 1, unit = "pt"),
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
  )

p2 <- ggplot() +
  geom_rect(
    data = pwc_age_rate_boundaries,
    aes(
      xmin = x_min,
      xmax = x_max,
      ymin = -Inf,
      ymax = Inf
    )
  ) +
  geom_point(
    data = df,
    aes(age, .data[[params$hearing_var]])
  ) +
  geom_ribbon(
    data = output_age_rate$predictions$smooth,
    aes(
      x = age,
      ymin = ci_lower + coef(output_age_rate$models$smooth)["(Intercept)"],
      ymax = ci_upper + coef(output_age_rate$models$smooth)["(Intercept)"]
    )
  ) +
  geom_line(
    data = output_age_rate$predictions$smooth,
    aes(
      x = age,
      y = estimate + coef(output_age_rate$models$smooth)["(Intercept)"]
    )
  ) +
  construct_axes(p_cfg$x$age, p_cfg$y$hl) +
  p_cfg$plot_theme +
  theme(
    plot.margin = margin(1, 1, 0, 0, unit = "pt"),
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
  )

p3 <- ggplot() +
  geom_hline(yintercept = 0) +
  geom_line(
    data = output_age_rate$predictions$linear_derivative,
    aes(x = age, y = derivative)
  ) +
  geom_ribbon(
    data = output_age_rate$predictions$linear_derivative,
    aes(x = age, ymin = ci_lower, ymax = ci_upper)
  ) +
  p_cfg$plot_theme +
  construct_axes(p_cfg$x$age, p_cfg$y$hl_rate) +
  theme(plot.margin = margin(0, 0, 0, 1, unit = "pt"))

p4 <- ggplot() +
  geom_hline(yintercept = 0) +
  geom_line(
    data = output_age_rate$predictions$quadratic_derivative,
    aes(x = age, y = derivative)
  ) +
  geom_ribbon(
    data = output_age_rate$predictions$quadratic_derivative,
    aes(x = age, ymin = ci_lower, ymax = ci_upper)
  ) +
  construct_axes(p_cfg$x$age, p_cfg$y$hl_rate) +
  p_cfg$plot_theme +
  theme(
    plot.margin = margin(0, 0, 0, 1, unit = "pt"),
    axis.title.y = element_blank()
  )

p5 <- ggplot() +
  geom_rect(
    data = pwc_age_rate_boundaries,
    aes(
      xmin = x_min,
      xmax = x_max,
      ymin = -Inf,
      ymax = Inf
    )
  ) +
  geom_hline(yintercept = 0) +
  geom_ribbon(
    data = output_age_rate$predictions$derivative_smooth,
    aes(x = age, ymin = ci_lower, ymax = ci_upper)
  ) +
  geom_line(
    data = output_age_rate$predictions$derivative_smooth,
    aes(x = age, y = derivative)
  ) +
  construct_axes(p_cfg$x$age, p_cfg$y$hl_rate) +
  p_cfg$plot_theme +
  theme(
    plot.margin = margin(0, 1, 0, 0, unit = "pt"),
    axis.title.y = element_blank()
  )

p6 <- ggplot() +
  geom_rect(
    data = pwc_age_rate_boundaries,
    aes(
      xmin = x_min,
      xmax = x_max,
      ymin = -Inf,
      ymax = Inf
    )
  ) +
  geom_hline(yintercept = 0) +
  geom_ribbon(
    data = output_age_rate$predictions$derivative_smooth,
    aes(x = age, ymin = ci_lower_adj, ymax = ci_upper_adj)
  ) +
  geom_line(
    data = output_age_rate$predictions$derivative_smooth,
    aes(x = age, y = derivative_adj)
  ) +
  construct_axes(p_cfg$x$age, p_cfg$y$hl_diff) +
  p_cfg$plot_theme +
  theme(plot.margin = margin(0, 0, 1, 0, unit = "pt"))


p <- align_plots(p0, p1, p2, p3, p4, p5, p6, align = "hv")

p <- plot_grid(
  plot_grid(
    plot_grid(
      p[[1]],
      NULL,
      p[[2]],
      NULL,
      p[[3]],
      labels = c("A", "", "B", "", "C"),
      nrow = 1,
      rel_widths = c(1, -0.01, 1, -0.01, 1),
      label_x = 0.04,
      label_y = 1.02
    ),
    NULL,
    plot_grid(
      p[[4]],
      NULL,
      p[[5]],
      NULL,
      p[[6]],
      labels = c("D", "", "E", "", "F"),
      nrow = 1,
      rel_widths = c(1, -0.01, 1, -0.01, 1),
      label_x = 0.04,
      label_y = 1.02
    ),
    ncol = 1,
    rel_heights = c(1, -0.15, 1)
  ),
  NULL,
  plot_grid(
    NULL,
    p[[7]],
    NULL,
    ncol = 1,
    rel_heights = c(0.23, 0.54, 0.23),
    labels = c("", "G", ""),
    label_x = 0.04,
    label_y = 1.1
  ),
  ncol = 3,
  rel_widths = c(3, 0.1, 1.005)
)

sameplot(
  p,
  file = here(p_cfg$figure_dir, "age_rate_panel.png"),
  width = params$double_column,
  height = 3,
  res = 600,
  save = params$save_figures
)
```

#### Additional plots

##### Simplified Rate of Aging

```{r plot_age_rate_simple}
if (isFALSE(params$telolength)) {
  p <- align_plots(p0, p1, p2, p3, p4, p5, align = "hv")

  p <-
    plot_grid(
      plot_grid(
        p[[1]],
        NULL,
        p[[2]],
        NULL,
        p[[3]],
        labels = c("A", "", "B", "", "C"),
        nrow = 1,
        rel_widths = c(1, -0.01, 1, -0.01, 1),
        label_x = 0.04,
        label_y = 1.02
      ),
      NULL,
      plot_grid(
        p[[4]],
        NULL,
        p[[5]],
        NULL,
        p[[6]],
        labels = c("D", "", "E", "", "F"),
        nrow = 1,
        rel_widths = c(1, -0.01, 1, -0.01, 1),
        label_x = 0.04,
        label_y = 1.02
      ),
      ncol = 1,
      rel_heights = c(1, -0.15, 1)
    )

  sameplot(
    p,
    file = here(
      p_cfg$figure_dir,
      "Extras",
      "simplified_age_rate_panel.png"
    ),
    width = params$double_column * 1.8,
    height = 3 * 1.8,
    scaling = 1.8,
    res = 600,
    save = params$save_figures
  )
}
```

##### Blank rate of aging

```{r plot_blank_age_rate}
if (isFALSE(params$telolength)) {
  p0 <- ggplot() +
    geom_point(
      data = df,
      aes(age, .data[[params$hearing_var]])
    ) +
    construct_axes(p_cfg$x$age, p_cfg$y$hl) +
    p_cfg$plot_theme +
    theme(
      plot.margin = margin(1, 0, 0, 1, unit = "pt"),
      axis.title.x = element_blank()
    )

  p1 <- ggplot() +
    geom_point(
      data = df,
      aes(age, .data[[params$hearing_var]])
    ) +
    construct_axes(p_cfg$x$age, p_cfg$y$hl) +
    p_cfg$plot_theme +
    theme(
      plot.margin = margin(1, 0, 0, 1, unit = "pt"),
      axis.title.y = element_blank(),
      axis.title.x = element_blank()
    )

  p2 <- ggplot() +
    geom_point(
      data = df,
      aes(age, .data[[params$hearing_var]])
    ) +
    construct_axes(p_cfg$x$age, p_cfg$y$hl) +
    p_cfg$plot_theme +
    theme(
      plot.margin = margin(1, 1, 0, 0, unit = "pt"),
      axis.title.y = element_blank(),
      axis.title.x = element_blank()
    )

  p3 <- ggplot() +
    geom_hline(
      yintercept = 0
    ) +
    geom_line(
      data = output_age_rate$predictions$linear_derivative,
      aes(x = age, y = derivative),
      alpha = 0
    ) +
    construct_axes(p_cfg$x$age, p_cfg$y$hl_rate) +
    p_cfg$plot_theme +
    theme(plot.margin = margin(0, 0, 0, 1, unit = "pt"))

  p4 <- ggplot() +
    geom_hline(
      yintercept = 0
    ) +
    geom_line(
      data = output_age_rate$predictions$quadratic_derivative,
      aes(x = age, y = derivative),
      alpha = 0
    ) +
    construct_axes(p_cfg$x$age, p_cfg$y$hl_rate) +
    p_cfg$plot_theme +
    theme(
      plot.margin = margin(0, 0, 0, 1, unit = "pt"),
      axis.title.y = element_blank()
    )

  p5 <- ggplot() +
    geom_hline(
      yintercept = 0
    ) +
    geom_line(
      data = output_age_rate$predictions$derivative_smooth,
      aes(x = age, y = derivative),
      alpha = 0
    ) +
    construct_axes(p_cfg$x$age, p_cfg$y$hl_rate) +
    p_cfg$plot_theme +
    theme(
      plot.margin = margin(0, 1, 0, 0, unit = "pt"),
      axis.title.y = element_blank()
    )

  p <- align_plots(p0, p1, p2, p3, p4, p5, align = "hv")

  p <-
    plot_grid(
      plot_grid(
        p[[1]],
        NULL,
        p[[2]],
        NULL,
        p[[3]],
        labels = c("A", "", "B", "", "C"),
        nrow = 1,
        rel_widths = c(1, -0.01, 1, -0.01, 1),
        label_x = 0.04,
        label_y = 1.02
      ),
      NULL,
      plot_grid(
        p[[4]],
        NULL,
        p[[5]],
        NULL,
        p[[6]],
        labels = c("D", "", "E", "", "F"),
        nrow = 1,
        rel_widths = c(1, -0.01, 1, -0.01, 1),
        label_x = 0.04,
        label_y = 1.02
      ),
      ncol = 1,
      rel_heights = c(1, -0.15, 1)
    )

  sameplot(
    p,
    file = here(p_cfg$figure_dir, "Extras", "blank_age_rate_panel.png"),
    width = params$double_column * 1.8,
    height = 3 * 1.8,
    scaling = 1.8,
    res = 600,
    save = params$save_figures
  )
}
```

##### Rate of aging gender

###### Calculate gender GAM

```{r calc_gender_GAM}
if (isFALSE(params$telolength)) {
  gam_gender <- bam(
    as.formula(
      paste0(
        params$hearing_var,
        "~ s(age, k = 10) + s(gender, age, bs = 'sz', k = 10)"
      )
    ),
    data = df,
    method = "REML",
    family = gaussian()
  )

  gender_predictions <- as.data.frame(confint(
    emmeans(
      gam_gender,
      as.formula(paste("~ age + gender")),
      at = setNames(list(seq(
        min(df[["age"]]), max(df[["age"]]),
        by = 1
      )), "age"),
      type = "response",
      data = df
    ),
    adjust = "mvt"
  )) %>%
    rename(
      "estimate" = emmean,
      "ci_lower" = lower.CL,
      "ci_upper" = upper.CL
    )

  gender_predictions_der <- data.frame(confint(
    emtrends(
      gam_gender,
      as.formula(paste("~ age + gender")),
      var = "age",
      at = setNames(list(seq(
        min(df[["age"]]), max(df[["age"]]),
        by = 1
      )), "age"),
      data = df
    ),
    adjust = "mvt"
  )) %>%
    rename(
      "derivative" = !!paste0("age", ".trend"),
      "ci_lower" = lower.CL,
      "ci_upper" = upper.CL
    )
}
```

##### Plot rate of aging gender

```{r plot_age_rate_gender}
if (isFALSE(params$telolength)) {
  p0 <- ggplot() +
    geom_point(
      data = df,
      aes(age, .data[[params$hearing_var]])
    ) +
    geom_line(
      data = output_age_rate$predictions$linear,
      aes(x = age, y = estimate)
    ) +
    geom_ribbon(
      data = output_age_rate$predictions$linear,
      aes(x = age, ymin = ci_lower, ymax = ci_upper)
    ) +
    construct_axes(p_cfg$x$age, p_cfg$y$hl) +
    p_cfg$plot_theme +
    theme(
      axis.title.x = element_blank(),
      plot.margin = margin(1, 0, 0, 1, unit = "pt")
    )

  p1 <- ggplot() +
    geom_point(
      data = df,
      aes(age, .data[[params$hearing_var]])
    ) +
    geom_line(
      data = output_age_rate$predictions$quadratic,
      aes(x = age, y = estimate)
    ) +
    geom_ribbon(
      data = output_age_rate$predictions$quadratic,
      aes(x = age, ymin = ci_lower, ymax = ci_upper)
    ) +
    construct_axes(p_cfg$x$age, p_cfg$y$hl) +
    p_cfg$plot_theme +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      plot.margin = margin(1, 0, 0, 1, unit = "pt")
    )

  p2 <- ggplot() +
    geom_point(
      data = df,
      aes(age, .data[[params$hearing_var]])
    ) +
    geom_ribbon(
      data = gender_predictions,
      aes(
        x = age,
        ymin = ci_lower,
        ymax = ci_upper,
        fill = gender
      ),
      show.legend = FALSE
    ) +
    geom_line(
      data = gender_predictions,
      aes(x = age, y = estimate, color = gender),
      show.legend = FALSE
    ) +
    construct_axes(p_cfg$x$age, p_cfg$y$hl) +
    p_cfg$colors$gender +
    p_cfg$plot_theme +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      plot.margin = margin(1, 1, 0, 0, unit = "pt")
    )

  p3 <- ggplot() +
    geom_hline(
      yintercept = 0
    ) +
    geom_line(
      data = output_age_rate$predictions$linear_derivative,
      aes(x = age, y = derivative)
    ) +
    geom_ribbon(
      data = output_age_rate$predictions$linear_derivative,
      aes(x = age, ymin = ci_lower, ymax = ci_upper)
    ) +
    construct_axes(p_cfg$x$age, p_cfg$y$hl_rate) +
    p_cfg$plot_theme +
    theme(plot.margin = margin(0, 0, 0, 1, unit = "pt"))

  p4 <- ggplot() +
    geom_hline(
      yintercept = 0
    ) +
    geom_line(
      data = output_age_rate$predictions$quadratic_derivative,
      aes(x = age, y = derivative)
    ) +
    geom_ribbon(
      data = output_age_rate$predictions$quadratic_derivative,
      aes(x = age, ymin = ci_lower, ymax = ci_upper)
    ) +
    construct_axes(p_cfg$x$age, p_cfg$y$hl_rate) +
    p_cfg$plot_theme +
    theme(
      axis.title.y = element_blank(),
      plot.margin = margin(0, 0, 0, 1, unit = "pt")
    )

  p5 <- ggplot() +
    geom_hline(
      yintercept = 0
    ) +
    geom_ribbon(
      data = gender_predictions_der,
      aes(
        x = age,
        ymin = ci_lower,
        ymax = ci_upper,
        fill = gender
      ),
      show.legend = FALSE
    ) +
    geom_line(
      data = gender_predictions_der,
      aes(x = age, y = derivative, color = gender),
      show.legend = FALSE
    ) +
    construct_axes(p_cfg$x$age, p_cfg$y$hl_rate) +
    p_cfg$colors$gender +
    p_cfg$plot_theme +
    theme(
      axis.title.y = element_blank(),
      plot.margin = margin(0, 1, 0, 0, unit = "pt")
    )

  p <- align_plots(p0, p1, p2, p3, p4, p5, align = "hv")

  p <-
    plot_grid(
      plot_grid(
        p[[1]],
        NULL,
        p[[2]],
        NULL,
        p[[3]],
        labels = c("A", "", "B", "", "C"),
        nrow = 1,
        rel_widths = c(1, -0.01, 1, -0.01, 1),
        label_x = 0.04,
        label_y = 1.02
      ),
      NULL,
      plot_grid(
        p[[4]],
        NULL,
        p[[5]],
        NULL,
        p[[6]],
        labels = c("D", "", "E", "", "F"),
        nrow = 1,
        rel_widths = c(1, -0.01, 1, -0.01, 1),
        label_x = 0.04,
        label_y = 1.02
      ),
      ncol = 1,
      rel_heights = c(1, -0.15, 1)
    )

  sameplot(
    p,
    file = here(
      p_cfg$figure_dir,
      "Extras",
      "simplified_age_rate_panel_gender.png"
    ),
    width = params$double_column * 1.8,
    height = 3 * 1.8,
    scaling = 1.8,
    res = 600,
    save = params$save_figures
  )
}
```

##### Plot aging means with and without confounders

###### Plot aging mean without confounders

```{r plot_aging_mean}
if (isFALSE(params$telolength)) {
  p0 <- ggplot() +
    geom_point(
      data = df,
      aes(age, .data[[params$hearing_var]])
    ) +
    geom_line(
      data = output_age_rate$predictions$smooth,
      aes(x = age, y = estimate),
      color = "grey80"
    ) +
    geom_line(
      data = df %>%
        group_by(age) %>%
        summarize(hearing_var = mean(.data[[params$hearing_var]]), .groups = "drop"),
      aes(x = age, y = hearing_var)
    ) +
    construct_axes(p_cfg$x$age, p_cfg$y$hl) +
    p_cfg$plot_theme

  sameplot(
    p0,
    file = here(p_cfg$figure_dir, "Extras", "age_rate_mean.png"),
    width = params$single_column,
    height = 3,
    res = 600,
    save = params$save_figures
  )
}
```

###### Plot aging mean per gender

```{r plot_gender_aging_mean}
if (isFALSE(params$telolength)) {
  p0 <- ggplot() +
    geom_point(
      data = df,
      aes(age, .data[[params$hearing_var]])
    ) +
    geom_line(
      data = gender_predictions,
      aes(x = age, y = estimate, color = gender),
      show.legend = FALSE
    ) +
    geom_line(
      data = df %>%
        group_by(age, gender) %>%
        summarize(hearing_var = mean(.data[[params$hearing_var]]), .groups = "drop"),
      aes(x = age, y = hearing_var, color = gender),
      show.legend = FALSE
    ) +
    construct_axes(p_cfg$x$age, p_cfg$y$hl) +
    p_cfg$colors$gender +
    p_cfg$plot_theme

  sameplot(
    p0,
    file = here(p_cfg$figure_dir, "Extras", "age_rate_mean.png"),
    width = params$single_column,
    height = 3,
    res = 600,
    save = params$save_figures
  )
}
```

###### Plot aging mean per race

```{r plot_race_aging_mean}
if (isFALSE(params$telolength)) {
  p0 <- ggplot() +
    geom_point(
      data = df,
      aes(age, .data[[params$hearing_var]])
    ) +
    geom_line(
      data = df %>%
        group_by(age, race) %>%
        summarize(hearing_var = mean(.data[[params$hearing_var]]), .groups = "drop"),
      aes(x = age, y = hearing_var, color = race)
    ) +
    construct_axes(p_cfg$x$age, p_cfg$y$hl) +
    p_cfg$colors$race +
    p_cfg$plot_theme

  sameplot(
    p0,
    file = here(p_cfg$figure_dir, "Extras", "age_rate_mean_race.png"),
    width = params$single_column,
    height = 3,
    res = 600,
    save = TRUE
  )
}
```

### Bioage prediction panel

#### Bioage MSE plot

```{r plot_bioage}
update_geom_defaults("point", p_cfg$geom_args$mse_point)
update_geom_defaults("line", p_cfg$geom_args$mse_line)

p <- df_summary_bioage %>%
  ggplot(aes(x = age_group, y = mean_delta_mse)) +
  geom_hline(yintercept = 0) +
  geom_ribbon(
    aes(
      fill = predictors,
      group = predictors,
      ymin = ci_low_delta_mse,
      ymax = ci_high_delta_mse
    ),
    show.legend = FALSE
  ) +
  geom_line(
    aes(color = predictors, group = predictors),
    show.legend = TRUE
  ) +
  geom_point(
    aes(color = predictors),
    show.legend = TRUE
  ) +
  geom_text(
    data = . %>%
      filter(p_adj != "ns"),
    aes(
      x = age_group,
      y = ci_high_delta_mse + 0.25,
      label = signif,
      color = predictors
    ),
    size = p_cfg$signif_size / (14 / 5)
  ) +
  construct_axes(p_cfg$x$age_group, p_cfg$y$dmse_bioage) +
  p_cfg$colors$bioage +
  p_cfg$plot_theme

sameplot(
  p,
  file = here(p_cfg$figure_dir, "bioage_panel.png"),
  width = params$single_column,
  height = 2.15,
  res = 600,
  save = params$save_figures
)
```

#### Bioage RMSE plot (paper)

```{r}
p <- df_summary_bioage %>%
  ggplot(aes(x = age_group, y = delta_rmse)) +
  geom_hline(yintercept = 0) +
  geom_ribbon(
    aes(
      fill = predictors,
      group = predictors,
      ymin = ci_low_delta_rmse,
      ymax = ci_high_delta_rmse
    ),
    show.legend = FALSE
  ) +
  geom_line(
    aes(color = predictors, group = predictors),
    show.legend = TRUE
  ) +
  geom_point(
    aes(color = predictors),
    show.legend = TRUE
  ) +
  geom_text(
    data = . %>%
      filter(p_adj != "ns"),
    aes(
      x = age_group,
      y = ci_high_delta_rmse + 0.25,
      label = signif,
      color = predictors
    ),
    size = p_cfg$signif_size / (14 / 5)
  ) +
  construct_axes(p_cfg$x$age_group, p_cfg$y$drmse_bioage) +
  p_cfg$colors$bioage +
  p_cfg$plot_theme

sameplot(
  p,
  file = here(p_cfg$figure_dir, "bioage_panel.png"),
  width = params$single_column,
  height = 2.15,
  res = 600,
  save = params$save_figures
)
```



#### Additional Plots

##### Plot PhenoAge Only

```{r plot_phenoage_only, warning=FALSE}
if (isFALSE(params$telolength)) {
  p <- df_summary_bioage %>%
    filter(predictors %in% c("age", "age, phenoage")) %>%
    ggplot(aes(x = age_group, y = mean_delta_mse)) +
    geom_hline(
      yintercept = 0
    ) +
    geom_ribbon(
      aes(
        fill = predictors,
        group = predictors,
        ymin = ci_low_delta_mse,
        ymax = ci_high_delta_mse
      ),
      show.legend = FALSE
    ) +
    geom_line(
      aes(color = predictors, group = predictors),
      show.legend = TRUE
    ) +
    geom_point(
      aes(color = predictors),
      show.legend = TRUE
    ) +
    geom_text(
      data = . %>%
        filter(p_adj != "ns"),
      aes(
        x = age_group,
        y = ci_high_delta_mse + 0.25,
        label = signif,
        color = predictors
      ),
      manual = TRUE,
      size = p_cfg$signif_size / (14 / 5)
    ) +
    construct_axes(p_cfg$x$age_group, p_cfg$y$dmse_bioage) +
    p_cfg$colors$phenoage +
    p_cfg$plot_theme

  sameplot(
    p,
    file = here(
      p_cfg$figure_dir,
      "Extras",
      "phenoage_only_bioage_panel.png"
    ),
    width = params$single_column * 1.8,
    height = 2.15 * 1.8,
    scaling = 1.8,
    res = 600,
    save = params$save_figures
  )
}
```

##### Advance bioage plot

```{r }
p <- df %>%
  ggplot(aes(x = age, y = phenoage_advance)) +
  geom_density_ridges(aes(x = phenoage_advance, y = age_group)) +
  annotate(
    "text",
    x = Inf,
    y = Inf,
    label = "Decelerated Aging",
    hjust = 2.7,
    vjust = 1,
    size = p_cfg$text_size_primary / (14 / 5),
    color = "darkgreen"
  ) +
  annotate(
    "text",
    x = Inf,
    y = Inf,
    label = "Accelerated Aging",
    hjust = 1,
    vjust = 1,
    size = p_cfg$text_size_primary / (14 / 5),
    color = "darkred"
  ) +
  construct_axes(p_cfg$x$age_accel, p_cfg$y$age_group) +
  p_cfg$colors$bioage +
  p_cfg$plot_theme

sameplot(
  p,
  file = here(p_cfg$figure_dir, "Extras", "age_acceleration.png"),
  width = params$single_column,
  height = 2.15,
  res = 600,
  save = params$save_figures
)
```



##### Confounders predictive capacity plot

```{r plot_confounders, warning=FALSE}
if (isFALSE(params$telolength)) {
  update_geom_defaults("point", p_cfg$geom_args$mse_point)
  update_geom_defaults("line", p_cfg$geom_args$mse_line)

  p <- df_summary_confounders %>%
    ggplot(aes(x = age_group, y = mean_delta_mse)) +
    geom_hline(
      yintercept = 0
    ) +
    geom_ribbon(
      aes(
        fill = predictors,
        group = predictors,
        ymin = ci_low_delta_mse,
        ymax = ci_high_delta_mse
      ),
      show.legend = FALSE
    ) +
    geom_line(
      aes(color = predictors, group = predictors),
      show.legend = TRUE
    ) +
    geom_point(
      aes(color = predictors),
      show.legend = TRUE
    ) +
    geom_text(
      data = . %>%
        filter(p_adj != "ns"),
      aes(
        x = age_group,
        y = ci_high_delta_mse,
        label = signif,
        color = predictors
      ),
      manual = TRUE,
      size = p_cfg$signif_size / (14 / 5)
    ) +
    construct_axes(p_cfg$x$age_group, p_cfg$y$dmse_conf) +
    p_cfg$colors$confounders +
    p_cfg$plot_theme

  sameplot(
    p,
    file = here(p_cfg$figure_dir, "confounder_panel.png"),
    width = params$single_column,
    height = 2.15,
    res = 600,
    save = params$save_figures
  )
}
```

##### Plot additional variance confounders

```{r plot_confounders_var, warning=FALSE}
if (isFALSE(params$telolength)) {
  p <- df_summary_confounders %>%
    ggplot(aes(x = age_group, y = mean_additional_variance)) +
    geom_hline(
      yintercept = 0
    ) +
    geom_ribbon(
      aes(
        fill = predictors,
        group = predictors,
        ymin = ci_low_additional_variance,
        ymax = ci_high_additional_variance
      ),
      show.legend = FALSE
    ) +
    geom_line(
      aes(color = predictors, group = predictors),
      show.legend = TRUE
    ) +
    geom_point(
      aes(color = predictors),
      show.legend = TRUE
    ) +
    geom_text(
      data = . %>%
        filter(p_adj != "ns"),
      aes(
        x = age_group,
        y = ci_high_additional_variance,
        label = signif,
        color = predictors
      ),
      manual = TRUE,
      size = p_cfg$signif_size / (14 / 5)
    ) +
    construct_axes(p_cfg$x$age_group, p_cfg$y$add_pred_conf) +
    p_cfg$colors$confounders +
    p_cfg$plot_theme

  sameplot(
    p,
    file = here(p_cfg$figure_dir, "confounder_panel_var.png"),
    width = params$single_column,
    height = 2.15,
    res = 600,
    save = params$save_figures
  )
}
```

##### Biomarker plots

###### Plot Delta mse Biomarkers (only relevant ones)

```{r plot_biomarkers, include=FALSE}
if (isFALSE(params$telolength)) {
  p1 <- df_summary_biomarkers %>%
    filter(predictors %in% c("age, crp", "age, hba1c", "age, albumin_gL")) %>%
    mutate(
      label = predictors,
      label = case_when(
        label == "age, crp" ~ "Age + CRP",
        label == "age, hba1c" ~ "Age + HbA1c",
        label == "age, albumin_gL" ~ "Age + ALB",
        TRUE ~ NA
      ),
      predictors = factor(
        predictors,
        levels = c("age, crp", "age, hba1c", "age, albumin_gL")
      )
    ) %>%
    ggplot(aes(x = age_group, y = mean_delta_mse)) +
    geom_hline(
      yintercept = 0
    ) +
    geom_label(
      data = . %>%
        filter(age_group == "20-30"),
      aes(
        x = "20-30",
        y = 3.65,
        label = label,
        fill = predictors
      ),
      size = 8 / (14 / 5),
      hjust = 0,
      label.padding = unit(0.3, "lines"),
      nudge_x = -0.1,
      color = "white",
      show.legend = FALSE
    ) +
    geom_ribbon(
      aes(
        fill = predictors,
        group = predictors,
        ymin = ci_low_delta_mse,
        ymax = ci_high_delta_mse
      ),
      show.legend = FALSE
    ) +
    geom_line(
      aes(color = predictors, group = predictors),
      show.legend = FALSE
    ) +
    geom_point(
      aes(color = predictors),
      show.legend = FALSE
    ) +
    facet_wrap(
      ~predictors,
      ncol = 1,
      nrow = 5,
      labeller = labeller(
        predictors = c(
          "age, crp" = "Age + CRP",
          "age, hba1c" = "Age + HbA1c",
          "age, albumin_gL" = "Age + ALB"
        )
      )
    ) +
    construct_axes(p_cfg$x$age_group, p_cfg$y$dmse_best_biomark) +
    p_cfg$colors$best_biomarkers +
    p_cfg$plot_theme +
    theme(strip.text.x = element_blank())

  sameplot(
    p1,
    file = here(p_cfg$figure_dir, "biomarkers_biomarkers_panel.png"),
    width = 2.5,
    height = 3,
    res = 600,
    save = FALSE
  )
}
```

###### Plot biomarkers together vs biomarkers + phenoage

```{r plot_explained_var, include=FALSE}
if (isFALSE(params$telolength)) {
  p2 <- df_summary_biomarkers %>%
    filter(
      predictors %in% c(
        "age, phenoage",
        "age, crp, hba1c, albumin_gL",
        "age, crp, hba1c, albumin_gL, phenoage"
      )
    ) %>%
    mutate(predictors = factor(
      predictors,
      levels = c(
        "age, phenoage",
        "age, crp, hba1c, albumin_gL",
        "age, crp, hba1c, albumin_gL, phenoage"
      )
    )) %>%
    ggplot(aes(x = age_group, y = mean_delta_mse)) +
    geom_hline(
      yintercept = 0,
    ) +
    geom_ribbon(
      aes(
        fill = predictors,
        group = predictors,
        ymin = ci_low_delta_mse,
        ymax = ci_high_delta_mse
      ),
      show.legend = FALSE
    ) +
    geom_line(
      aes(color = predictors, group = predictors),
      show.legend = TRUE
    ) +
    geom_point(
      aes(color = predictors),
      show.legend = FALSE
    ) +
    construct_axes(p_cfg$x$age_group, p_cfg$y$dmse_bioage) +
    p_cfg$colors$biomarker_diff +
    p_cfg$plot_theme +
    theme(
      axis.title.y = element_blank(),
      plot.margin = margin(1, 1, 0, 0, unit = "pt")
    )

  sameplot(
    p2,
    file = here(p_cfg$figure_dir, "biomarkers_v_phenoage.png"),
    width = 3,
    height = 4,
    res = 600,
    save = FALSE
  )
}
```

###### Composite plot

```{r plot_biomark_panel}
if (isFALSE(params$telolength)) {
  p <- ggarrange(
    p1,
    NULL,
    p2,
    nrow = 1,
    widths = c(0.6, 0, 1),
    labels = c("A", "", "B"),
    label.x = 0.02,
    label.y = 1.01,
    align = "v"
  )

  sameplot(
    p,
    file = here(p_cfg$figure_dir, "biomarkers_fig_panel.png"),
    width = params$double_column,
    height = 3,
    res = 600,
    save = params$save_figures
  )
}
```

##### Biomarkers supplemental plots

```{r plot_biomarker_all, warning=FALSE}
if (isFALSE(params$telolength)) {
  update_geom_defaults("linerange", p_cfg$geom_args$biomarker_linerange)
  update_geom_defaults("point", p_cfg$geom_args$biomarker_point)

  p <- df_summary_biomarkers_overall %>%
    filter(!(
      predictors %in% c(
        "age",
        "age, phenoage",
        "age, crp, hba1c, albumin_gL",
        "age, crp, hba1c, albumin_gL, phenoage"
      )
    )) %>%
    mutate(predictors = factor(
      predictors,
      levels = c(
        "age, sbp",
        "age, bun",
        "age, mcv",
        "age, lymph",
        "age, rdw",
        "age, totchol",
        "age, creat",
        "age, uap",
        "age, glucose",
        "age, wbc",
        "age, alp",
        "age, albumin_gL",
        "age, hba1c",
        "age, crp"
      ),
      labels = c(
        "SBP",
        "BUN",
        "MCV",
        "Lymphs",
        "RDW",
        "TC",
        "Creat",
        "UA",
        "Glc",
        "WBC",
        "ALP",
        "ALB",
        "HbA1c",
        "CRP"
      )
    )) %>%
    ggplot(aes(x = predictors, y = mean_delta_mse)) +
    geom_point(
      aes(color = predictors, fill = predictors),
      position = position_dodge(0.8),
      show.legend = FALSE
    ) +
    geom_linerange(
      aes(
        color = predictors,
        ymin = ci_low_delta_mse,
        ymax = ci_high_delta_mse
      ),
      position = position_dodge(0.8),
      lineend = "round",
      show.legend = FALSE
    ) +
    geom_text(
      data = . %>%
        filter(p_adj != "ns"),
      aes(
        x = predictors,
        y = ci_high_delta_mse + 0.25,
        label = signif,
        color = predictors
      ),
      size = p_cfg$signif_size / (14 / 5),
      show.legend = FALSE
    ) +
    construct_axes(p_cfg$y$dmse_biomark) +
    p_cfg$plot_theme +
    theme(
      panel.grid.major.y = element_line(color = "grey95", linewidth = 0.25),
      axis.text.x = element_markdown(
        "sans",
        size = 8,
        angle = 45,
        hjust = 1,
        vjust = 1
      ),
      axis.title.x = element_blank()
    )

  sameplot(
    p,
    file = here(p_cfg$figure_dir, "biomarkers_panel.png"),
    width = params$single_column,
    height = 2.15,
    res = 600,
    save = params$save_figures
  )
}
```
