# Visualization

## Tables

#### Descriptives table

#### Format table

Round R² and delta RMSE, rename variables, set p values to bold and to " <
0.001" when very small.

```{r format_table}
format_table <- function(summary_data, per_age_group) {
  if (per_age_group) {
    table <- summary_data %>%
      mutate(r2 = round(r2, 3),
             delta_rmse = round(delta_rmse, 2)) %>%
      rename(
        "Predictors" = predictors,
        "Age" = age_group,
        "∆RMSE (vs age, dB HL)" = delta_rmse,
        "R²" = r2,
        "p (vs age)" = p_adj
      ) %>%
      mutate(
        `p (vs age)` = case_when(
          signif == "***" ~
            cell_spec("< 0.001", "html", bold = TRUE),
          suppressWarnings(!is.na(as.numeric(`p (vs age)`)) &
                             as.numeric(`p (vs age)`) < 0.05) &
            `p (vs age)` != "< 0.001" ~
            cell_spec(suppressWarnings(round(
              as.numeric(`p (vs age)`), 3
            )), "html", bold = TRUE),
          TRUE ~ `p (vs age)`
        ),
        `p (vs age)` = as.character(`p (vs age)`),
        `p (vs age)` = case_when(Predictors == "age" ~ "", TRUE ~ `p (vs age)`),
        `∆RMSE (vs age, dB HL)` = as.character(`∆RMSE (vs age, dB HL)`),
        `∆RMSE (vs age, dB HL)` = case_when(Predictors == "age" ~ "", TRUE ~ `∆RMSE (vs age, dB HL)`)
      ) %>%
      select(Predictors, Age, n, `R²`, `∆RMSE (vs age, dB HL)`, `p (vs age)`)
  } else {
    table <- summary_data %>%
      mutate(r2 = round(r2, 3),
             delta_rmse = round(delta_rmse, 2)) %>%
      rename(
        "Predictors" = predictors,
        "∆RMSE (vs age, dB HL)" = delta_rmse,
        "R²" = r2,
        "p (vs age)" = p_adj
      ) %>%
      mutate(
        `p (vs age)` = case_when(
          signif == "***" ~
            cell_spec("< 0.001", "html", bold = TRUE),
          suppressWarnings(!is.na(as.numeric(`p (vs age)`)) &
                             as.numeric(`p (vs age)`) < 0.05) &
            `p (vs age)` != "< 0.001" ~
            cell_spec(suppressWarnings(round(
              as.numeric(`p (vs age)`), 3
            )), "html", bold = TRUE),
          TRUE ~ `p (vs age)`
        ),
        `p (vs age)` = as.character(`p (vs age)`),
        `p (vs age)` = case_when(Predictors == "age" ~ "", TRUE ~ `p (vs age)`),
        `∆RMSE (vs age, dB HL)` = as.character(`∆RMSE (vs age, dB HL)`),
        `∆RMSE (vs age, dB HL)` = case_when(Predictors == "age" ~ "", TRUE ~ `∆RMSE (vs age, dB HL)`)
      ) %>%
      select(Predictors, n, `R²`, `∆RMSE (vs age, dB HL)`, `p (vs age)`)
  }
  table <- table %>%
    rename("p (vs full model)" = `p (vs age)`,
           "∆RMSE (vs full model, dB HL)" = `∆RMSE (vs age, dB HL)`,
    )
  
  return(table)
}

table_bioage <- format_table(df_summary_bioage, per_age_group = TRUE)
table_bioage_overall <- format_table(df_summary_bioage_overall, per_age_group = FALSE)

if (isFALSE(params$telolength)) {
  table_biomarkers <- format_table(df_summary_biomarkers, per_age_group = TRUE)
  table_biomarkers_overall <- format_table(df_summary_biomarkers_overall, per_age_group = FALSE)
  
  table_confounders <- format_table(df_summary_confounders, per_age_group = TRUE)
  table_confounders_overall <- format_table(df_summary_confounders_overall, per_age_group = FALSE)
}
```

#### Order, Filter, and Rename Tables

```{r order_table}
rename_bioage_table <- function(bioage_table) {
  if (isFALSE(params$telolength)) {
    bioage_table <- bioage_table %>%
      mutate(
        Predictors = case_when(
          Predictors == "age" ~ "Age",
          Predictors == "age, hd_log" ~ "Age + HD",
          Predictors == "age, kdm" ~ "Age + KDM",
          Predictors == "age, phenoage" ~ "Age + PhenoAge"
        ),
        Predictors = factor(
          Predictors,
          levels = c("Age", "Age + HD", "Age + KDM", "Age + PhenoAge")
        )
      ) %>%
      arrange(Predictors)
  } else {
    bioage_table <- bioage_table %>%
      mutate(
        Predictors = case_when(
          Predictors == "age" ~ "Age",
          Predictors == "age, hd_log" ~ "Age + HD",
          Predictors == "age, kdm" ~ "Age + KDM",
          Predictors == "age, phenoage" ~ "Age + PhenoAge",
          Predictors == "age, telolength" ~ "Age + Telolength"
        ),
        Predictors = factor(
          Predictors,
          levels = c(
            "Age",
            "Age + HD",
            "Age + KDM",
            "Age + PhenoAge",
            "Age + Telolength"
          )
        )
      ) %>%
      arrange(Predictors)
  }
  
  return(bioage_table)
}

rename_biomarker_table <- function(biomarker_table) {
  biomarker_table <- biomarker_table %>%
    filter(!(
      Predictors %in% c(
      "age",
      "age, phenoage",
      "age, crp, hba1c, albumin_gL",
      "age, crp, hba1c, albumin_gL, phenoage"
      )
    )) %>% mutate(
      Predictors = case_when(
        Predictors == "age, rdw" ~ "Age + RDW",
        Predictors == "age, alp" ~ "Age + ALP",
        Predictors == "age, crp" ~ "Age + CRP",
        Predictors == "age, totchol" ~ "Age + TC",
        Predictors == "age, albumin_gL" ~ "Age + ALB",
        Predictors == "age, hba1c" ~ "Age + HbA1c",
        Predictors == "age, glucose" ~ "Age + Glc",
        Predictors == "age, sbp" ~ "Age + SBP",
        Predictors == "age, bun" ~ "Age + BUN",
        Predictors == "age, mcv" ~ "Age + MCV",
        Predictors == "age, wbc" ~ "Age + WBC",
        Predictors == "age, lymph" ~ "Age + Lymphs",
        Predictors == "age, uap" ~ "Age + UA",
        Predictors == "age, creat" ~ "Age + Creat"
      ),
      Predictors = factor(Predictors, levels = c(
      "Age + SBP",
      "Age + BUN",
      "Age + MCV",
      "Age + Lymphs",
      "Age + RDW",
      "Age + TC",
      "Age + Creat",
      "Age + UA",
      "Age + Glc",
      "Age + WBC",
      "Age + ALP",
      "Age + ALB",
      "Age + HbA1c",
      "Age + CRP"
      )
    )) %>%
    arrange(Predictors)
  
  return(biomarker_table)
}

rename_confounder_table <- function(confounders_table) {
  confounders_table <- confounders_table %>%
    filter(!(
      Predictors %in% c(
      "full"
      )
    )) %>% mutate(
      Predictors = case_when(
        Predictors == "full_with_phenoage" ~ "Phenoage",
        Predictors == "no_gender" ~ "Gender",
        Predictors == "no_smoking" ~ "Smoking",
        Predictors == "no_race" ~ "Race",
        Predictors == "no_education" ~ "Education",
        Predictors == "no_gender, race" ~ "Gender * Race"
      ),
      Predictors = factor(Predictors, levels = c(
      "Phenoage",
      "Gender",
      "Smoking",
      "Race",
      "Education",
      "Gender * Race"
      )
    )) %>%
    arrange(Predictors)
  
  return(confounders_table)
}

table_bioage <- rename_bioage_table(table_bioage)
table_bioage_overall <- rename_bioage_table(table_bioage_overall)

if (isFALSE(params$telolength)) {
table_biomarkers <- rename_biomarker_table(table_biomarkers)
table_biomarkers_overall <- rename_biomarker_table(table_biomarkers_overall)

table_confounders <- rename_confounder_table(table_confounders)
table_confounders_overall <- rename_confounder_table(table_confounders_overall)
}
```

#### Render table

```{r store_tables}
render_and_save_table <- function(table, path) {
  html_table <- table %>%
    kable(
      format = "html",
      align = c("l", "l", "r", "r", "r", "r", "r", "r", "r"),
      escape = FALSE
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE,
      position = "center"
    ) %>%
    collapse_rows(columns = 1, valign = "top")
  
  if (isTRUE(params$save_figures)) {
  save_kable(html_table, path)
  }
  
  return(html_table)
}

render_and_save_table(table_bioage, path = here("Output", "Figures", params$hearing_var, "table_bioage.html"))
render_and_save_table(table_bioage_overall, path = here("Output", "Figures", params$hearing_var, "table_bioage_overall.html"))

if (isFALSE(params$telolength)) {
render_and_save_table(table_biomarkers, path = here("Output", "Figures", params$hearing_var, "table_biomarkers.html")) %>%
  scroll_box(width = "100%", height = "400px")
render_and_save_table(table_biomarkers_overall, path = here("Output", "Figures", params$hearing_var, "table_biomarkers_overall.html"))

render_and_save_table(table_confounders, path = here("Output", "Figures", params$hearing_var, "table_confounders.html")) %>%
  scroll_box(width = "100%", height = "400px")
render_and_save_table(table_confounders_overall, path = here("Output", "Figures", params$hearing_var, "table_confounders_overall.html"))
}
```

### Confounders table

#### Build table and recode values

```{r build_conf_table}
build_confounder_table <- function(data, y = params$hearing_var) {
  variables = c("Gender", "Smoking", "Race", "Education")
  
  data <- data %>%
    mutate(
      gender = dplyr::recode(gender, "1" = "Male", "2" = "Female"),
      smoking = dplyr::recode(smoking, "1" = "Smoker", "2" = "Non-smoker"),
      race = dplyr::recode(
        race,
        "1" = "Hispanic",
        "3" = "White",
        "4" = "Black",
        "5" = "Other"
      ),
      education = dplyr::recode(
        education,
        "1" = "< 9th grade",
        "2" = "9-11th grade",
        "3" = "High school",
        "4" = "Some college or AA degree",
        "5" = "College graduate or higher"
      )
    ) %>%
    rename(
      "Smoking" = smoking,
      "Gender" = gender,
      "Race" = race,
      "Education" = education
    )

bind_rows(lapply(variables, function(variable)
  data %>%
    mutate(level = .data[[variable]]) %>%
    group_by(age_group, level) %>%
    summarise(
      mean = mean(.data[[y]], na.rm = TRUE),
      sd   = sd(.data[[y]],   na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(variable = factor(variable, levels = c("Gender", "Smoking", "Race", "Education")), .before = 1)
))
}

if (isFALSE(params$telolength)) {
table_confounders_descriptive <- build_confounder_table(df)
}
```

#### Pivot table and merge mean and standard deviation

```{r pivot_conf_table}
if (isFALSE(params$telolength)) {
table_confounders_descriptive <- table_confounders_descriptive %>%
  mutate(stat = sprintf("%.2f (%.2f)", mean, sd)) %>%
  select(variable, level, age_group, stat) %>%
  pivot_wider(names_from = age_group, values_from = stat)
}
```

#### Format table

```{r format_Conf_table}
if (isFALSE(params$telolength)) {
table_confounders_descriptive <- table_confounders_descriptive %>%
  group_by(variable) %>%
  mutate(variable = if_else(row_number() == 1L, variable, "")) %>%
  ungroup() %>%
  rename(
    " " = variable,
    "  " = level
  )
}
```

#### Save table

```{r store_conf_table}
if (isFALSE(params$telolength)) {
render_and_save_table(table_confounders_descriptive, path = here("Output", "Figures", params$hearing_var, "table_confounders_descriptives.html"))
}
```

## Figures

### Rate of aging panel

#### Plot Rate of Aging

```{r plot_aging, warning=FALSE}
p_config <- load_plot_config(output_type = params$plot_config, params)

p0_a <- ggplot() +
  geom_point(
    data = df,
    aes(age, .data[[params$hearing_var]]),
    alpha = 0.1,
    size = 0.5,
    shape = 1,
    stroke = 0.25,
    color = "grey30"
  ) +
  geom_line(
    data = output_age_rate$predictions$linear,
    aes(x = age, y = estimate),
    linewidth = 0.6,
    alpha = 1
  ) +
  geom_ribbon(
    data = output_age_rate$predictions$linear,
    aes(x = age, ymin = ci_lower, ymax = ci_upper),
    alpha = 0.15
  ) +
  scale_y_continuous(breaks = c(0, 30, 60, 90)) +
  scale_x_continuous(breaks = seq(20, 80, 10)) +
  coord_cartesian(ylim = c(-10, 100), xlim = c(20, 82))

p0_a <- p_config$preset(p0_a, "age_rate") +
  theme(axis.title.x = element_blank(),
        plot.margin = margin(1, 0, 0, 1, unit = "pt"))

p1_a <- ggplot() +
  geom_point(
    data = df,
    aes(age, .data[[params$hearing_var]]),
    alpha = 0.1,
    size = 0.5,
    shape = 1,
    stroke = 0.25,
    color = "grey30"
  ) +
  geom_line(
    data = output_age_rate$predictions$quadratic,
    aes(x = age, y = estimate),
    linewidth = 0.6,
    alpha = 1
  ) +
  geom_ribbon(
    data = output_age_rate$predictions$quadratic,
    aes(x = age, ymin = ci_lower, ymax = ci_upper),
    alpha = 0.15
  ) +
  scale_y_continuous(breaks = c(0, 30, 60, 90)) +
  scale_x_continuous(breaks = seq(20, 80, 10)) +
  coord_cartesian(ylim = c(-10, 100), xlim = c(20, 82))

p1_a <- p_config$preset(p1_a, "age_rate") +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(1, 0, 0, 1, unit = "pt")
  )

p2_a <- ggplot() +
  geom_rect(
    data = pwc_age_rate_boundaries,
    aes(
      xmin = x_min,
      xmax = x_max,
      ymin = -Inf,
      ymax = Inf
    ),
    alpha = 0.125,
    fill = "#ff0000"
  ) +
  geom_point(
    data = df,
    aes(age, .data[[params$hearing_var]]),
    alpha = 0.1,
    size = 0.5,
    shape = 1,
    stroke = 0.25,
    color = "grey30"
  ) +
  geom_ribbon(
    data = output_age_rate$predictions$smooth,
    aes(
      x = age,
      ymin = ci_lower + coef(output_age_rate$models$smooth)["(Intercept)"],
      ymax = ci_upper + coef(output_age_rate$models$smooth)["(Intercept)"]
    ),
    alpha = 0.15
  ) +
  geom_line(
    data = output_age_rate$predictions$smooth,
    aes(
      x = age,
      y = estimate + coef(output_age_rate$models$smooth)["(Intercept)"]
    ),
    linewidth = 0.6
  ) +
  scale_y_continuous(breaks = c(0, 30, 60, 90)) +
  scale_x_continuous(breaks = seq(20, 80, 10)) +
  coord_cartesian(ylim = c(-10, 100), xlim = c(20, 82))

p2_a <- p_config$preset(p2_a, "age_rate") +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.margin = margin(1, 1, 0, 0, unit = "pt")
  )

p3_a <- ggplot() +
  geom_hline(
    yintercept = 0,
    linetype = "22",
    color = "grey20",
    linewidth = 0.2
  )  +
  geom_line(
    data = output_age_rate$predictions$linear_derivative,
    aes(x = age, y = derivative),
    linewidth = 0.6
  ) +
  geom_ribbon(
    data = output_age_rate$predictions$linear_derivative,
    aes(x = age, ymin = ci_lower, ymax = ci_upper),
    alpha = 0.15
  ) +
  scale_y_continuous(breaks = c(0, 1, 2)) +
  scale_x_continuous(breaks = seq(20, 80, 10)) +
  coord_cartesian(ylim = c(-0.5, 2.6), xlim = c(20, 82))

p3_a <- p_config$preset(p3_a, "age_rate_der") +
  theme(plot.margin = margin(0, 0, 0, 1, unit = "pt"))

p4_a <- ggplot() +
  geom_hline(
    yintercept = 0,
    linetype = "22",
    color = "grey20",
    linewidth = 0.2
  )  +
  geom_line(
    data = output_age_rate$predictions$quadratic_derivative,
    aes(x = age, y = derivative),
    linewidth = 0.6
  ) +
  geom_ribbon(
    data = output_age_rate$predictions$quadratic_derivative,
    aes(x = age, ymin = ci_lower, ymax = ci_upper),
    alpha = 0.15
  ) +
  scale_y_continuous(breaks = c(0, 1, 2)) +
  scale_x_continuous(breaks = seq(20, 80, 10)) +
  coord_cartesian(ylim = c(-0.5, 2.6), xlim = c(20, 82))

p4_a <- p_config$preset(p4_a, "age_rate_der") +
  theme(axis.title.y = element_blank(),
        plot.margin = margin(0, 0, 0, 1, unit = "pt"))

p5_a <- ggplot() +
  geom_rect(
    data = pwc_age_rate_boundaries,
    aes(
      xmin = x_min,
      xmax = x_max,
      ymin = -Inf,
      ymax = Inf
    ),
    alpha = 0.125,
    fill = "#ff0000"
  ) +
  geom_hline(
    yintercept = 0,
    linetype = "22",
    color = "grey20",
    linewidth = 0.2
  )  +
  geom_ribbon(
    data = output_age_rate$predictions$derivative_smooth,
    aes(x = age, ymin = ci_lower, ymax = ci_upper),
    alpha = 0.15
  ) +
  geom_line(
    data = output_age_rate$predictions$derivative_smooth,
    aes(x = age, y = derivative),
    linewidth = 0.6
  ) +
  scale_y_continuous(breaks = c(0, 1, 2)) +
  scale_x_continuous(breaks = seq(20, 80, 10)) +
  coord_cartesian(ylim = c(-0.5, 2.6), xlim = c(20, 82))

p5_a <- p_config$preset(p5_a, "age_rate_der") +
  theme(axis.title.y = element_blank(),
        plot.margin = margin(0, 1, 0, 0, unit = "pt"))

p6 <- ggplot() +
  geom_rect(
    data = pwc_age_rate_boundaries,
    aes(
      xmin = x_min,
      xmax = x_max,
      ymin = -Inf,
      ymax = Inf
    ),
    alpha = 0.125,
    fill = "#ff0000"
  ) +
  geom_hline(
    yintercept = 0,
    linetype = "22",
    color = "grey20",
    linewidth = 0.2
  )  +
  geom_ribbon(
    data = output_age_rate$predictions$derivative_smooth,
    aes(x = age, ymin = ci_lower_adj, ymax = ci_upper_adj),
    alpha = 0.15
  ) +
  geom_line(
    data = output_age_rate$predictions$derivative_smooth,
    aes(x = age, y = derivative_adj),
    linewidth = 0.6
  ) +
  scale_y_continuous(breaks = c(0, 1, 2)) +
  scale_x_continuous(breaks = seq(20, 80, 10)) +
  coord_cartesian(ylim = c(-0.5, 2.6), xlim = c(20, 82))

p6 <- p_config$preset(p6, "age_rate_der") +
  theme(plot.margin = margin(0, 0, 1, 0, unit = "pt"))

p <- align_plots(p0_a, p1_a, p2_a, p3_a, p4_a, p5_a, p6, align = 'hv')

p <- plot_grid(
  plot_grid(
    plot_grid(
      p[[1]],
      NULL,
      p[[2]],
      NULL,
      p[[3]],
      labels = c("A", "", "B", "", "C"),
      nrow = 1,
      rel_widths = c(1, -0.01, 1, -0.01, 1),
      label_x = 0.04,
      label_y = 1.02
    ),
    NULL,
    plot_grid(
      p[[4]],
      NULL,
      p[[5]],
      NULL,
      p[[6]],
      labels = c("D", "", "E", "", "F"),
      nrow = 1,
      rel_widths = c(1, -0.01, 1, -0.01, 1),
      label_x = 0.04,
      label_y = 1.02
    ),
    ncol = 1,
    rel_heights = c(1, -0.15, 1)
  ),
  NULL,
  plot_grid(
    NULL,
    p[[7]],
    NULL,
    ncol = 1,
    rel_heights = c(0.23, 0.54, 0.23),
    labels = c("", "G", ""),
    label_x = 0.04,
    label_y = 1.1
  ),
  ncol = 3,
  rel_widths = c(3, 0.1, 1.005)
)

sameplot(
  p,
  file = here(p_config$base_path, "age_rate_panel.png"),
  width = params$double_column,
  height = 3,
  res = 600,
  save = params$save_figures
) 
```

### Bioage panel

#### Plot Delta mse

```{r plot_bioage, warning=FALSE}
p_config <- load_plot_config(output_type = params$plot_config, params)

p <- df_summary_bioage %>%
  ggplot(aes(x = age_group, y = mean_delta_mse))  +
  geom_hline(
    yintercept = 0,
    linetype = "22",
    color = "grey20",
    linewidth = 0.2
  )  +
  geom_ribbon(
    aes(
      fill = predictors,
      group = predictors,
      ymin = ci_low_delta_mse,
      ymax = ci_high_delta_mse
    ),
    alpha = 0.15,
    show.legend = FALSE
  ) +
  geom_line(
    aes(color = predictors, group = predictors),
    linewidth = 0.7,
    alpha = 0.75,
    show.legend = TRUE
  ) +
  geom_point(
    aes(color = predictors),
    size = 1.25,
    alpha = 1,
    show.legend = TRUE
  ) +
  geom_text(
    data = . %>%
      filter(p_adj != "ns"),
    aes(
      x = age_group,
      y = ci_high_delta_mse + 0.25,
      label = signif,
      color = predictors
    ),
    size = p_config$signif_size / (14 / 5)
  ) +
  scale_y_continuous(breaks = seq(0, 8, 2)) +
  coord_cartesian(ylim = c(-1, 8.5), xlim = c(1.4, 5.6))

p <- p_config$preset(p, "bioage") +
  theme(plot.margin = margin(1, 1, 0, 1, unit = "pt"))

sameplot(
  p,
  file = here(p_config$base_path, "bioage_panel.png"),
  width = params$single_column,
  height = 2.15,
  res = 600,
  save = params$save_figures
) 
```


```{r}
p <- df_summary_bioage %>%
  ggplot(aes(x = age_group, y = mean_delta_mse))  +
  geom_hline(
    yintercept = 0,
    linetype = "22",
    color = "grey20",
    linewidth = 0.2
  )  +
  geom_ribbon(
    aes(
      fill = predictors,
      group = predictors,
      ymin = ci_low_delta_mse,
      ymax = ci_high_delta_mse
    ),
    alpha = 0.15,
    show.legend = FALSE
  ) +
  geom_line(
    aes(color = predictors, group = predictors),
    linewidth = 0.7,
    alpha = 0.75,
    show.legend = TRUE
  ) +
  geom_point(
    aes(color = predictors),
    size = 1.25,
    alpha = 1,
    show.legend = TRUE
  ) +
  geom_text(
    data = . %>%
      filter(p_adj != "ns"),
    aes(
      x = age_group,
      y = ci_high_delta_mse + 0.25,
      label = signif,
      color = predictors
    ),
    size = p_config$signif_size / (14 / 5)
  )

plot(p)
```

### Supplemental Plots

#### Advance bioage plot

```{r }
p_config <- load_plot_config(output_type = params$plot_config, params)

p <- df %>%
  ggplot(aes(x = age, y = phenoage_advance)) +
  geom_density_ridges(aes(x = phenoage_advance, y = age_group)) +
  annotate(
    "text",
    x = Inf,
    y = Inf,
    label = "Decelerated Aging",
    hjust = 2.7,
    vjust = 1,
    size = p_config$base_text_size / (14 / 5),
    color = "darkgreen"
  ) +
  annotate(
    "text",
    x = Inf,
    y = Inf,
    label = "Accelerated Aging",
    hjust = 1,
    vjust = 1,
    size = p_config$base_text_size / (14 / 5),
    color = "darkred"
  ) +
  scale_x_continuous(breaks = seq(-20, 20, 20)) +
  coord_cartesian(xlim = c(-20, 20), ylim = c(1, 7))

p <- p_config$preset(p, "advance_bioage") +
  theme(plot.margin = margin(1, 1, 0, 1, unit = "pt"))

sameplot(
  p,
  file = here(p_config$base_path, "age_acceleration.png"),
  width = params$single_column,
  height = 2.15,
  res = 600,
  save = params$save_figures
)
```

#### Biomarkers supplemental plots

```{r plot_biomarker_all, warning=FALSE}
if (isFALSE(params$telolength)) {
  p_config <- load_plot_config(output_type = params$plot_config, params)
  
  p <- df_summary_biomarkers_overall %>%
    filter(!(
      predictors %in% c(
        "age",
        "age, phenoage",
        "age, crp, hba1c, albumin_gL",
        "age, crp, hba1c, albumin_gL, phenoage"
      )
    )) %>% mutate(predictors = factor(
      predictors,
      levels = c(
        "age, sbp",
        "age, bun",
        "age, mcv",
        "age, lymph",
        "age, rdw",
        "age, totchol",
        "age, creat",
        "age, uap",
        "age, glucose",
        "age, wbc",
        "age, alp",
        "age, albumin_gL",
        "age, hba1c",
        "age, crp"
      ),
      labels = c(
        "SBP",
        "BUN",
        "MCV",
        "Lymphs",
        "RDW",
        "TC",
        "Creat",
        "UA",
        "Glc",
        "WBC",
        "ALP",
        "ALB",
        "HbA1c",
        "CRP"
      )
    )) %>%
    ggplot(aes(x = predictors, y = mean_delta_mse)) +
    geom_point(
      aes(color = predictors, fill = predictors),
      size = 2.25,
      position = position_dodge(0.8),
      alpha = 1,
      show.legend = FALSE
    ) +
    geom_linerange(
      aes(
        color = predictors,
        ymin = ci_low_delta_mse,
        ymax = ci_high_delta_mse
      ),
      position = position_dodge(0.8),
      lineend = "round",
      linewidth = 0.75,
      show.legend = FALSE
    ) +
    geom_text(
      data = . %>%
        filter(p_adj != "ns"),
      aes(
        x = predictors,
        y = ci_high_delta_mse + 0.25,
        label = signif,
        color = predictors
      ),
      manual = TRUE,
      size = p_config$signif_size / (14 / 5),
      show.legend = FALSE
    ) +
    scale_y_continuous(breaks = seq(0, 2, 1)) +
    coord_cartesian(ylim = c(-0.25, 1.5))
  
  p <- p_config$preset(p, "biomarkers") +
    theme(
      plot.margin = margin(1, 1, 0, 1, unit = "pt"),
      axis.title.x = element_blank(),
      axis.text.x = element_markdown(
        "sans",
        size = 8,
        angle = 45,
        hjust = 1,
        vjust = 1
      )
    )
  
  sameplot(
    p,
    file = here(p_config$base_path, "biomarkers_panel.png"),
    width = params$single_column,
    height = 2.15,
    res = 600,
    save = params$save_figures
  )
}
```

### Additional Plots

#### Simplified Rate of Aging

```{r plot_age_rate_simple}
if (isFALSE(params$telolength)) {
  p <- align_plots(p0_a, p1_a, p2_a, p3_a, p4_a, p5_a, align = 'hv')
  
  p <-
    plot_grid(
      plot_grid(
        p[[1]],
        NULL,
        p[[2]],
        NULL,
        p[[3]],
        labels = c("A", "", "B", "", "C"),
        nrow = 1,
        rel_widths = c(1, -0.01, 1, -0.01, 1),
        label_x = 0.035,
        label_y = 1.02
      ),
      NULL,
      plot_grid(
        p[[4]],
        NULL,
        p[[5]],
        NULL,
        p[[6]],
        labels = c("D", "", "E", "", "F"),
        nrow = 1,
        rel_widths = c(1, -0.01, 1, -0.01, 1),
        label_x = 0.035,
        label_y = 1.02
      ),
      ncol = 1,
      rel_heights = c(1, -0.15, 1)
    )
  
  sameplot(
    p,
    file = here(
      p_config$base_path,
      "Extras",
      "simplified_age_rate_panel.png"
    ),
    width = params$double_column * 1.8,
    height = 3 * 1.8,
    scaling = 1.8,
    res = 600,
    save = params$save_figures
  )
}
```

#### Blank rate of aging

```{r plot_blank_age_rate}
if (isFALSE(params$telolength)) {
  p_config <- load_plot_config(output_type = params$plot_config, params)
  
  p0 <- ggplot() +
    geom_point(
      data = df,
      aes(age, .data[[params$hearing_var]]),
      alpha = 0.1,
      size = 0.5,
      shape = 1,
      stroke = 0.25,
      color = "grey30"
    ) +
    scale_y_continuous(breaks = c(0, 30, 60, 90)) +
    scale_x_continuous(breaks = seq(20, 80, 10)) +
    coord_cartesian(ylim = c(-10, 100), xlim = c(20, 82))
  
  p0 <- p_config$preset(p0, "age_rate") +
    theme(plot.margin = margin(1, 0, 0, 1, unit = "pt"))
  
  p1 <- ggplot() +
    geom_point(
      data = df,
      aes(age, .data[[params$hearing_var]]),
      alpha = 0.1,
      size = 0.5,
      shape = 1,
      stroke = 0.25,
      color = "grey30"
    ) +
    scale_y_continuous(breaks = c(0, 30, 60, 90)) +
    scale_x_continuous(breaks = seq(20, 80, 10)) +
    coord_cartesian(ylim = c(-10, 100), xlim = c(20, 82))
  
  p1 <- p_config$preset(p1, "age_rate") +
    theme(plot.margin = margin(1, 0, 0, 1, unit = "pt"),
          axis.title.y = element_blank())
  
  p2 <- ggplot() +
    geom_point(
      data = df,
      aes(age, .data[[params$hearing_var]]),
      alpha = 0.1,
      size = 0.5,
      shape = 1,
      stroke = 0.25,
      color = "grey30"
    ) +
    scale_y_continuous(breaks = c(0, 30, 60, 90)) +
    scale_x_continuous(breaks = seq(20, 80, 10)) +
    coord_cartesian(ylim = c(-10, 100), xlim = c(20, 82))
  
  p2 <- p_config$preset(p2, "age_rate") +
    theme(plot.margin = margin(1, 1, 0, 0, unit = "pt"),
          axis.title.y = element_blank())
  
  p3 <- ggplot() +
    geom_hline(
      yintercept = 0,
      linetype = "22",
      color = "grey20",
      linewidth = 0.2
    )  +
    geom_line(
      data = output_age_rate$predictions$linear_derivative,
      aes(x = age, y = derivative),
      linewidth = 0.6,
      color = "white"
    ) +
    scale_y_continuous(breaks = c(0, 1, 2)) +
    scale_x_continuous(breaks = seq(20, 80, 10)) +
    coord_cartesian(ylim = c(-0.5, 2.6), xlim = c(20, 82))
  
  
  p3 <- p_config$preset(p3, "age_rate_der") +
    theme(plot.margin = margin(1, 0, 0, 1, unit = "pt"))
  
  
  p4 <- ggplot() +
    geom_hline(
      yintercept = 0,
      linetype = "22",
      color = "grey20",
      linewidth = 0.2
    )  +
    geom_line(
      data = output_age_rate$predictions$linear_derivative,
      aes(x = age, y = derivative),
      linewidth = 0.6,
      color = "white"
    ) +
    scale_y_continuous(breaks = c(0, 1, 2)) +
    scale_x_continuous(breaks = seq(20, 80, 10)) +
    coord_cartesian(ylim = c(-0.5, 2.6), xlim = c(20, 82))
  
  p4 <- p_config$preset(p4, "age_rate_der") +
    theme(plot.margin = margin(0, 0, 0, 1, unit = "pt"),
          axis.title.y = element_blank())
  
  p5 <- ggplot() +
    geom_hline(
      yintercept = 0,
      linetype = "22",
      color = "grey20",
      linewidth = 0.2
    )  +
    geom_line(
      data = output_age_rate$predictions$linear_derivative,
      aes(x = age, y = derivative),
      linewidth = 0.6,
      color = "white"
    ) +
    scale_y_continuous(breaks = c(0, 1, 2)) +
    scale_x_continuous(breaks = seq(20, 80, 10)) +
    coord_cartesian(ylim = c(-0.5, 2.6), xlim = c(20, 82))
  
  p5 <- p_config$preset(p5, "age_rate_der") +
    theme(plot.margin = margin(0, 1, 0, 0, unit = "pt"),
          axis.title.y = element_blank())
  
  p <- align_plots(p0, p1, p2, p3, p4, p5, align = 'hv')
  
  p <-
    plot_grid(
      plot_grid(
        p[[1]],
        NULL,
        p[[2]],
        NULL,
        p[[3]],
        labels = c("A", "", "B", "", "C"),
        nrow = 1,
        rel_widths = c(1, -0.01, 1, -0.01, 1),
        label_x = 0.035,
        label_y = 1.02
      ),
      NULL,
      plot_grid(
        p[[4]],
        NULL,
        p[[5]],
        NULL,
        p[[6]],
        labels = c("D", "", "E", "", "F"),
        nrow = 1,
        rel_widths = c(1, -0.01, 1, -0.01, 1),
        label_x = 0.035,
        label_y = 1.02
      ),
      ncol = 1,
      rel_heights = c(1, -0.15, 1)
    )
  
  sameplot(
    p,
    file = here(
      "Output",
      "Figures",
      params$hearing_var,
      "Extras",
      "blank_age_rate_panel.png"
    ),
    width = params$double_column * 1.8,
    height = 3 * 1.8,
    scaling = 1.8,
    res = 600,
    save = params$save_figures
  )
}
```

### Confounders panel

#### Plot Delta mse confounders

```{r plot_confounders, warning=FALSE}
if (isFALSE(params$telolength)) {
    p_config <- load_plot_config(output_type = params$plot_config, params)

p <- df_summary_confounders %>%
  ggplot(aes(x = age_group, y = mean_delta_mse))  +
  geom_hline(
    yintercept = 0,
    linetype = "22",
    color = "grey20",
    linewidth = 0.2
  )  +
  geom_ribbon(
    aes(
      fill = predictors,
      group = predictors,
      ymin = ci_low_delta_mse,
      ymax = ci_high_delta_mse
    ),
    alpha = 0.15,
    show.legend = FALSE
  ) +
  geom_line(
    aes(color = predictors, group = predictors),
    linewidth = 0.7,
    alpha = 0.75,
    show.legend = TRUE
  ) +
  geom_point(
    aes(color = predictors),
    size = 1.25,
    alpha = 1,
    show.legend = TRUE
  ) +
  scale_color_manual(
    values =
      c(
        "full" = "#000000",
        "full_with_phenoage" = "#256DAB",
        "no_education" = "#DF5C24",
        "no_gender" = "#059748",
        "no_gender, race" = "#7B3A96",
        "no_race" = "#E5126F",
        "no_smoking" = "#9D722A"
      ),
    labels = c(
      "Full",
      "Phenoage",
      "Education",
      "Gender",
      "Gender*Race",
      "Race",
      "Smoking"
    )
  ) +
    scale_fill_manual(
    values =
      c(
        "full" = "#8C8C8C",
        "full_with_phenoage" = "#88BDE6",
        "no_education" = "#FBB258",
        "no_gender" = "#90CD97",
        "no_gender, race" = "#BC99C7",
        "no_race" = "#F6AAC9",
        "no_smoking" = "#BFA554"
      ),
    labels = c(
      "Full",
      "Phenoage",
      "Education",
      "Gender",
      "Gender*Race",
      "Race",
      "Smoking"
    )
  ) +
  guides(
    colour = guide_legend(
      override.aes = list(size = 1.25),
      keywidth = unit(0.3, "cm"),
      keyheight = unit(0.3, "cm"),
      label.theme = element_markdown(
        "sans",
        size = 9,
        margin = margin(l = 3, t = 1.5, unit = "pt")
      ),
    )
  ) +
  scale_y_continuous(breaks = seq(0, 40, 5)) +
  coord_cartesian(ylim = c(-1, 40), xlim = c(1.4, 5.6)) +
  labs(x = "Age Group<br><span style='font-size:8pt; color:#4D4D4D'>(Years)</span>", y = "Additional Predictive Capacity<br><span style='font-size:8pt; color:#4D4D4D'>(Δ MSE (dB HL)<sup>2</sup>)</span>") +
  theme_few() +
  theme(
    axis.text.x = element_markdown("sans", size = 8),
    axis.text.y = element_markdown("sans", size = 8),
    axis.title.x = element_markdown("sans", size = 9, lineheight = 1.1),
    axis.title.y = element_markdown("sans", size = 9, lineheight = 1.1),
    legend.title              = element_blank(),
    legend.background         = element_rect(fill = "transparent", color = NA),
    legend.box.background     = element_rect(fill = "transparent", color = NA),
    legend.key                = element_rect(fill = "transparent"),
    legend.position           = "inside",
    legend.justification.inside    = c(0, 1),
    legend.key.spacing.y      = unit(1, "pt"),
    panel.border = element_part_rect(
      colour = "grey30",
      side = "lb",
      linewidth = 0.5
    ),
    plot.margin = margin(
      t = 1,
      r = 1,
      b = 0,
      l = 1,
      unit = "pt"
    )
  ) + geom_text(
    data = . %>%
      filter(p_adj != "ns"),
    aes(
      x = age_group,
      y = ci_high_delta_mse,
      label = signif,
      color = predictors
    ),
    manual = TRUE,
    size = 14 / (14 / 5)
  )

sameplot(
  p,
  file = here(
    "Output",
    "Figures",
    "confounder_panel.png"
  ),
  width = params$single_column,
  height = 2.15,
  res = 600,
  save = params$save_figures
) 
}
```

#### Plot additional variance confounders

```{r plot_confounders_var, warning=FALSE}
if (isFALSE(params$telolength)) {
p <- df_summary_confounders %>%
  ggplot(aes(x = age_group, y = mean_additional_variance))  +
  geom_hline(
    yintercept = 0,
    linetype = "22",
    color = "grey20",
    linewidth = 0.2
  )  +
  geom_ribbon(
    aes(
      fill = predictors,
      group = predictors,
      ymin = ci_low_additional_variance,
      ymax = ci_high_additional_variance
    ),
    alpha = 0.15,
    show.legend = FALSE
  ) +
  geom_line(
    aes(color = predictors, group = predictors),
    linewidth = 0.7,
    alpha = 0.75,
    show.legend = TRUE
  ) +
  geom_point(
    aes(color = predictors),
    size = 1.25,
    alpha = 1,
    show.legend = TRUE
  ) +
  scale_color_manual(
    values =
      c(
        "full" = "#000000",
        "full_with_phenoage" = "#256DAB",
        "no_education" = "#DF5C24",
        "no_gender" = "#059748",
        "no_gender, race" = "#7B3A96",
        "no_race" = "#E5126F",
        "no_smoking" = "#9D722A"
      ),
    labels = c(
      "Full",
      "Phenoage",
      "Education",
      "Gender",
      "Gender*Race",
      "Race",
      "Smoking"
    )
  ) +
    scale_fill_manual(
    values =
      c(
        "full" = "#8C8C8C",
        "full_with_phenoage" = "#88BDE6",
        "no_education" = "#FBB258",
        "no_gender" = "#90CD97",
        "no_gender, race" = "#BC99C7",
        "no_race" = "#F6AAC9",
        "no_smoking" = "#BFA554"
      ),
    labels = c(
      "Full",
      "Phenoage",
      "Education",
      "Gender",
      "Gender*Race",
      "Race",
      "Smoking"
    )
  ) +
  guides(
    colour = guide_legend(
      override.aes = list(size = 1.25),
      keywidth = unit(0.3, "cm"),
      keyheight = unit(0.3, "cm"),
      label.theme = element_markdown(
        "sans",
        size = 9,
        margin = margin(l = 3, t = 1.5, unit = "pt")
      ),
    )
  ) +
  scale_y_continuous(breaks = seq(0, 8, 2)) +
  coord_cartesian(ylim = c(-1, 15), xlim = c(1.4, 5.6)) +
  labs(x = "Age Group<br><span style='font-size:8pt; color:#4D4D4D'>(Years)</span>", y = "Additional Predictive Capacity<br><span style='font-size:8pt; color:#4D4D4D'>(Δ % Variance Explained)</span>") +
  theme_few() +
  theme(
    axis.text.x = element_markdown("sans", size = 8),
    axis.text.y = element_markdown("sans", size = 8),
    axis.title.x = element_markdown("sans", size = 9, lineheight = 1.1),
    axis.title.y = element_markdown("sans", size = 9, lineheight = 1.1),
    legend.title              = element_blank(),
    legend.background         = element_rect(fill = "transparent", color = NA),
    legend.box.background     = element_rect(fill = "transparent", color = NA),
    legend.key                = element_rect(fill = "transparent"),
    legend.position           = "inside",
    legend.justification.inside    = c(0, 1),
    legend.key.spacing.y      = unit(1, "pt"),
    panel.border = element_part_rect(
      colour = "grey30",
      side = "lb",
      linewidth = 0.5
    ),
    plot.margin = margin(
      t = 1,
      r = 1,
      b = 0,
      l = 1,
      unit = "pt"
    )
  ) + geom_text(
    data = . %>%
      filter(p_adj != "ns"),
    aes(
      x = age_group,
      y = ci_high_additional_variance,
      label = signif,
      color = predictors
    ),
    manual = TRUE,
    size = 14 / (14 / 5)
  )

sameplot(
  p,
  file = here(
    "Output",
    "Figures",
    "confounder_panel_var.png"
  ),
  width = params$single_column,
  height = 2.15,
  res = 600,
  save = params$save_figures
) 
}
```

#### Plot PhenoAge Only

```{r plot_phenoage_only, warning=FALSE}
if (isFALSE(params$telolength)) {
p <- df_summary_bioage %>% filter(
  predictors %in% c("age", "age, phenoage")
) %>%
  ggplot(aes(x = age_group, y = mean_delta_mse))  +
  geom_hline(
    yintercept = 0,
    linetype = "22",
    color = "grey20",
    linewidth = 0.2
  )  +
  geom_ribbon(
    aes(
      fill = predictors,
      group = predictors,
      ymin = ci_low_delta_mse,
      ymax = ci_high_delta_mse
    ),
    alpha = 0.15,
    show.legend = FALSE
  ) +
  geom_line(
    aes(color = predictors, group = predictors),
    linewidth = 0.7,
    alpha = 0.75,
    show.legend = TRUE
  ) +
  geom_point(
    aes(color = predictors),
    size = 1.25,
    alpha = 1,
    show.legend = TRUE
  ) +
  scale_color_manual(
    values = c(
      "age" = "#000000",
      "age, phenoage" = "#CB2027"
    ),
    labels = c("Age", "Age + PhenoAge")
  ) +
  scale_fill_manual(
    values = c(
      "age" = "#8C8C8C",
      "age, phenoage" = "#F07E6E"
    ),
    labels = c("Age","Age + PhenoAge")
  ) +
  guides(
    colour = guide_legend(
      override.aes = list(size = 1.25),
      keywidth = unit(0.3, "cm"),
      keyheight = unit(0.3, "cm"),
      label.theme = element_markdown(
        "sans",
        size = 9,
        margin = margin(l = 3, t = 1.5, unit = "pt")
      ),
    )
  ) +
  scale_y_continuous(breaks = seq(0, 8, 2)) +
  coord_cartesian(ylim = c(-1, 8.5), xlim = c(1.4, 5.6)) +
  labs(x = "Age Group<br><span style='font-size:8pt; color:#4D4D4D'>(Years)</span>", y = "Additional Predictive Capacity<br><span style='font-size:8pt; color:#4D4D4D'>(Δ MSE (dB HL)<sup>2</sup>)</span>") +
  theme_few() +
  theme(
    axis.text.x = element_markdown("sans", size = 8),
    axis.text.y = element_markdown("sans", size = 8),
    axis.title.x = element_markdown("sans", size = 9, lineheight = 1.1),
    axis.title.y = element_markdown("sans", size = 9, lineheight = 1.1),
    legend.title              = element_blank(),
    legend.background         = element_rect(fill = "transparent", color = NA),
    legend.box.background     = element_rect(fill = "transparent", color = NA),
    legend.key                = element_rect(fill = "transparent"),
    legend.position           = "inside",
    legend.position.inside    = c(0.215, 0.82),
    legend.key.spacing.y      = unit(1, "pt"),
    panel.border = element_part_rect(
      colour = "grey30",
      side = "lb",
      linewidth = 0.5
    ),
    plot.margin = margin(
      t = 1,
      r = 1,
      b = 0,
      l = 1,
      unit = "pt"
    )
  ) + geom_text(
    data = . %>%
      filter(p_adj != "ns"),
    aes(
      x = age_group,
      y = ci_high_delta_mse + 0.25,
      label = signif,
      color = predictors
    ),
    manual = TRUE,
    size = 14 / (14 / 5)
  )

sameplot(
  p,
  file = here("Output", "Figures", params$hearing_var, "Extras", "phenoage_only_bioage_panel.png"),
  width = params$single_column * 1.8,
  height = 2.15 * 1.8,
  scaling = 1.8,
  res = 600,
  save = params$save_figures
) 
}
```

#### Rate of aging gender

##### Calculate gender GAM

```{r calc_gender_GAM}
if (isFALSE(params$telolength)) {
gam_gender <- bam(
  as.formula(paste0(params$hearing_var, "~ s(age, k = 10) + s(gender, age, bs = 'sz', k = 10)")),
  data = df,
  method = "REML",
  family = gaussian()
)

gender_predictions <- as.data.frame(confint(
    emmeans(
      gam_gender,
      as.formula(paste("~ age + gender")),
      at = setNames(list(seq(
        min(df[["age"]]), max(df[["age"]]), by = 1
      )), "age"),
      type = "response",
      data = df
    ),
    adjust = "mvt"
  )) %>%
    rename("estimate" = emmean,
           "ci_lower" = lower.CL,
           "ci_upper" = upper.CL)    

gender_predictions_der <-  data.frame(confint(
    emtrends(
      gam_gender,
      as.formula(paste("~ age + gender")),
      var = "age",
      at = setNames(list(seq(
        min(df[["age"]]), max(df[["age"]]), by = 1
      )), "age"),
      data = df
    ),
    adjust = "mvt"
  )) %>%
    rename(
      "derivative" = !!paste0("age", ".trend"),
      "ci_lower" = lower.CL,
      "ci_upper" = upper.CL
    )
}
```

##### Plot rate of aging gender

```{r plot_age_rate_gender}
if (isFALSE(params$telolength)) {
p0 <- ggplot() +
  geom_point(
    data = df,
    aes(age, .data[[params$hearing_var]]),
    alpha = 0.1,
    size = 0.5,
    shape = 1,
    stroke = 0.25,
    color = "grey30"
  ) +
  geom_line(
    data = output_age_rate$predictions$linear,
    aes(x = age, y = estimate),
    linewidth = 0.6,
    alpha = 1
  ) +
      geom_ribbon(
    data = output_age_rate$predictions$linear,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    ),
    alpha = 0.15
  ) +
  scale_y_continuous(breaks = c(0, 30, 60, 90)) +
  scale_x_continuous(breaks = seq(20, 80, 10)) +
  coord_cartesian(ylim = c(-10, 100), xlim = c(20, 82)) +
  labs(y = "High Fletcher Index<br><span style='font-size:8pt; color:#4D4D4D'>(dB HL)</span>") +
  theme_few() +
  theme(
    axis.text.x = element_markdown("sans", size = 8),
    axis.text.y = element_markdown("sans", size = 8),
    axis.title.x = element_blank(),
    axis.title.y = element_markdown("sans", size = 9, lineheight = 1.1),
    panel.border = element_part_rect(
      colour = "grey30",
      side = "lb",
      linewidth = 0.5
    ),
    plot.margin = margin(
      t = 1,
      r = 0,
      b = 0,
      l = 1,
      unit = "pt"
    )
  )

p1 <- ggplot() +
  geom_point(
    data = df,
    aes(age, .data[[params$hearing_var]]),
    alpha = 0.1,
    size = 0.5,
    shape = 1,
    stroke = 0.25,
    color = "grey30"
  ) +
  geom_line(
    data = output_age_rate$predictions$quadratic,
    aes(x = age, y = estimate),
    linewidth = 0.6,
    alpha = 1
  ) +
      geom_ribbon(
    data = output_age_rate$predictions$quadratic,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    ),
    alpha = 0.15
  ) +
  scale_y_continuous(breaks = c(0, 30, 60, 90)) +
  scale_x_continuous(breaks = seq(20, 80, 10)) +
  coord_cartesian(ylim = c(-10, 100), xlim = c(20, 82)) +
  theme_few() +
  theme(
    axis.text.x = element_markdown("sans", size = 8),
    axis.text.y = element_markdown("sans", size = 8),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_part_rect(
      colour = "grey30",
      side = "lb",
      linewidth = 0.5
    ),
    plot.margin = margin(
      t = 1,
      r = 0,
      b = 0,
      l = 1,
      unit = "pt"
    )
  )

p2 <- ggplot() +
  geom_point(
    data = df,
    aes(age, .data[[params$hearing_var]]),
    alpha = 0.1,
    size = 0.5,
    shape = 1,
    stroke = 0.25,
    color = "grey30"
  ) +
  geom_ribbon(
    data = gender_predictions,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper,
      fill = gender
    ),
    alpha = 0.15,
    show.legend = FALSE
  ) +
  geom_line(
    data = gender_predictions,
    aes(
      x = age,
      y = estimate,
      color = gender
    ),
    linewidth = 0.6,
    show.legend = FALSE
  ) +
  scale_color_manual(values = c("1" = "#2986cc", "2" = "#c90076")) +
  scale_fill_manual(values = c("1" = "#6fa8dc", "2" = "#d5a6bd")) +
  scale_y_continuous(breaks = c(0, 30, 60, 90)) +
  scale_x_continuous(breaks = seq(20, 80, 10)) +
  coord_cartesian(ylim = c(-10, 100), xlim = c(20, 82)) +
  theme_few() +
  theme(
    axis.text.x = element_markdown("sans", size = 8),
    axis.text.y = element_markdown("sans", size = 8),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_part_rect(
      colour = "grey30",
      side = "lb",
      linewidth = 0.5
    ),
    plot.margin = margin(
      t = 1,
      r = 1,
      b = 0,
      l = 0,
      unit = "pt"
    )
  )

p3 <- ggplot() +
  geom_hline(
    yintercept = 0,
    linetype = "22",
    color = "grey20",
    linewidth = 0.2
  )  +
  geom_line(
    data = output_age_rate$predictions$linear_derivative,
    aes(
      x = age,
      y = derivative
    ),
    linewidth = 0.6
  ) +
    geom_ribbon(
    data = output_age_rate$predictions$linear_derivative,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    ),
    alpha = 0.15
  ) +
  scale_y_continuous(breaks = c(0, 1, 2)) +
  scale_x_continuous(breaks = seq(20, 80, 10)) +
  coord_cartesian(ylim = c(-0.5, 2.6), xlim = c(20, 82)) +
  labs(x = "Age<br><span style='font-size:8pt; color:#4D4D4D'>(Years)</span>", y = "Rate of HL<br><span style='font-size:8pt; color:#4D4D4D'>(dB HL/Year)</span>") +
  theme_few() +
theme(
    axis.text.x = element_markdown("sans", size = 8),
    axis.text.y = element_markdown("sans", size = 8),
    axis.title.x = element_markdown("sans", size = 9, lineheight = 1.1),
    axis.title.y = element_markdown("sans", size = 9, lineheight = 1.1),
    panel.border = element_part_rect(
      colour = "grey30",
      side = "lb",
      linewidth = 0.5
    ),
    plot.margin = margin(
      t = 0,
      r = 0,
      b = 0,
      l = 1,
      unit = "pt"
    )
  )

p4 <- ggplot() +
  geom_hline(
    yintercept = 0,
    linetype = "22",
    color = "grey20",
    linewidth = 0.2
  )  +
  geom_line(
    data = output_age_rate$predictions$quadratic_derivative,
    aes(
      x = age,
      y = derivative
    ),
    linewidth = 0.6
  ) +
    geom_ribbon(
    data = output_age_rate$predictions$quadratic_derivative,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper
    ),
    alpha = 0.15
  ) +
  scale_y_continuous(breaks = c(0, 1, 2)) +
  scale_x_continuous(breaks = seq(20, 80, 10)) +
  coord_cartesian(ylim = c(-0.5, 2.6), xlim = c(20, 82)) +
  labs(x = "Age<br><span style='font-size:8pt; color:#4D4D4D'>(Years)</span>", y = "Rate of HL<br><span style='font-size:8pt; color:#4D4D4D'>(dB HL/Year)</span>") +
  theme_few() +
theme(
    axis.text.x = element_markdown("sans", size = 8),
    axis.text.y = element_markdown("sans", size = 8),
    axis.title.x = element_markdown("sans", size = 9, lineheight = 1.1),
    axis.title.y = element_blank(),
    panel.border = element_part_rect(
      colour = "grey30",
      side = "lb",
      linewidth = 0.5
    ),
    plot.margin = margin(
      t = 0,
      r = 0,
      b = 0,
      l = 1,
      unit = "pt"
    )
  )

p5 <- ggplot() +
  geom_hline(
    yintercept = 0,
    linetype = "22",
    color = "grey20",
    linewidth = 0.2
  )  +
  geom_ribbon(
    data = gender_predictions_der,
    aes(
      x = age,
      ymin = ci_lower,
      ymax = ci_upper,
      fill = gender
    ),
    alpha = 0.15,
    show.legend = FALSE
  ) +
  geom_line(
    data = gender_predictions_der,
    aes(
      x = age,
      y = derivative,
      color = gender
    ),
    linewidth = 0.6,
    show.legend = FALSE
  ) +
  scale_color_manual(values = c("1" = "#2986cc", "2" = "#c90076")) +
  scale_fill_manual(values = c("1" = "#6fa8dc", "2" = "#d5a6bd")) +
  scale_y_continuous(breaks = c(0, 1, 2)) +
  scale_x_continuous(breaks = seq(20, 80, 10)) +
  coord_cartesian(ylim = c(-0.5, 2.6), xlim = c(20, 82)) +
  labs(x = "Age<br><span style='font-size:8pt; color:#4D4D4D'>(Years)</span>", y = "Nonlinear Deviation Rate of HL<br><span style='font-size:8pt; color:#4D4D4D'>(dB HL/Year)</span>") +
  theme_few() +
theme(
    axis.text.x = element_markdown("sans", size = 8),
    axis.text.y = element_markdown("sans", size = 8),
    axis.title.x = element_markdown("sans", size = 9, lineheight = 1.1),
    axis.title.y = element_blank(),
    panel.border = element_part_rect(
      colour = "grey30",
      side = "lb",
      linewidth = 0.5
    ),
    plot.margin = margin(
      t = 0,
      r = 1,
      b = 0,
      l = 0,
      unit = "pt"
    )
  )

p <- align_plots(p0, p1, p2, p3, p4, p5, align = 'hv')

p <- 
  plot_grid(
    plot_grid(p[[1]], NULL, p[[2]], NULL, p[[3]],
      labels = c("A", "", "B", "", "C"),
      nrow = 1, rel_widths = c(1, -0.01, 1, -0.01, 1),
      label_x = 0.04, label_y = 1.02
    ),
    NULL,
    plot_grid(p[[4]], NULL, p[[5]], NULL, p[[6]],
      labels = c("D", "", "E", "", "F"),
      nrow = 1, rel_widths = c(1, -0.01, 1, -0.01, 1),
      label_x = 0.04, label_y = 1.02
    ),
    ncol = 1, rel_heights = c(1, -0.15, 1)
  )

sameplot(
  p,
  file = here("Output", "Figures", params$hearing_var, "Extras", "simplified_age_rate_panel_gender.png"),
  width = params$double_column * 1.8,
  height = 3 * 1.8,
  scaling = 1.8,
  res = 600,
  save = params$save_figures
) 
}
```

#### Plot aging means with and without confounders

##### Plot aging mean without confounders

```{r plot_aging_mean}
if (isFALSE(params$telolength)) {
p0 <- ggplot() +
  geom_point(
    data = df,
    aes(age, .data[[params$hearing_var]]),
    alpha = 0.1,
    size = 0.5,
    shape = 1,
    stroke = 0.25,
    color = "grey30"
  ) +
  geom_line(
    data = output_age_rate$predictions$smooth,
    aes(x = age, y = estimate),
    linewidth = 0.6,
    alpha = 1,
    color = "grey80"
  ) +
    geom_line(
    data = df %>%
  group_by(age) %>%
  summarize(
    hearing_var = mean(.data[[params$hearing_var]]),
    .groups = "drop"
  ),
    aes(x = age, y = hearing_var),
    linewidth = 0.6,
    alpha = 1
  ) +
  scale_y_continuous(breaks = c(0, 30, 60, 90)) +
  scale_x_continuous(breaks = seq(20, 80, 10)) +
  coord_cartesian(ylim = c(-10, 100), xlim = c(20, 82)) +
  labs(y = "High Fletcher Index<br><span style='font-size:8pt; color:#4D4D4D'>(dB HL)</span>") +
  theme_few() +
  theme(
    axis.text.x = element_markdown("sans", size = 8),
    axis.text.y = element_markdown("sans", size = 8),
    axis.title.x = element_markdown("sans", size = 9, lineheight = 1.1),
    axis.title.y = element_markdown("sans", size = 9, lineheight = 1.1),
    panel.border = element_part_rect(
      colour = "grey30",
      side = "lb",
      linewidth = 0.5
    ),
    plot.margin = margin(
      t = 1,
      r = 0,
      b = 0,
      l = 1,
      unit = "pt"
    )
  )

sameplot(
  p0,
  file = here("Output", "Figures", params$hearing_var, "Extras", "age_rate_mean.png"),
  width = params$single_column,
  height = 3,
  res = 600,
  save = params$save_figures
) 
}
```

##### Plot aging mean per gender

```{r plot_gender_aging_mean}
if (isFALSE(params$telolength)) {
p0 <- ggplot() +
  geom_point(
    data = df,
    aes(age, .data[[params$hearing_var]]),
    alpha = 0.1,
    size = 0.5,
    shape = 1,
    stroke = 0.25,
    color = "grey30"
  ) +
  geom_line(
    data = gender_predictions,
    aes(
      x = age,
      y = estimate,
      color = gender
    ),
    linewidth = 0.6,
    show.legend = FALSE
  ) +
    geom_line(
    data = df %>%
  group_by(age, gender) %>%
  summarize(
    hearing_var = mean(.data[[params$hearing_var]]),
    .groups = "drop"
  ),
    aes(x = age, y = hearing_var, color = gender),
    linewidth = 0.6,
    alpha = 1
  ) +
  scale_y_continuous(breaks = c(0, 30, 60, 90)) +
  scale_x_continuous(breaks = seq(20, 80, 10)) +
  coord_cartesian(ylim = c(-10, 100), xlim = c(20, 82)) +
    scale_color_manual(values = c("1" = "#2986cc", "2" = "#c90076"), labels = c("1" = "Male",
                                                                                "2" = "Female")) +
  scale_fill_manual(values = c("1" = "#6fa8dc", "2" = "#d5a6bd"), labels = c("1" = "Male",
                                                                                "2" = "Female")) +
  labs(y = "High Fletcher Index<br><span style='font-size:8pt; color:#4D4D4D'>(dB HL)</span>") +
  theme_few() +
  theme(
    axis.text.x = element_markdown("sans", size = 8),
    axis.text.y = element_markdown("sans", size = 8),
    axis.title.x = element_markdown("sans", size = 9, lineheight = 1.1),
    axis.title.y = element_markdown("sans", size = 9, lineheight = 1.1),
    panel.border = element_part_rect(
      colour = "grey30",
      side = "lb",
      linewidth = 0.5
    ),
    plot.margin = margin(
      t = 1,
      r = 0,
      b = 0,
      l = 1,
      unit = "pt"
    )
  )

sameplot(
  p0,
  file = here("Output", "Figures", params$hearing_var, "Extras", "age_rate_mean.png"),
  width = params$single_column,
  height = 3,
  res = 600,
  save = params$save_figures
) 
}
```

##### Plot aging mean per race

```{r plot_race_aging_mean}
if (isFALSE(params$telolength)) {
p0 <- ggplot() +
  geom_point(
    data = df,
    aes(age, .data[[params$hearing_var]]),
    alpha = 0.1,
    size = 0.5,
    shape = 1,
    stroke = 0.25,
    color = "grey30"
  ) +
    geom_line(
    data = df %>%
  group_by(age, race) %>%
  summarize(
    hearing_var = mean(.data[[params$hearing_var]]),
    .groups = "drop"
  ),
    aes(x = age, y = hearing_var, color = race),
  linewidth = 0.6, alpha = 1
    ) +
  scale_y_continuous(breaks = c(0, 30, 60, 90)) +
  scale_x_continuous(breaks = seq(20, 80, 10)) +
  coord_cartesian(ylim = c(-10, 100), xlim = c(20, 82)) +
  labs(y = "High Fletcher Index<br><span style='font-size:8pt; color:#4D4D4D'>(dB HL)</span>") +
  theme_few() +
  theme(
    axis.text.x = element_markdown("sans", size = 8),
    axis.text.y = element_markdown("sans", size = 8),
    axis.title.x = element_markdown("sans", size = 9, lineheight = 1.1),
    axis.title.y = element_markdown("sans", size = 9, lineheight = 1.1),
    panel.border = element_part_rect(
      colour = "grey30",
      side = "lb",
      linewidth = 0.5
    ),
    plot.margin = margin(
      t = 1,
      r = 0,
      b = 0,
      l = 1,
      unit = "pt"
    )
  )

sameplot(
  p0,
  file = here("Output", "Figures", params$hearing_var, "Extras", "age_rate_mean_race.png"),
  width = params$single_column,
  height = 3,
  res = 600,
  save = TRUE
) 
}
```

#### Biomarkers

##### Plot Delta mse Biomarkers (only relevant ones)

```{r plot_biomarkers}
if (isFALSE(params$telolength)) {
p1 <- df_summary_biomarkers %>%
  filter(predictors %in% c("age, crp", "age, hba1c", "age, albumin_gL")) %>%
  mutate(
    label = predictors,
    label = case_when(
      label == "age, crp" ~ "Age + CRP",
      label == "age, hba1c" ~ "Age + HbA1c",
      label == "age, albumin_gL" ~ "Age + ALB",
      TRUE ~ NA
    ),
    predictors = factor(
      predictors,
      levels = c("age, crp", "age, hba1c", "age, albumin_gL")
    )
  ) %>%
  ggplot(aes(x = age_group, y = mean_delta_mse))  +
  geom_hline(
    yintercept = 0,
    linetype = "22",
    color = "grey20",
    linewidth = 0.2
  )  +
  geom_label(
    data = . %>%
      filter(age_group == "20-30"),
    aes(
      x = "20-30",
      y = 3.65,
      label = label,
      fill = predictors
    ),
    size = 8 / (14 / 5),
    hjust = 0,
    label.padding = unit(0.3, "lines"),
    nudge_x = -0.1,
    color = "white",
    show.legend = FALSE
  ) +
  geom_ribbon(
    aes(
      fill = predictors,
      group = predictors,
      ymin = ci_low_delta_mse,
      ymax = ci_high_delta_mse
    ),
    alpha = 0.2,
    show.legend = FALSE
  ) +
  geom_line(
    aes(color = predictors, group = predictors),
    linewidth = 0.7,
    alpha = 0.75,
    show.legend = FALSE
  ) +
  geom_point(
    aes(color = predictors),
    size = 1.25,
    alpha = 1,
    show.legend = FALSE
  ) +
  scale_color_manual(
    values = c(
      "age, crp" = "#256DAB",
      "age, hba1c" = "#E5126F",
      "age, albumin_gL" = "#DF5C24"
    )
  ) +
  scale_fill_manual(
    values = c(
      "age, crp" = "#88BDE6",
      "age, hba1c" = "#F6AAC9",
      "age, albumin_gL" = "#FBB258"
    )
  ) +
  facet_wrap(
    ~ predictors,
    ncol = 1,
    nrow = 5,
    labeller = labeller(
      predictors = c(
        "age, crp" = "Age + CRP",
        # creatinine in mg/dL
        "age, hba1c" = "Age + HbA1c",
        # plasma uric acid in mg/dL
        "age, albumin_gL" = "Age + ALB"
      )
    )
  ) + # in billions per liter
  labs(x = "Age Group<br><span style='font-size:8pt; color:#4D4D4D'>(Years)</span>", y = "Additional Predictive Capacity<br><span style='font-size:8pt; color:#4D4D4D'>(Δ MSE (dB HL)<sup>2</sup>)</span>") +
  scale_y_continuous(breaks = c(0, 2)) +
  coord_cartesian(ylim = c(-1.5, 4), xlim = c(1.4, 5.6)) +
  theme_few() +
  theme(
    axis.text.x = element_markdown("sans", size = 8),
    axis.text.y = element_markdown("sans", size = 8),
    axis.title.x = element_markdown("sans", size = 9, lineheight = 1.1),
    axis.title.y = element_markdown("sans", size = 9, lineheight = 1.1),
    strip.text = element_blank(),
    legend.title              = element_blank(),
    legend.background         = element_rect(fill = "transparent", color = NA),
    legend.box.background     = element_rect(fill = "transparent", color = NA),
    legend.key                = element_rect(fill = "transparent"),
    legend.position           = "inside",
    legend.position.inside    = c(0.18, 0.83),
    legend.key.spacing.y      = unit(1, "pt"),
    panel.border = element_part_rect(
      colour = "grey30",
      side = "lb",
      linewidth = 0.5
    ),
    plot.margin = margin(
      t = 1,
      r = 0,
      b = 0,
      l = 1,
      unit = "pt"
    )
  )

sameplot(
  p1,
  file = here("Output", "Figures", params$hearing_var, "biomarkers_biomarkers_panel.png"),
  width = 2.5,
  height = 3,
  res = 600,
  save = params$save_figures
)
}
```

##### Plot biomarkers together vs biomarkers + phenoage

```{r plot_explained_var}
if (isFALSE(params$telolength)) {
p2 <- df_summary_biomarkers %>% filter(
  predictors %in% c(
    "age, phenoage",
    "age, crp, hba1c, albumin_gL",
    "age, crp, hba1c, albumin_gL, phenoage"
  )
) %>%
  mutate(predictors = factor(
    predictors,
    levels = c(
    "age, phenoage",
    "age, crp, hba1c, albumin_gL",
    "age, crp, hba1c, albumin_gL, phenoage"
    )
  )) %>%
  ggplot(aes(x = age_group, y = mean_delta_mse))  +
  geom_hline(
    yintercept = 0,
    linetype = "22",
    color = "grey20",
    linewidth = 0.2
  )  +
  geom_ribbon(
    aes(
      fill = predictors,
      group = predictors,
      ymin = ci_low_delta_mse,
      ymax = ci_high_delta_mse
    ),
    alpha = 0.2,
    show.legend = FALSE
  ) +
  geom_line(
    aes(color = predictors, group = predictors),
    linewidth = 0.7,
    alpha = 0.75,
    show.legend = TRUE
  ) +
  geom_point(
    aes(color = predictors),
    size = 1.25,
    alpha = 1,
    show.legend = FALSE
  ) +
  scale_color_manual(
    values = c(
      "age, crp, hba1c, albumin_gL, phenoage" = "#CB2027",
      "age, crp, hba1c, albumin_gL" = "#059748",
      "age, phenoage" = "grey65"
    ),
    labels = c(
      "Age + CRP + HbA1c + ALB + PhenoAge",
      "Age + CRP + HbA1c + ALB",
      "Age + PhenoAge"
    ),
    breaks = c(
      "age, crp, hba1c, albumin_gL, phenoage",
      "age, crp, hba1c, albumin_gL",
      "age, phenoage"
    )
  ) +
  scale_fill_manual(
    values = c(
      "age, crp, hba1c, albumin_gL, phenoage" = "#F07E6E",
      "age, crp, hba1c, albumin_gL" = "#90CD97",
      "age, phenoage" = "grey80"
    ),
    labels = c(
      "Age + CRP + HbA1c + ALB + PhenoAge",
      "Age + CRP + HbA1c + ALB",
      "Age + PhenoAge"
    ),
    breaks = c(
      "age, crp, hba1c, albumin_gL, phenoage",
      "age, crp, hba1c, albumin_gL",
      "age, phenoage"
    )
    
  ) +
  guides(
    colour = guide_legend(
      override.aes = list(size = 1.25),
      keywidth = unit(0.3, "cm"),
      keyheight = unit(0.3, "cm"),
      label.theme = element_markdown(
        "sans",
        size = 8,
        margin = margin(l = 3, t = 1.5, unit = "pt")
      ),
    )
  ) +
  scale_y_continuous(breaks = c(0, 2, 4, 6, 8)) +
  coord_cartesian(ylim = c(-3, 8.5), xlim = c(1.4, 5.6)) +
  labs(x = "Age Group<br><span style='font-size:8pt; color:#4D4D4D'>(Years)</span>") +
  theme_few() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.x = element_markdown("sans", size = 8),
    axis.text.y = element_markdown("sans", size = 8),
    axis.title.x = element_markdown("sans", size = 9, lineheight = 1.1),
    axis.title.y = element_blank(),
    legend.title              = element_blank(),
    legend.background         = element_rect(fill = "transparent", color = NA),
    legend.box.background     = element_rect(fill = "transparent", color = NA),
    legend.key                = element_rect(fill = "transparent"),
    legend.position           = "inside",
    legend.position.inside    = c(0.3, 0.91),
    legend.key.spacing.y      = unit(1, "pt"),
    strip.text.x              = element_markdown("sans", size = 9, lineheight = 1.1),
    panel.border = element_part_rect(
      colour = "grey30",
      side = "lb",
      linewidth = 0.5
    ),
    plot.margin = margin(
      t = 1,
      r = 1,
      b = 0,
      l = 0,
      unit = "pt"
    )
  )

sameplot(
  p2,
  file = here("Output", "Figures", params$hearing_var, "biomarkers_v_phenoage.png"),
  width = 3,
  height = 4,
  res = 600,
  save = params$save_figures
) 
}
```

##### Composite plot

```{r plot_biomark_panel}
if (isFALSE(params$telolength)) {
p <- ggarrange(
  p1,
  NULL,
  p2,
  nrow = 1,
  widths = c(0.6, 0, 1),
  labels = c("A", "", "B"),
  label.x = 0.02,
  label.y = 1.01,
  align = "v"
)

sameplot(
   p,
  file = here("Output", "Figures", params$hearing_var, "biomarkers_fig_panel.png"),
  width = params$double_column,
  height = 4,
  res = 600,
  save = params$save_figures
) 
}
```
