---
title: "Predictive Modelling of Hearing Effect of Bioage"
author: "Landy Qu & Tom Naber"
date: "`r format(Sys.Date(), '%B %d, %Y')`"

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, params = "ask", output_file = here::here("Output", "HTML", "NHANES_FI_high_Telolength_Analysis.html"))})
  
output:
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    highlight: tango
    self_contained: true
    fig_width: 7
    fig_height: 4
    code_folding: hide
    
editor_options:
  markdown:
    wrap: 80
  chunk_output_type: inline
  
params:
  save_figures:
    label: "Save figures"
    value: FALSE
  construct_data:
    label: "Build NHANES dataframe from scratch"
    value: FALSE
  use_kwon_bioage:
    label: "Use bioages calculated by Kwon"
    value: FALSE
  use_advance_bioages:
    label: "Use difference between chronological and biological age only"
    value: FALSE
  global_k:
    label: "K (n BAM base functions)"
    value: 50
    input: numeric
  global_discrete: 
    label: "Discretize BAM (much faster but slightly worse fitting)"
    value: TRUE
  single_column_width:
    label: "Set width of single column figure (inches)"
    value: 3.19
    input: numeric
  double_column_width:
    label: "Set width of double column figure (inches)"
    value: 6.96
    input: numeric    
---

# Set-up & Data import

## Import Packages & Set Flags

```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(fig.width = 6.4,
                      fig.height = 4.8,
                      dpi = 600,
                      fig.retina = 1,
                      out.width = "100%",
                      out.height = "100%",
                      dev = "ragg_png")

knitr::opts_knit$set(root.dir = here::here())

options(contrasts = c("contr.sum", "contr.poly"))

library(mgcv) # used to train model
library(MASS)
library(tidyverse)
library(ggpubr)
library(ggthemes)
library(ggtext)
library(kableExtra)
library(splines)
library(ggh4x)
library(afex)
library(emmeans)
library(car)
library(correctR)
library(flextable)
library(officer)
library(matrixStats) # used for weighted variance
library(gratia) # used for derivative coefficients
library(cowplot)
library(data.table)
library(ragg)
library(here)
library(sameplot)

source(here("Scripts", "universal_helpers.R"))
```

## Print params for .md

```{r}
kable(as.data.frame(unclass(params)))
```

## Import Data

```{r}
import_nhanes_data <- function(path) {
  path
  files <- list.files(
    path,
    pattern = "\\.xpt$",
    full.names = TRUE,
    ignore.case = TRUE
  )
  
  meta <- data.table(file  = files,
                     year = str_extract(tools::file_path_sans_ext(basename(files)), "\\d{4}-\\d{4}$"))
  
  meta[, dt := lapply(file, \(f) as.data.table(read_xpt(f)))]
  by_year <- split(meta$dt, meta$year)
  
  as.data.frame(rbindlist(
    lapply(by_year, \(dts)
           Reduce(\(x, y) merge(
             x, y, by = "SEQN", all = TRUE
           ), dts)),
    fill = TRUE,
    use.names = TRUE,
    idcol = "year"
  ))
}

if (isTRUE(params$construct_data)) {
  df_raw <- import_nhanes_data(here("Data", "Raw data"))
} else {
  df_raw <- read.csv(here("Data", "dataset_bioage.csv"))
}
```

### Insert Kwon data

If `use_kwon_bioage` is `TRUE`, load NHANES4 (cloned from Kwon's GitHub), ensure
sampleIDs that are in BOTH Kwon's data and in the raw data are retained, then
select only the columns in NHANES4 and remove bioage measures if they're in the
raw_data file (either composed from scratch or from Landy).

```{r}
load_kwon_data <- function() {
  library(BioAge)
  
  #HD using NHANES (separate training for men and women)
  hd = hd_nhanes(
    biomarkers = c(
      "albumin",
      "alp",
      "lncrp",
      "totchol",
      "lncreat",
      "hba1c",
      "sbp",
      "bun",
      "uap",
      "lymph",
      "mcv",
      "wbc"
    )
  )
  
  #KDM bioage using NHANES (separate training for men and women)
  kdm = kdm_nhanes(
    biomarkers = c(
      "albumin",
      "alp",
      "lncrp",
      "totchol",
      "lncreat",
      "hba1c",
      "sbp",
      "bun",
      "uap",
      "lymph",
      "mcv",
      "wbc"
    )
  )
  
  #phenoage using NHANES
  phenoage = phenoage_nhanes(
    biomarkers = c(
      "albumin_gL",
      "alp",
      "lncrp",
      "totchol",
      "lncreat_umol",
      "hba1c",
      "sbp",
      "bun",
      "uap",
      "lymph",
      "mcv",
      "wbc"
    )
  )
  
  merge(hd$data, kdm$data) %>% merge(., phenoage$data)
}

if (isTRUE(params$use_kwon_bioage)) {
  NHANES4 <- load_kwon_data()
  df_raw <- df_raw %>%
    select(-any_of(c("hd", "hd_log", "kdm", "phenoage")))
  
  df_raw <- df_raw %>% filter(sampleID %in% NHANES4$sampleID) %>% arrange(sampleID) %>% relocate(sampleID, .before = 1)
  NHANES4 <- NHANES4 %>% filter(sampleID %in% df_raw$sampleID) %>% arrange(sampleID)
  
  
  
  df_raw <- df_raw %>%
    left_join(NHANES4 %>% select(any_of(c(
      "sampleID", setdiff(names(NHANES4), names(df_raw))
    ))), by = "sampleID")
}
```

# Use advance versions

```{r}
if (isTRUE(params$use_kwon_bioage) && isTRUE(params$use_advance_bioages)) {
  df_raw <- df_raw %>%
    rename("phenoage_bak" = phenoage,
           "kdm_bak" = kdm) %>%
    mutate(phenoage = phenoage_advance,
           kdm = kdm_advance)
} else if (isFALSE(params$use_kwon_bioage) && isTRUE(params$use_advance_bioages)) {
  stop("Set use_kwon_bioage to TRUE when using use_advance_bioages")
}
```

# Data Preprocessing

## Remove Unnecessary Columns & Rename

We filter out education level 9 because there are only 2 occurences. Smoking has
more than 2 levels but we only need two. Race 1 (Mexican American) and 2 (other
hispanic) are merged. Income could be added but we'd have to exclude 12, 13, 77,
99 and missing. 12 is over 20k, 13 is under 20k, 77 is refused and 99 is don't
know. This would drop the size of the dataset by about 20%.

Code levels: **Smoking** 1. Smoker (L1) 2. Non-smoker L2) 7. Refused 9. Don't
know **Gender** 1. Male (L1) 2. Female (L2) **Race** 1. Mexican/other hispanic
(L1) 3. White (L2) 4. Black (L3) 5. Other (L4) **Education** 1. Less than 9th
grade (L1) 2. 9-11th grade (includes 12th grade with no diploma) (L2) 3. High
school (L3) 4. Some college or AA degree (L4) 5. College graduate or above 7.
Refused 9. Don't know

```{r format_df}
df <- df_raw %>%
  select(
    age,
    smoking,
    gender,
    race,
    education,
    income,
    kdm,
    phenoage,
    hd_log,
    telolength,
    FI_high,
    albumin_gL,
    alp,
    bun,
    creat,
    crp,
    glucose,
    hba1c,
    lymph,
    mcv,
    rdw,
    sbp,
    totchol,
    wbc,
    uap
  ) %>%
  filter(education != 9, smoking %in% c(1, 2), !is.na(telolength)) %>%
  mutate(
    gender = factor(gender),
    smoking = factor(smoking),
    race = case_when(race == 2 ~ 1, TRUE ~ race),
    race = factor(race),
    education = factor(education))
```

## Prepare age group column

```{r set_age_groups}
df <- df %>% mutate(
  age_group = case_when(
    age >= 20 & age < 30 ~ "20-30",
    age >= 30 & age < 40 ~ "30-40",
    age >= 40 & age < 50 ~ "40-50",
    age >= 50 & age < 60 ~ "50-60",
    age >= 60 & age < 70 ~ "60-70"
  ), confounders = interaction(gender, smoking)
)
```

# Modelling

## K fold splitting

### Fold-maker function

This function takes the indices of rows each age group, randomizes them, then
assigns numbers of 1-10. We could integrate this into the k-fold cross
validation function, but by doing it separately it's easier to inspect any
possible issues.

```{r fun_make_folds}
make_repeated_folds <- function(data, grouping_variable, k, repeats) {
  # helper function to make random k-fold indices for one repeat
  kfold_split_stratified <- function(k, grouping_variable = NULL) {
    grouping_variable <- as.integer(as.factor(grouping_variable)) # convert factor of age_group to number
    group_len <- length(unique(grouping_variable)) # check number of age groups
    observations <- length(grouping_variable) # check how many observations there are to get a number
    
    indices <- numeric() # empty list to add fold numbers to
    
    for (l in 1:group_len) {
      # run loop for each age group
      indices_grouping_variable <- which(grouping_variable == l) # find which rows have the age group
      indices <- c(indices, sample(indices_grouping_variable)) # randomly shuffle indegrouping_variablees within each age group, so that the numbers 1-10 will equally be spaced throughout the age group, creating even age group distributions in each fold
    }
    
    bins <- rep(observations) # create an output with the length of the number of observations
    bins[indices] <- rep(1:k, ceiling(observations / k))[1:observations] # using the random order, 1-10 is repeated until we've reached the last observation
    return(bins) # return shuffled k-folds, which is pasted into a new column in 'data'
  }
  
  for (i in 1:repeats) {
    # per repeat, make shuffled k-folds of indices in a new column
    data[[paste0("fold_r", i)]] <- kfold_split_stratified(k = k, grouping_variable = data[[grouping_variable]]) # writes k-folds to a new column for each repeat
  }
  
  return(data) # return the dataframe with a k-fold column per repeat
}
```

### Make folds

If k-fold validation throws errors that testing data contains factor levels that
weren't in the training data, it may be needed to make an interaction variable
of age_group with the problematic factor (or condense levels) to ensure the
smaller factor levels are represented in each fold.

```{r make_folds}
set.seed(2)
df <- make_repeated_folds(
  data = df,
  grouping_variable = "age_group",
  k = 10,
  repeats = 10
)
```

### Inspect folds

We check if each fold number is assigned equal amounts of times in two repeats
for each age group. Thereafter, we save the row number and plot the fold number
of points (with some jitter to spread things out) to check if bands appear
homogenous. We know it should be homogenous in the y-axis within a band because
the jitter is random. If there are similarities between bands, then the shuffle
isn't randomized for each repeat. If the bands don't appear of similar intensity
over the x-axis, then the number assignment is not random.

```{r inspect_folds, fig.height = 10, fig.width = 10}
df %>%
  group_by(age_group) %>%
  mutate(total_fold1 = n()) %>%
  group_by(age_group, fold_r1, total_fold1) %>%
  summarize(n_fold1 = n()
            , .groups = "drop") %>% rename("fold" = fold_r1) %>%
  left_join(
    df %>%
      group_by(age_group) %>%
      mutate(total_fold2 = n()) %>%
      group_by(age_group, fold_r2, total_fold2) %>%
      summarize(n_fold2 = n()
                , .groups = "drop") %>% select(age_group, fold_r2, total_fold2, n_fold2, ) %>%
      rename("fold" = fold_r2),
    by = c("age_group", "fold")
  ) %>% kable("html", caption = "Inspect Fold Number Frequencies") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position = "left",
    font_size = 12,
    fixed_thead = TRUE
  ) %>%
  scroll_box(width = "100%", height = "400px")

p <- df %>%
  mutate(row_index = row_number()) %>%
  pivot_longer(cols = starts_with("fold_r"),
               names_to = "r",
               values_to = "fold") %>%
  ggplot(aes(x = row_index, y = fold)) +
  geom_point(
    size = 1,
    alpha = 0.1,
    color = "grey80",
    position = position_jitter(width = 0, height = 0.4)
  ) +
  facet_wrap( ~ r) +
  theme_few()

plot(p)
```

Fold numbers are consistent between groups and correctly split upon multiple
repeats. each fold number has the same (within 1) frequency, so there are no
imbalances between folds and repeats. Plot shows no non-random changes in
intensity over the X dimension and random dissimilarities between bands in the Y
dimension (indicating repeats have different samples). **Stratified fold
randomization succeeded**

## Confounder Analysis

### Check Interaction vs Offsets

We want to get a sense which confounders improve the model and which interact
with age.

```{r assess_confounders}
assess_confounders <- function(df, confounders, k) {
  f0 <- paste0("FI_high ~ s(age, k = ", k, ")")
  m0 <- bam(as.formula(f0), data = df, method = "REML")
  
  results <- data.frame()
  
  for (confounder in confounders) {
    f1 <- paste0("FI_high ~ ", confounder, " + s(age, k = ", k, ")")
    f2 <- paste0("FI_high ~ s(age) + s(age, ", confounder, ", bs = 'sz', k = ",
                 k,
                 ")")
    
    m1 <- bam(as.formula(f1), data = df, method = "REML")
    m2 <- bam(as.formula(f2), data = df, method = "REML")
    
    aov <- as.data.frame(anova(m0, m1, m2))
    aic <- as.data.frame(AIC(m0, m1, m2))
    bic <- as.data.frame(BIC(m0, m1, m2))
    
    results <- rbind(results, cbind(tibble(
      confounder = confounder, formula = c(f0, f1, f2)
    ), aic, bic, aov))
  }
  
  return(results)
}

assess_confounders(df, confounders = c(
  "gender", "smoking", "race", "education"), k = params$global_k)
```

**We see that for each confounder the interaction model fits better than the
intercept-only model**

### Investigate Additional Information

First we see if the confounders are correlated (e.g., smoking is more prevalent
in males than females so gender is partly coded in smoking), to see if all
confounders actually contribute.

```{r compare_added_info}
compare_confounders <- function(df, k) {
  m0 <- bam(FI_high ~ s(age, k = k), data = df, method = "REML")
  m1 <- bam(
    FI_high ~ s(age, k = k) + s(age, gender, bs = "sz", k = k),
    data = df,
    method = "REML"
  )
  m2 <- bam(
    FI_high ~ s(age, k = k) + s(age, gender, bs = "sz", k = k) + s(age, race, bs = "sz", k = k),
    data = df,
    method = "REML"
  )
  m3 <- bam(
    FI_high ~ s(age, k = k) + s(age, gender, bs = "sz", k = k) + s(age, race, bs = "sz", k = k) + s(age, smoking, bs = "sz", k = k),
    data = df,
    method = "REML"
  )
  m4 <- bam(
    FI_high ~ s(age, k = k) + s(age, gender, bs = "sz", k = k) + s(age, race, bs = "sz", k = k) + s(age, smoking, bs = "sz", k = k) + s(age, education, bs = "sz", k = k),
    data = df,
    method = "REML"
  )
  
  print(cbind(
    as.data.frame(AIC(m0, m1, m2, m3, m4)),
    as.data.frame(BIC(m0, m1, m2, m3, m4)),
    as.data.frame(anova(m0, m1, m2, m3, m4))
  ))
  print(summary(m4))
  plot(m4)
}

compare_confounders(df, k = params$global_k)
```

**Each factor significantly improves the model and seems to interact with age in
some way**

### Check shared wiggliness

We could insist that different smooths have the same amount of smoothness, i.e.,
that there's no difference in the wiggliness of the GAM between men and women.
The plots from the previous chunk (where different wiggliness was permitted)
suggest that there may be equal amount of wiggliness, so that each smooth has
the same basis functions but with different coefficients. There's no specific
reason to assume that wiggliness is the same in many cases, so if aiming to
maximize model flexibility/predictive power, we should leave this unconstrained,
but we do eat up a lot of degrees of freedom and may be 'wasting' the data
estimating these global trends. It may also be better to spend the processing
time on a higher k rather than estimating smooth parameters.

```{r compare_wiggliness}
compare_wiggliness <- function(df, k) {
  m1 <- bam(
    FI_high ~
      s(age, k = k) +
      
      s(age, gender, bs = "sz", k = k, id = 1) +
      s(age, race, bs = "sz", k = k, id = 2) +
      s(age, smoking, bs = "sz", k = k, id = 3) +
      s(age, education, bs = "sz", k = k, id = 4),
    data = df,
    method = "REML"
  )
  
  m2 <- bam(
    FI_high ~
      s(age, k = k) +
      
      s(age, gender, bs = "sz", k = k) +
      s(age, race, bs = "sz", k = k) +
      s(age, smoking, bs = "sz", k = k) +
      s(age, education, bs = "sz", k = k),
    data = df,
    method = "REML"
  )
  
  print(cbind(
    as.data.frame(AIC(m1, m2)),
    as.data.frame(BIC(m1, m2)),
    as.data.frame(anova(m1, m2))
  ))
}

compare_wiggliness(df, k = params$global_k)
```

### Check factor three-way interactions

<https://stats.stackexchange.com/questions/637423/conceptual-interpretation-of-bs-fs-and-by-term-in-gam>.
The free wiggliness model takes about 15 times longer to compute but has a 0.37%
drop in residual deviance. That is significant but probably not worth it
compared to a higher k. Moreover, the improvement in fit must stem entirely from
three-way interactions, which are likely noisy, so it is unlikely that out of
fold performance would improve substantially. BIC prefers simpler model and AIC
prefers more complex model.

```{r conf_interactions}
compare_confounder_interactions <- function(df, k, equal_wiggliness, return_model = FALSE) {
  if (equal_wiggliness) {
    m0 <- bam(
      FI_high ~
        s(age, k = k) +
        
        s(gender, age, bs = "sz", k = k, id = 1) +
        s(race, age, bs = "sz", k = k, id = 2) +
        s(smoking, age, bs = "sz", k = k, id = 3) +
        s(education, age, bs = "sz", k = k, id = 4) +
        
        s(gender, race, age, bs = "sz", k = k, id = 5) +
        s(gender, smoking, age, bs = "sz", k = k, id = 6) +
        s(gender, education, age, bs = "sz", k = k, id = 7) +
        s(race, smoking, age, bs = "sz", k = k, id = 8) +
        s(race, education, age, bs = "sz", k = k, id = 9) +
        s(smoking, education, age, bs = "sz", k = k, id = 10),
      data = df,
      method = "REML"
    )
  } else {
    m0 <- bam(
      FI_high ~
        s(age, k = k) +
        
        s(gender, age, bs = "sz", k = k) +
        s(race, age, bs = "sz", k = k) +
        s(smoking, age, bs = "sz", k = k) +
        s(education, age, bs = "sz", k = k) +
        
        s(gender, race, age, bs = "sz", k = k) +
        s(gender, smoking, age, bs = "sz", k = k) +
        s(gender, education, age, bs = "sz", k = k) +
        s(race, smoking, age, bs = "sz", k = k) +
        s(race, education, age, bs = "sz", k = k) +
        s(smoking, education, age, bs = "sz", k = k),
      data = df,
      method = "REML"
    )
  }
      
  print(summary(m0))
  plot(m0)
  
  if (return_model) {
    return(m0)
  }
}

system.time({
compare_confounder_interactions(df, k = params$global_k, equal_wiggliness = TRUE)
})
```

**We see that age \* gender \* race is a significant three-way interaction but
it takes 3 times longer to compute. Only add on a final, very long run if
processing power permits.**

## Repeated k-fold modeling

### Repeated k-fold validation function per age group

```{r fun_CV}
cross_validation <- function(data, formulas, k, repeats, discrete = FALSE, keep_models = TRUE) {
  age_groups = c("20-30", "30-40", "40-50", "50-60", "60-70") # initialize age groups to stratify by explicitly
  fold_columns <- paste0("fold_r", 1:repeats) # initialize names of repeat columns to iterate over
  confounders <- c("gender", "smoking", "race", "education") # initialize vector of confounders
  
  names_of_predictors <- unique(unlist(lapply(formulas, extract_predictors))) # save names of all predictor columns to select for the pred vs obs dataset
  
  results <- data.frame() # initialize empty dataframe
  models <- list() # intialize list of models
  pred_vs_obs <- data.frame() # intialize df with fitted vs predicted
  
  for (formula in formulas) {
    for (r in fold_columns) {
      for (f in 1:k) {
        train <- data %>% filter(.data[[r]] != f) # set training data
        test  <- data %>% filter(.data[[r]] == f) # set testing data
        
        if (discrete) {
          mod <- bam(
            as.formula(formula),
            data = train,
            family = gaussian(),
            method = "fREML",
            discrete = TRUE
          ) # fit GAM
        } else {
          mod <- bam(
            as.formula(formula),
            data = train,
            family = gaussian(),
            method = "REML",
            discrete = FALSE
          ) # fit GAM
        }
        
        predictors =   paste(extract_predictors(formula), collapse = ", ") # extract predictors from formula
        
        rep = as.numeric(sub(".*fold_r(\\d+).*", "\\1", r)) # extract repetition number only
        
        if (keep_models) {
          model_id <- paste0(predictors, "_r", rep, "_f", f)
          models[[model_id]] <- mod # wrap model into a list to keep the models separated, attach it to the list of models, and name the model according to predictors, repeat, and fold
        }
        
        # calculate performance per age group
        for (age_group in age_groups) {
          test_subset <- test[test$age_group == age_group, ]# subset testing data for age group
          train_subset <- train[train$age_group == age_group, ] # subset training data for age group
          
          pred <- predict(mod, newdata = test_subset, type = "response") # use GAM to predict responses in current age group
          
          results <- bind_rows(
            results,
            tibble(
              predictors = predictors,
              age_group = age_group,
              r = rep,
              k = f,
              n_train = nrow(train_subset),
              n_test = nrow(test_subset),
              mse = mean((test_subset$FI_high - pred) ^ 2) # calculate squared difference between predicted vs real in this age group
            )
          )
          
          pred_vs_obs <- bind_rows(
            pred_vs_obs,
            test_subset %>%
              mutate(
                predictors = predictors,
                age_group = age_group,
                k = f,
                r = rep,
                predicted = pred,
              ) %>%
              select(
                predictors,
                r,
                k,
                age_group,
                all_of(names_of_predictors),
                FI_high,
                predicted
              ) # add predictions used to calculate MSE to a dataframe
          )
        }
      }
    }
  }
  
  return(list(
    results = results,
    models = models,
    predictions = pred_vs_obs
  )) # return models, dataframe with results, and a fitted vs predicted dataframe
}
```

### Formula builder function

The `k` determines the number of functions (linear or non-linear).The parameter
with the lowest degrees of freedom (DF) determines the maximum basis dimensions
(k in mgcv). Albumin has 26 distinct levels, so the maximal k value/number of
basic functions, equivalent to the DF, is 25 if each level is in each fold. We
must compute the DF for each fold and pick the lowest to avoid errors. The GAM
is a composite of all these underlying functions, with different weights for
each function.

-   `k`: number of basis functions.
-   `max_k`: sets k to the theoretical limit for each parameter (DF), note that
    for non-integers, like the bioage measures, this will be equivalent to the
    length of the dataframe minus one.
-   `auto_limit_k`: if k exceeds the DF for a parameter (e.g., if k \> 25 for
    albumin), it will automatically be limited to the DF.

```{r fun_formula_gen}
formula_generator <- function(data,
                              formulas,
                              k = 25,
                              max_k = FALSE,
                              auto_limit_k = FALSE,
                              equal_wiggliness = TRUE) {
  fold_columns <- grep("^fold_r", names(data), value = TRUE) # initialize the repeat columns with folds
  folds = sort(unique(unlist(data[fold_columns]), use_name = FALSE)) # get each unique fold number in folds
  formula_list <- list() # initialize a list to store the compiled formulas in
  
  for (formula in formulas) {
    predictors <- trimws(unlist(strsplit(formula, ","))) # unlist, split by comma, and get rid of white spaces for the current predictors string
    sm_terms <- list()
    
    # make a term for each predictor
    for (predictor in predictors) {
      DF_list <- list() # initialize a list to store the DF for each fold in
      
      for (rep in fold_columns) {
        for (fold in folds) {
          DF_list <- append(DF_list, length(unique(data[data[[rep]] != fold, ][[predictor]]))) # store the DF for each fold in each repeat in a list
        }
      }
      
      DF <- min(unlist(DF_list)) - 1 # pick the smallest value in the list and subtract 1 to obtain degrees of freedom
      
      if (max_k) {
        k_formula <- DF
      } else {
        k_formula <- k
      }
      
      if (auto_limit_k) {
        if (k_formula > DF) {
          k_formula <- DF
        }
      }
      if ((predictor == "age") && (equal_wiggliness)) {
        sm_terms <- append(
          sm_terms,
          paste0(
            "s(age, k = ",
            k_formula,
            ") + s(gender, age, bs = 'sz', k = ",
            k_formula,
            ", id = 1) + s(race, age, bs = 'sz', k = ",
            k_formula,
            ", id = 2) + s(smoking, age, bs = 'sz', k = ",
            k_formula,
            ", id = 3) + s(education, age, bs = 'sz', k = ",
            k_formula,
            ", id = 4) + s(gender, race, age, bs = 'sz', k = ",
            k_formula,
            ", id = 5)"
          )
        )
      } else if ((predictor == "age") && !(equal_wiggliness)) {
        sm_terms <- append(
          sm_terms,
          paste0(
            "s(age, k = ",
            k_formula,
            ") + s(gender, age, bs = 'sz', k = ",
            k_formula,
            ") + s(race, age, bs = 'sz', k = ",
            k_formula,
            ") + s(smoking, age, bs = 'sz', k = ",
            k_formula,
            ") + s(education, age, bs = 'sz', k = ",
            k_formula,
            ") + s(gender, race, age, bs = 'sz', k = ",
            k_formula,
            ")"
          )
        )
      }
      else {
        sm_terms <- append(sm_terms,
                           paste0("s(", predictor, ", k = ", k_formula, ")"))
      }
    }
    
    formula_list <- append(formula_list, paste("FI_high ~", paste(unlist(sm_terms), collapse = " + "))) # put predictor terms in formula and separate by +
  }
  
  return(formula_list)
}
```

### Extract variables from formula function

```{r fun_extr_pred}
extract_predictors <- function(formula, collapse = "") {
  confounders <- c("gender", "smoking", "race", "education")
  
  s_terms <- unlist(str_extract_all(formula, "s\\([^\\)]*\\)"))
  keep    <- !str_detect(s_terms, paste0("\\b(", paste(confounders, collapse = "|"), ")\\b"))
  str_match(s_terms[keep], "^s\\(\\s*([^,\\)\\s]+)")[, 2]
}
```

### Generate formulas

```{r build_formulas}
formula_list_bioage <- formula_generator(
  data = df,
  k = params$global_k,
  auto_limit_k = TRUE,
  equal_wiggliness = TRUE,
  formulas = c("age", "age, phenoage", "age, kdm", "age, hd_log", "age, telolength")
)

as.data.frame(cbind(Formulas = unlist(formula_list_bioage))) %>% kable("html", caption = "Bioage Formulas") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position = "left",
    font_size = 12,
    fixed_thead = TRUE
  ) %>%
  scroll_box(width = "100%", height = "400px")
```

### Inspect k-values

To assess k-value, we fit a model on the whole dataset using the formula lists,
build a list of models, make a helper function to run the k.check function from
mgcv, then turn it in a dataframe and return the output. The k-value determines
the number of base functions that the GAM is constructed from. A GAM adds these
different base functions together, some of which capture linear trends, some of
which are just splines or polynomials with a bump in one region. The GAM gives
different weights to each model to get as good a fit as possible, but a GAM also
measures wiggliness and gives a penalty to wiggliness. This is what makes it
generalized, so to perform well in general, even to unseen data. Therefore, a
GAM optimizes to get as good a fit as possible with as little wiggliness as
possible. Because of this wiggliness penalty, adding more base functions to the
GAM doesn't make it overfit. Therefore, k should be as large as possible within
computation power or degrees of freedom. If k isn't large enough, there will be
patterns left in the residuals. Therefore, by measuring patterns in the
residuals, we can assess if k is large enough. We want to capture all global
patterns without fitting to any noise. This is what we assess when we look at
k-index.

For reference, k = 10 takes \~ 15 min, k = 25 takes \~ 1h, k = 50 takes \~ 4h.

```{r inspect_k}
inspect_k_values <- function(data, formulas, subsample, reps) {
  set.seed(2)

  models <- list()
  
  for (formula in formulas) {
    mod <- bam(
      as.formula(formula),
      data = data,
      method = "REML",
      family = gaussian()
    )
    
    predictors = paste(extract_predictors(formula), collapse = ", ")
    
    models[[predictors]] <- mod
  }
  
  inspect_k <- function(model) {
    k.check(model, subsample = subsample, n.rep = reps)
  }
  
  k_output <- lapply(models, inspect_k)
  k_output <- imap_dfr(k_output, ~ {
    as.data.frame(.x) %>%
      rownames_to_column(var = "Predictors") %>%
      mutate(model = .y)
  })
  
  return(k_output)
}

inspect_k_values(df,
                 formula_list_bioage,
                 subsample = 5000,
                 reps = 400) %>% kable("html", caption = "Inspect K Bioage") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position = "left",
    font_size = 12,
    fixed_thead = TRUE
  ) %>%
  scroll_box(width = "100%", height = "400px")
```

### Estimate modeling time

bam speeds up the fitting compared to gam and limits the memory footprint.
Data.table is negligibly faster than data.frame (less than 1%) because the
majority of computing time is spent on fitting the model as opposed to data
operations.

With discrete = FALSE: k = 5, age + phenoage: 0.5 k = 10, age + phenoage: 0.78 k
= 15, age + phenoage: 1.34 k = 20, age + phenoage: 2.76 k = 25, age + phenoage:
4.98 k = 30, age + phenoage: 8.16 k = 40, age + phenoage: 18.28 k = 50, age +
phenoage: 31.97

With discrete = TRUE: k = 5, age + phenoage: 0.04 k = 10, age + phenoage: 0.06 k
= 15, age + phenoage: 0.09 k = 20, age + phenoage: 0.15 k = 25, age + phenoage:
0.24 k = 30, age + phenoage: 0.35 k = 40, age + phenoage: 0.78 k = 50, age +
phenoage: 1.28

```{r check_modeling_time}
estimate_modeling_time <- function(df, k, discrete = FALSE) {
    system.time({
      bam(FI_high ~ 
        s(age, k = k) +
        
        s(gender, age, bs = "sz", k = k, id = 1) +
        s(race, age, bs = "sz", k = k, id = 2) +
        s(smoking, age, bs = "sz", k = k, id = 3) +
        s(education, age, bs = "sz", k = k, id = 4) +
        
        s(gender, race, age, bs = "sz", k = k, id = 5) +

        s(phenoage, k = k),
        data = df, discrete = discrete, method = "fREML"
      )
    })
}

estimate_modeling_time(df, k = params$global_k, discrete = params$global_discrete)
```

**Without id = x, k = 15 took 39.65 theoretically, but in practice the 4 bioage
formulas took 58s each. Similarly, k = 5 with id should have taken \~400s for
one formula but took 650s, so add \~50% time to estimate processing time.**

### Test k-fold validation

Test k-fold validation on a single formula with small k. k = 5, discrete = FALSE
takes 50s, k = 5, discrete = TRUE takes 15s

```{r test_k_fold_val}
formula_list_test <- formula_generator(
  data = df,
  k = params$global_k,
  auto_limit_k = TRUE,
  equal_wiggliness = TRUE,
  formulas = c("age, phenoage")
)

system.time({
  output_bioage <- cross_validation(
    df,
    formula_list_test,
    k = 10,
    repeats = 10,
    discrete = params$global_discrete
  )
})
```

### Model bioage

With k = 40 discrete = FALSE takes 6759s With k = 50 discrete = FALSE takes
13506s

```{r bioage_cv}
system.time({
  output_bioage <- cross_validation(
    df,
    formula_list_bioage,
    k = 10,
    repeats = 10,
    discrete = params$global_discrete,
    keep_models = FALSE
  )
})
```

# Data Processing

## Process Result Comparison dfs

### Reformat df, Calculate Delta MSE Bioage, and set data types

```{r reformat_df}
reformat_df <- function(data) {
  reformatted_data <- data$results %>%
    left_join(
      data$results %>%
        filter(predictors == "age") %>%
        mutate(mse_age = mse) %>%
        select(-mse, -predictors, -n_train, -n_test),
      by = c("age_group", "r", "k")
    ) %>%
    mutate(
      delta_mse = mse_age - mse,
      additional_variance = delta_mse / mse_age * 100,
      fold_repeat_id = interaction(r, k),
      predictors = factor(predictors),
      age_group = factor(age_group),
      r = as.integer(r)
    )
  return(reformatted_data)
}

df_results_bioage <- reformat_df(output_bioage)
```

### Calculate CI

For the overall, non-age-group-stratified summary statistics, we first calculate
mse per fold by weighting the age groups. Then, we calculate the number of
repeats and folds, the mse of each outcome variable (weighted per number of
observations, which produces a small difference over unweighted because the
n_test per fold differs by 1 at most), and the variance of mse `s2_mse`, also
weighted per n_test, grouped by predictors and possibly age group.

The Nadeau Bengio corrected t-test estimates variance using the training and
testing dataset sizes, because cross validations are not independent, violating
the assumption of a paired t-test, inflating type 1 errors (false positives). We
can use the corrected variance estimate for the confidence intervals too. First,
we calculate t-crit using degrees of freedom (n-1). The t-test should be
one-tailed because we only want to know if age + bioage is better than age, not
if one is worse/different, so then a tcrit of 0.05 is appropriate, but for the
confidence intervals we want to know whether the true mean is within 95% of the
sample mean, not just where the true mean would be if it would only be above.
The variance is used based on `n_train` and `n_test`. It doesn't matter whether
we use the mean n per fold or the total over all folds because the ratio is the
same.

In <https://www.jmlr.org/papers/volume6/markatou05a/markatou05a.pdf> Analysis of
Variance of Cross-Validation Estimators of the Generalization Error, the authors
also report the variance estimate as S²\* (1/J + n2/n1), where S2 is the
variance (squared standard deviation), J is the total number of folds, n2 is the
validation/test set and n1 is the training set size. They also show that when J
is appropriate (10\*10), it provides a good variance estimate. Variance is
already in squared units, as its just the square of the standard deviation.

```{r calculate_ci}
calculate_nb_corrected_cis <- function(data, per_age_group) {
  if (per_age_group) {
    grouping_variables <- c("predictors", "age_group")
  } else {
    grouping_variables <- c("predictors")
  }
  
  if (!per_age_group) {
    data <- data %>%
      group_by(predictors, r, k) %>%
      summarize(
        mse = weighted.mean(mse, w = n_test),
        delta_mse = weighted.mean(delta_mse, w = n_test),
        additional_variance = weighted.mean(additional_variance, w = n_test),
        n_train = sum(n_train),
        n_test = sum(n_test),
        .groups = "drop"
      )
  }
  
  summary_data <- data %>%
    group_by(across(all_of(grouping_variables))) %>%
    summarize(
      r = length(unique(r)),
      k = length(unique(k)),
      mean_mse = weighted.mean(mse, w = n_test),
      mean_delta_mse = weighted.mean(delta_mse, w = n_test),
      mean_additional_variance = weighted.mean(additional_variance, w = n_test),
      s2_mse = weightedVar(mse, w = n_test),
      s2_delta_mse = weightedVar(delta_mse, w = n_test),
      s2_additional_variance = weightedVar(additional_variance, w = n_test),
      sum_n_train = sum(n_train),
      sum_n_test = sum(n_test),
      .groups = "drop"
    ) %>%
    mutate(
      n = r * k,
      df = n - 1,
      t_crit = qt(1 - 0.025, df),
      ci_low_mse = mean_mse - t_crit * (sqrt(s2_mse * (
        1 / n + sum_n_test / sum_n_train
      ))),
      ci_high_mse = mean_mse + t_crit * (sqrt(s2_mse * (
        1 / n + sum_n_test / sum_n_train
      ))),
      ci_low_delta_mse = mean_delta_mse - t_crit * (sqrt(
        s2_delta_mse * (1 / n + sum_n_test / sum_n_train)
      )),
      ci_high_delta_mse = mean_delta_mse + t_crit * (sqrt(
        s2_delta_mse * (1 / n + sum_n_test / sum_n_train)
      )),
      ci_low_additional_variance = mean_additional_variance - t_crit * (sqrt(
        s2_additional_variance * (1 / n + sum_n_test / sum_n_train)
      )),
      ci_high_additional_variance = mean_additional_variance + t_crit * (sqrt(
        s2_additional_variance * (1 / n + sum_n_test / sum_n_train)
      ))
    ) %>% select(-r, -k, -s2_mse, -s2_additional_variance, -df, -t_crit)
  
  return(summary_data)
}

df_summary_bioage <- calculate_nb_corrected_cis(df_results_bioage, per_age_group = TRUE)

df_summary_bioage_overall <- calculate_nb_corrected_cis(df_results_bioage, per_age_group = FALSE)
```

### Calculate R2 and RMSE

**Calculating descriptives**: First we import `n_train` and `n_test` from the
results dataframe per fold, either grouped or not grouped by `age_group`. If
grouped by `age_group`, we can import directly because the results dataframe is
already binned by age group. Then, we import the mean fletcher index from the
whole prediction dataset for each model/set of predictors, repeat, and fold. As
a sanity check, because each model should have the same folds and repeats, I
checked whether the mean fletcher index is always the same at repeat 1 fold 1
(it is). The mean FI is then called in the summarize function using median(),
which gives the same result as mean(), because all predicted/real values within
a model, fold, and repeat have the same `mean_FI_high`, as told by the left_join
function. Thereafter, we group over the 100 (10x10) folds per model to calculate
the total number of predictions (on the testing data), the mean training data
size, mean testing data size, R² (using sum of squares), MSE, and RMSE. For
RMSE, we could either calculate it using sqrt(mse) or sqrt(sse/n_predictions).
It gives the same result.

**Total sum of squares calculation when binning R² per age group**: It's
imperative that for the age group stratified descriptives, the global mean is
used. Otherwise it's as if the total sum of squares is calculated within each
age-bin, which is as if a linear regression is fitted through the mean fletcher
index per age group, explaining a lot of variation while it shouldn't. We want
to assess the predictive effect on age, so the total sum of squares should be
calculated by taking the distance from each fletcher index within an age group
to the global mean. Of course, the models are trained on the whole dataset, not
just one age group. We're just assessing performance per age group.

<https://chatgpt.com/share/68da6507-0008-8010-857e-bc6e9b76df58> - see
interpretation of R² differences

```{r calculate_descr}
calculate_RMSE_R2 <- function(summary_data,
                              results_data,
                              predictions,
                              per_age_group) {
  if (per_age_group) {
    grouping_variables <- c("predictors", "age_group")
  } else {
    grouping_variables <- c("predictors")
  }
  
  if (!per_age_group) {
    results_data <- results_data %>%
      group_by(predictors, r, k) %>%
      summarize(
        n_train = sum(n_train),
        n_test = sum(n_test),
        .groups = "drop"
      )
  }
  
  sample_sizes <- results_data %>%
    select(all_of(grouping_variables), r, k, n_train, n_test)
  
  mean_FI_high <- predictions %>%
    group_by(predictors, k, r) %>%
    summarize(mean_FI_high = mean(FI_high), .groups = "drop")
  
  updated_summary_data <- summary_data %>%
    left_join(
      predictions %>%
        left_join(sample_sizes, by = c(grouping_variables, "r", "k")) %>%
        left_join(mean_FI_high, by = c("predictors", "r", "k")) %>%
        group_by(across(all_of(
          grouping_variables
        ))) %>%
        summarize(
          n_predictions = n(),
          mean_n_train = round(mean(n_train), 1),
          mean_n_test = round(mean(n_test), 1),
          sse = sum((FI_high - predicted) ^ 2),
          sst = sum((FI_high - median(mean_FI_high)) ^ 2),
          r2  = 1 - sse / sst,
          rmse = sqrt(sse / n_predictions),
          .groups = "drop"
        ) %>% select(-sse, -sst),
      by = c(grouping_variables)
    )
  
  if (per_age_group) {
    updated_summary_data <- updated_summary_data %>%
      left_join(
        updated_summary_data %>%
          filter(predictors == "age") %>%
          mutate(rmse_age = rmse) %>%
          select(age_group, rmse_age),
        by = c("age_group")
      ) %>%
      mutate(delta_rmse = rmse_age - rmse) %>%
      select(-rmse_age)
  } else {
    updated_summary_data <- updated_summary_data %>% cross_join(
      updated_summary_data %>%
        filter(predictors == "age") %>%
        mutate(rmse_age = rmse) %>%
        select(rmse_age)
    ) %>%
      mutate(delta_rmse = rmse_age - rmse) %>%
      select(-rmse_age)
  }
  
  return(updated_summary_data)
}

df_summary_bioage <- calculate_RMSE_R2(df_summary_bioage,
                                       df_results_bioage,
                                       output_bioage$predictions,
                                       per_age_group = TRUE)

df_summary_bioage_overall <- calculate_RMSE_R2(
  df_summary_bioage_overall,
  df_results_bioage,
  output_bioage$predictions,
  per_age_group = FALSE
)
```

## Repeated Nadeau-Bengio Corrected T-Tests

### Check dependence

Holm error correction is the best method to control family-wise error rate when
comparisons are negatively dependent. However, if tests are either independent
or positively dependent, we can use Hommel, which has more power. To test
dependence, we correlate MSE of each metric over the same fold.

```{r assess_cor}
test_correlation <- function(data, grouping_variables, exclusions = NULL) {
  data <- data %>%
    filter(!predictors %in% exclusions) %>%
    group_by(across(all_of(grouping_variables)), r, k) %>%
    summarize(
      mse = weighted.mean(mse, w = n_test),
      delta_mse = weighted.mean(delta_mse, w = n_test),
      additional_variance = weighted.mean(additional_variance, w = n_test),
      n_train = sum(n_train),
      n_test = sum(n_test),
      .groups = "drop"
    ) %>%
    unite("interaction_column", all_of(grouping_variables), sep = "_") %>%
    mutate(n = (r - 1) * 10 + k) %>%
    select(n, interaction_column, delta_mse) %>%
    pivot_wider(names_from = interaction_column, values_from = delta_mse) %>%
    select(-n)
  
  correlation_df <- as.data.frame(cor(data, use = "pairwise.complete.obs"))
  
  eps = 0.2
  
  upper_indices <- which(upper.tri(correlation_df), arr.ind = TRUE)
  
  df_long <- tibble(
    row = rownames(correlation_df)[upper_indices[, 1]],
    col = colnames(correlation_df)[upper_indices[, 2]],
    values = correlation_df[upper_indices]
  )
  
  p <- ggplot(df_long, aes(values)) +
      geom_density(alpha = 0.15) +
      geom_vline(
        xintercept = c(-eps, eps),
        colour = "red",
        linetype = "dashed"
      ) + theme_few() + labs(title =  paste0("Mean correlation: ", round(
        mean(correlation_df[upper.tri(correlation_df)], na.rm = TRUE), 2
      )))
  
  return(p)
}

test_correlation(df_results_bioage,
                 grouping_variables = "predictors",
                 exclusions = "age")

test_correlation(
  df_results_bioage,
  grouping_variables = c("age_group", "predictors"),
  exclusions = "age"
)
```

All correlations trend positive. Overall bioage methods have positive
dependence, so we can use Hommel, for the other comparisons, correlations center
at 0 and are either mildly positively dependent or independent. There are some
negative correlations but practically all are between -0.25 and 0.25, so there
is probably no violation of the positive regression dependency in which it is
unlikely that an increase in one test statistic would lead to a decrease in
another. This also makes sense for bioage models. If phenoage works, kdm may
also work a little, because they are based on the same blood markers. Similarly,
if we get positive predictive value from one bioage method in one age group, it
probably positively affects predictive power in another age group. For
biomarkers, an increase in predictive power of one marker doesn't per se affect
another. If one of two closely-related biomarkers has higher predictive power,
it doesn't mean the other one won't. The only way to get a negative outcome is
to get overfitting, and I don't see in which scenario a higher predictive power
in one biomarker would increase the odds of overfitting for another biomarker,
so we can assume independence and use Hommel.

### Calculate p-values

Calculate t-statistic according to nadeau bengio, which is t =
m/(√(variance(1/n + n2/n1))). m is the difference between the sample mean and
hypothesized population mean. In this case, we've normalized it to age so the
null hypothesis (the hypothesized population mean) is 0, i.e., there is no
difference from age only. The variance we already calculated. n is equivalent to
the number of folds times the number of repeats. n2 is the number of
testing/validation data and n1 the number of training data. We then calculate
degrees of freedom by taking n - 1. Then we use pt (which automatically
calculates the critical value) using a one-tailed t-test of the upper tail. This
is because we're not interested in whether an age + bioage model is different
from age-only. We only want to know if its better. We then adjust p values with
Hommel or Holm error correction.

We can use Hommel error correction (more power), which makes blood urea nitrogen
significant with k = 10 but not with k = 25.

First, the question is, 'Does this bioage/biomarker add predictive value for
hearing loss in the population while controlling for chronological age?', which
we test with the overall comparisons. Only if it's significantly different do we
look at per age-group tests, also to control power. Another option is to ask
'Might bioage/biomarker only matter for specific age ranges, even if it doesn't
have an overall predictive effect?', in which case we could consider doing
pairwise on all, but we'd lose a lot of power and would have to use FDR and
accept a 5% false discovery rate to salvage it.

The CorrectR library has an implementation of the Nadeau-Bengio corrected t-test
(with support for repeated k-fold validations), but this makes the false
assumption that fold sizes are always equal, so there's a small error in
calculation of the sample mean and variance. The sample mean and variance should
be weighted by fold size. This gets relevant for age-group delta MSEs, because
for the 50-60 age group, some folds have a test size of 99 and others of 98.
That means that some folds weigh 1% more than another. Not correcting for that
could make the difference between significant or not signficant.

```{r t_test_models}
nb_corrected_repeated_t_test <- function(summary_data,
                                         exceptions = NULL,
                                         exception_include_exclude,
                                         adjust_method) {
  if (exception_include_exclude == "include") {
    summary_data <- summary_data %>%
      filter(predictors %in% exceptions)
  } else {
    summary_data <- summary_data %>%
      filter(!predictors %in% exceptions)
  }
  
  summary_data <- summary_data %>%
    mutate(
      t = mean_delta_mse / (sqrt(
        s2_delta_mse * (1 / n + sum_n_test / sum_n_train)
      )),
      df = n - 1,
      p = pt(t, df = df, lower.tail = FALSE),
      p_adj = p.adjust(p, method = adjust_method),
      signif = case_when(
        p_adj <= 0.001 ~ "***",
        p_adj <= 0.01 ~ "**",
        p_adj <= 0.05 ~ "*",
        TRUE ~ "ns"
      )
    ) %>%
    select(-t, -df, -p, -s2_delta_mse, -sum_n_test, -sum_n_train, -n)
  
  return(summary_data)
}

pwc_bioage_overall <- nb_corrected_repeated_t_test(
  df_summary_bioage_overall,
  exceptions = "age",
  exception_include_exclude = "exclude",
  adjust_method = "holm"
)

pwc_bioage <- nb_corrected_repeated_t_test(
  df_summary_bioage,
  exceptions = "age, phenoage",
  exception_include_exclude = "include",
  adjust_method = "holm"
)

pwc_bioage_overall %>% kable("html", caption = "Overall Bioage Pairwise Comparisons") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE,
    position = "left",
    font_size = 12,
    fixed_thead = TRUE
  ) %>%
  scroll_box(width = "100%", height = "400px")
```

### Add p-values to summary dataframe

```{r add_p_vals}
import_p_values <- function(summary_data, pwc_data, per_age_group) {
  if (per_age_group) {
    grouping_variables <- c("predictors", "age_group")
  } else {
    grouping_variables <- c("predictors")
  }
  
  updated_summary_data <- summary_data %>%
    left_join(pwc_data %>%
                select(all_of(grouping_variables), p_adj, signif), by = c(grouping_variables)) %>%
    mutate(
      p_adj = as.character(p_adj),
      p_adj = case_when(signif == "ns" ~ "ns", is.na(p_adj) ~ "ns", TRUE ~ p_adj),
      signif = case_when(is.na(signif) ~ "ns", TRUE ~ signif)
    )
  
  return(updated_summary_data)
}

df_summary_bioage <- import_p_values(df_summary_bioage, pwc_bioage, per_age_group = TRUE)
df_summary_bioage_overall <- import_p_values(df_summary_bioage_overall,
                                             pwc_bioage_overall,
                                             per_age_group = FALSE)
```

**It doesn't matter whether we use holm or hommel, so perhaps we should use
holm.**

### Add raw Ns to dataframe

```{r}
add_n_age_groups <- function(data, summary_data, per_age_group) {
  if (per_age_group) {
    summary_data <- summary_data %>%
      rename("n_folds_repeats" = n) %>%
      left_join(data %>% select(age_group) %>%
                  group_by(age_group) %>%
                  summarize(n = n()),
                by = "age_group")
  } else {
    summary_data <- summary_data %>%
      rename("n_folds_repeats" = n) %>%
      mutate(n = nrow(data))
  }
  
  return(summary_data)
}

df_summary_bioage <- add_n_age_groups(df, df_summary_bioage, per_age_group = TRUE)
df_summary_bioage_overall <- add_n_age_groups(df, df_summary_bioage_overall, per_age_group = FALSE)
```

# Visualization

## Descriptive Tables

### Format table

Round R² and delta RMSE, rename variables, set p values to bold and to " \<
0.001" when very small.

```{r format_table}
format_table <- function(summary_data, per_age_group) {
  if (per_age_group) {
    table <- summary_data %>%
      mutate(r2 = round(r2, 3),
             delta_rmse = round(delta_rmse, 2)) %>%
      rename(
        "Predictors" = predictors,
        "Age" = age_group,
        "∆RMSE (vs age, dB HL)" = delta_rmse,
        "R²" = r2,
        "p (vs age)" = p_adj
      ) %>%
      mutate(
        `p (vs age)` = case_when(
          signif == "***" ~
            cell_spec("< 0.001", "html", bold = TRUE),
          suppressWarnings(!is.na(as.numeric(`p (vs age)`)) &
                             as.numeric(`p (vs age)`) < 0.05) &
            `p (vs age)` != "< 0.001" ~
            cell_spec(suppressWarnings(round(
              as.numeric(`p (vs age)`), 3
            )), "html", bold = TRUE),
          TRUE ~ `p (vs age)`
        ), `p (vs age)` = as.character(`p (vs age)`),
        `p (vs age)` = case_when(Predictors == "age" ~ "",
                                    TRUE ~ `p (vs age)`),
        `∆RMSE (vs age, dB HL)` = as.character(`∆RMSE (vs age, dB HL)`),
        `∆RMSE (vs age, dB HL)` = case_when(Predictors == "age" ~ "",
                                    TRUE ~ `∆RMSE (vs age, dB HL)`)
      ) %>%
      select(Predictors, Age, n, `R²`, `∆RMSE (vs age, dB HL)`, `p (vs age)`)
  } else {
    table <- summary_data %>%
      mutate(r2 = round(r2, 3),
             delta_rmse = round(delta_rmse, 2)) %>%
      rename(
        "Predictors" = predictors,
        "∆RMSE (vs age, dB HL)" = delta_rmse,
        "R²" = r2,
        "p (vs age)" = p_adj
      ) %>%
      mutate(
        `p (vs age)` = case_when(
          signif == "***" ~
            cell_spec("< 0.001", "html", bold = TRUE),
          suppressWarnings(!is.na(as.numeric(`p (vs age)`)) &
                             as.numeric(`p (vs age)`) < 0.05) &
            `p (vs age)` != "< 0.001" ~
            cell_spec(suppressWarnings(round(
              as.numeric(`p (vs age)`), 3
            )), "html", bold = TRUE),
          TRUE ~ `p (vs age)`
        ), `p (vs age)` = as.character(`p (vs age)`),
        `p (vs age)` = case_when(Predictors == "age" ~ "",
                                    TRUE ~ `p (vs age)`),
        `∆RMSE (vs age, dB HL)` = as.character(`∆RMSE (vs age, dB HL)`),
        `∆RMSE (vs age, dB HL)` = case_when(Predictors == "age" ~ "",
                                    TRUE ~ `∆RMSE (vs age, dB HL)`)
      ) %>%
      select(Predictors, n, `R²`, `∆RMSE (vs age, dB HL)`, `p (vs age)`)
  }
  
  return(table)
}

table_bioage <- format_table(df_summary_bioage, per_age_group = TRUE)
table_bioage_overall <- format_table(df_summary_bioage_overall, per_age_group = FALSE)
```

### Order, Filter, and Rename Tables

```{r order_table}
rename_bioage_table <- function(bioage_table) {
  bioage_table <- bioage_table %>%
    mutate(
      Predictors = case_when(
        Predictors == "age" ~ "Age",
        Predictors == "age, hd_log" ~ "Age + HD",
        Predictors == "age, kdm" ~ "Age + KDM",
        Predictors == "age, phenoage" ~ "Age + PhenoAge",
        Predictors == "age, telolength" ~ "Age + Telolength"
      ),
      Predictors = factor(Predictors, levels = c(
        "Age",
        "Age + HD",
        "Age + KDM",
        "Age + PhenoAge",
        "Age + Telolength"
      )
    )) %>%
    arrange(Predictors)
  
  return(bioage_table)
}


table_bioage <- rename_bioage_table(table_bioage)
table_bioage_overall <- rename_bioage_table(table_bioage_overall)
```

### Render table

```{r}
render_and_save_table <- function(table, path) {
  html_table <- table %>%
    kable(
      format = "html",
      align = c("l", "l", "r", "r", "r", "r", "r", "r", "r"),
      escape = FALSE
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE,
      position = "center"
    ) %>%
    collapse_rows(columns = 1, valign = "top")
  
  if (isTRUE(params$save_figures)) {
  save_kable(html_table, path)
  }
  
  return(html_table)
}

render_and_save_table(table_bioage, path = here("Output", "Figures", "FI_high", "Telolength", "table_bioage.html"))
render_and_save_table(table_bioage_overall, path = here("Output", "Figures", "FI_high", "Telolength", "table_bioage_overall.html"))
```

## Bioage panel

### Plot Delta mse

```{r plot_bioage, warning=FALSE}
p <- df_summary_bioage %>%
  ggplot(aes(x = age_group, y = mean_delta_mse))  +
  geom_hline(
    yintercept = 0,
    linetype = "22",
    color = "grey20",
    linewidth = 0.2
  )  +
  geom_ribbon(
    aes(
      fill = predictors,
      group = predictors,
      ymin = ci_low_delta_mse,
      ymax = ci_high_delta_mse
    ),
    alpha = 0.15,
    show.legend = FALSE
  ) +
  geom_line(
    aes(color = predictors, group = predictors),
    linewidth = 0.7,
    alpha = 0.75,
    show.legend = TRUE
  ) +
  geom_point(
    aes(color = predictors),
    size = 1.25,
    alpha = 1,
    show.legend = TRUE
  ) +
  scale_color_manual(
    values = c(
      "age" = "#000000",
      "age, hd_log" = "#256DAB",
      "age, kdm" = "#059748",
      "age, phenoage" = "#CB2027",
      "age, telolength" = "#7B3A96"
    ),
    labels = c("Age", "Age + HD", "Age + KDM", "Age + PhenoAge", "Age + Telolength")
  ) +
  scale_fill_manual(
    values = c(
      "age" = "#8C8C8C",
      "age, hd_log" = "#88BDE6",
      "age, kdm" = "#90CD97",
      "age, phenoage" = "#F07E6E",
      "age, telolength" = "#B276B2"
    ),
    labels = c("Age", "Age + HD", "Age + KDM", "Age + PhenoAge", "Age + Telolength")
  ) +
  guides(
    colour = guide_legend(
      override.aes = list(size = 1.25),
      keywidth = unit(0.3, "cm"),
      keyheight = unit(0.3, "cm"),
      label.theme = element_markdown(
        "sans",
        size = 9,
        margin = margin(l = 3, t = 1.5, unit = "pt")
      ),
    )
  ) +
  scale_y_continuous(breaks = seq(0, 8, 2)) +
  coord_cartesian(ylim = c(-1, 8.5), xlim = c(1.4, 4.6)) +
  labs(x = "Age Group<br><span style='font-size:8pt; color:#4D4D4D'>(Years)</span>", y = "Additional Predictive Capacity<br><span style='font-size:8pt; color:#4D4D4D'>(Δ MSE (dB HL)<sup>2</sup>)</span>") +
  theme_few() +
  theme(
    axis.text.x = element_markdown("sans", size = 8),
    axis.text.y = element_markdown("sans", size = 8),
    axis.title.x = element_markdown("sans", size = 9, lineheight = 1.1),
    axis.title.y = element_markdown("sans", size = 9, lineheight = 1.1),
    legend.title              = element_blank(),
    legend.background         = element_rect(fill = "transparent", color = NA),
    legend.box.background     = element_rect(fill = "transparent", color = NA),
    legend.key                = element_rect(fill = "transparent"),
    legend.position           = "inside",
    legend.position.inside    = c(0.225, 0.79),
    legend.key.spacing.y      = unit(1, "pt"),
    panel.border = element_part_rect(
      colour = "grey30",
      side = "lb",
      linewidth = 0.5
    ),
    plot.margin = margin(
      t = 1,
      r = 1,
      b = 0,
      l = 1,
      unit = "pt"
    )
  ) + geom_text(
    data = . %>%
      filter(p_adj != "ns"),
    aes(
      x = age_group,
      y = ci_high_delta_mse + 0.25,
      label = signif,
      color = predictors
    ),
    manual = TRUE,
    size = 14 / (14 / 5),
    show.legend = FALSE
  ) 

sameplot(
  p,
  file = here("Output", "Figures", "FI_high", "Telolength", "bioage_panel.png"),
  width = params$single_column_width,
  height = 2.15,
  res = 600,
  save = params$save_figures
) 
```